### 2019-10-29 assignment1/submission1.tex
add tex-template

\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage[ngerman]{babel}
\usepackage{amsmath, amsthm, amssymb, amstext}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xparse}
\usepackage{fancyhdr}
\usepackage{algorithmicx}
\usepackage{parskip}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{enumerate}
%\usepackage{minted}
\usepackage {tikz}
\usetikzlibrary {positioning}

\pagestyle{fancy}

\rhead{{\newcommand\and\\\getauthors}}
\author{Felix Bühler\\2973410 \and Clemens Lieb\\xxxxxxx \and Deus\\xxxxxxx \and Fabian Bühler\\xxxxxxx}
\lhead{\textbf\gettitle}
\title{\gettitle}
\chead{\getsubtitle}
\subtitle{\getsubtitle}

\addtolength{\headheight}{2\baselineskip}
\renewcommand{\headrulewidth}{0pt}

\newcommand{\gettitle}{Verteilte Systeme I\\Winter Term 2019/20}
\newcommand{\getsubtitle}{G2T1 – Assignment 1 (theoretical part)}
\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Deus \and Fabian Bühler}
\setlength{\headheight}{53pt}

\begin{document}
\maketitle

\section*{1 Transparency Levels}
\subsection*{a)}
\subsection*{b)}
\subsection*{c)}
\subsection*{d)}
\subsubsection*{i.}
\subsubsection*{ii.}

\section*{2 System Models}
\subsection*{a)}
\subsection*{b)}
\subsubsection*{i.}
\subsubsection*{ii.}

\section*{3 Three-Army-Problem}
\subsection*{a)}
\subsection*{b)}

\section*{4 System Availability}
\subsection*{a)}
\subsection*{b)}
\subsection*{c)}
\subsection*{d)}

\end{document}

### 2019-10-30 assignment1/submission1.tex
started 1.4
Aufgabe 1 und Aufgabe 2 eingefügt

@@
 \rhead{{\newcommand\and\\\getauthors}}
-\author{Felix Bühler\\2973410 \and Clemens Lieb\\xxxxxxx \and Deus\\xxxxxxx \and Fabian Bühler\\xxxxxxx}
+\author{Felix Bühler\\2973410 \and Clemens Lieb\\xxxxxxx \and Steffen Wonner\\2862123 \and Fabian Bühler\\xxxxxxx}
 \lhead{\textbf\gettitle}
@@
 \subsection*{a)}
+The access is location transparent, because the location is hidden behinde an unresolved URL. 
 \subsection*{b)}
+The service is not replication transparent because ther is a unique link for every location that has to be checked separatly.
 \subsection*{c)}
+The access is replication transparent, because it is not apparent if Otto checking one or multiple web services. 
 \subsection*{d)}
 \subsubsection*{i.}
+It is only possible to write all copys at once so all copys are allways the same. That means independent from when or which copy is read the answer is allways up to date.
 \subsubsection*{ii.}
+Write operations are performed on n-1 copies.
+Read operations are performed on 2 copies, the client acceptes the new version.
 
@@
 \subsection*{a)}
+\begin{itemize}
+	\item No message is lost
+	\item The maximal possible delay is known
+\end{itemize}
 \subsection*{b)}
 \subsubsection*{i.}
+Every process i sends its own messages at time i * δt:
+t\textsubscript{i}\textsuperscript{send}(m) = i * δt
+The tranmission of a given message is not longer than δt, so if a process i starts sending after i * δt every process with a lower id has allready finished and all processes know there messages.
 \subsubsection*{ii.}
+The worst-case is n * δt as δt is reserved for every prozess as time to send und δt is the loangest time a prozess needs for transmitting a message
 
@@
 \subsection*{a)}
+$ A_x = \dfrac{80t}{100t} = 80\% $
+
+$ A_y = \dfrac{60t}{100t} = 60\%$
 \subsection*{b)}
+
+$ A_{A_x|A_y} = \dfrac{80t}{100t} = 80\% $
 \subsection*{c)}
+$ P(A_i, A_j) = $ observing node $ A_i $ as up given that node $ A_j $ is up
+
+$ P(A_x|A_y) = \dfrac{60t}{60t} = 100\% \rightarrow$ not dependent
+
+$ P(A_y|A_x) = \dfrac{60t}{80t} = 75\% \rightarrow$ is dependent
 \subsection*{d)}
+The availability depends on $ A_x = 80\% $ which is equal to $ A_{A_x|A_y} $.
 

### 2019-10-30 assignment1/submission1.tex
fix deltas in 1.2

@@
 \subsubsection*{i.}
-Every process i sends its own messages at time i * δt:
-t\textsubscript{i}\textsuperscript{send}(m) = i * δt
-The tranmission of a given message is not longer than δt, so if a process i starts sending after i * δt every process with a lower id has allready finished and all processes know there messages.
+Every process i sends its own messages at time $ i * \delta t$:
+t\textsubscript{i}\textsuperscript{send}(m) = $ i * \delta t $
+The tranmission of a given message is not longer than $ \delta $t, so if a process i starts sending after i * $ \delta $t every process with a lower id has allready finished and all processes know there messages.
 \subsubsection*{ii.}
-The worst-case is n * δt as δt is reserved for every prozess as time to send und δt is the loangest time a prozess needs for transmitting a message
+The worst-case is n * $ \delta $t as $ \delta $t is reserved for every prozess as time to send und $ \delta $t is the loangest time a prozess needs for transmitting a message
 

### 2019-10-30 assignment1/submission1.tex
add steffen to authors on new pages
Correct 1.4 (and wrap answers in enumerate)

@@
 \rhead{{\newcommand\and\\\getauthors}}
-\author{Felix Bühler\\2973410 \and Clemens Lieb\\xxxxxxx \and Steffen Wonner\\2862123 \and Fabian Bühler\\xxxxxxx}
+\author{Felix Bühler\\2973410 \and Clemens Lieb\\xxxxxxx \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
 \lhead{\textbf\gettitle}
@@
 \newcommand{\getsubtitle}{G2T1 – Assignment 1 (theoretical part)}
-\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Deus \and Fabian Bühler}
+\newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
 \setlength{\headheight}{53pt}
@@
 \section*{4 System Availability}
-\subsection*{a)}
-$ A_x = \dfrac{80t}{100t} = 80\% $
+
+\begin{enumerate}[a)]
+	\item $ A_x = \dfrac{80t}{100t} = 80\% $
 
 		$ A_y = \dfrac{60t}{100t} = 60\%$
-\subsection*{b)}
 
-$ A_{A_x|A_y} = \dfrac{80t}{100t} = 80\% $
-\subsection*{c)}
-$ P(A_i, A_j) = $ observing node $ A_i $ as up given that node $ A_j $ is up
+	\item Overall statistical availability:
 
-$ P(A_x|A_y) = \dfrac{60t}{60t} = 100\% \rightarrow$ not dependent
+		$ A_S = 1 - P(Y=\text{down} \cup X=\text{down}) = 1 - ((1 - A_x) \cdot (1 - A_y)) = 1 - (0.2 \cdot 0.4) = 1 - 0.08 = 0.92 = 92\%$
 
-$ P(A_y|A_x) = \dfrac{60t}{80t} = 75\% \rightarrow$ is dependent
-\subsection*{d)}
-The availability depends on $ A_x = 80\% $ which is equal to $ A_{A_x|A_y} $.
+		Availability in the concrete example:
+
+		$ A_S = \dfrac{80t}{100t} = 80\% $
+
+	\item $ P(Y=\text{up}|X=\text{up}) = \frac{P(Y=\text{up} \cap X=\text{up})}{P(X=\text{up})} = \frac{0.6}{0.8} = 75\% $
+
+		$ P(Y=\text{up}|X=\text{down}) = \frac{P(Y=\text{up} \cap X=\text{down})}{P(X=\text{down})} = \frac{0}{0.2} = 0\% $
+
+		Node $Y$ seems to depend on Node $X$ to be available.
+
+	\item The availability of the system ($ A_S $) depends on the node $X$.
+
+		$ A_S = P(Y=\text{up} \cup X=\text{up}) = P(X=\text{up}) = A_X = 0.8 $
+
+		This is different from the result in b) ($ A_S = 0.92 $) where the assumption was that the nodes have independent failures.
+		If the nodes are dependent then the overall system can not achieve a higher availability than any one of the nodes.
+
+\end{enumerate}
 

### 2019-11-02 assignment1/submission1.tex
Correct spelling for assignment 1.1

@@
 \rhead{{\newcommand\and\\\getauthors}}
-\author{Felix Bühler\\2973410 \and Clemens Lieb\\xxxxxxx \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
+\author{Felix Bühler\\2973410 \and Clemens Lieb\\3130838 \and Steffen Wonner\\2862123 \and Fabian Bühler\\2953320}
 \lhead{\textbf\gettitle}
@@
 \subsection*{a)}
-The access is location transparent, because the location is hidden behinde an unresolved URL.
+The access is location transparent, because the location is hidden behind an unresolved URL.
+The actual adress of the webservice is automatically resolved through DNS acting as service discovery mechanism.
 \subsection*{b)}
-The service is not replication transparent because ther is a unique link for every location that has to be checked separatly.
+The service is not replication transparent because there is a unique name for every location that has to be checked separatly.
 \subsection*{c)}
-The access is replication transparent, because it is not apparent if Otto checking one or multiple web services.
+The access is replication transparent, because it is not apparent whether Otto checks one or multiple web services.
+It's furthermore unknown which of the two replication locations did provide the file in the end.
 \subsection*{d)}
 \subsubsection*{i.}
-It is only possible to write all copys at once so all copys are allways the same. That means independent from when or which copy is read the answer is allways up to date.
+It is only possible to write all copies at once so all copies are always the same.
+This iplies that independently from when or which copy is read the answer is always up to date.
 \subsubsection*{ii.}
-Write operations are performed on n-1 copies.
-Read operations are performed on 2 copies, the client acceptes the new version.
+Write operations are performed on \(n-1\) copies, as required by the locking constraints.
+Read operations are performed on \(2\) different randomly selected copies, the client accepts the "latest available" version.
 

### 2019-11-02 assignment1/submission1.tex
Change 1.2 to account for clock-skew

@@
 	\item No message is lost
-	\item The maximal possible delay is known
+	\item The maximum possible delay is known
+	\item There is a limited amount of clock skew between the processes
 \end{itemize}
@@
 \subsubsection*{i.}
-Every process i sends its own messages at time $ i * \delta t$:
-t\textsubscript{i}\textsuperscript{send}(m) = $ i * \delta t $
-The tranmission of a given message is not longer than $ \delta $t, so if a process i starts sending after i * $ \delta $t every process with a lower id has allready finished and all processes know there messages.
+Assuming a limited clock skew of \(\delta c\) between the processes, the process \(P_i\) sends a broadcast message \(2 * \delta c + \delta t\) time units after receiving \(i - 1\) messages.
+This allows for instant delivery to a process skewed by the maximum amount of time ''forwards'' while keeping the order of messages intact when sending the message even with the maximum delay \(\delta t\) to a process skewed by the maximum amount of time ''backwards''.
 \subsubsection*{ii.}
-The worst-case is n * $ \delta $t as $ \delta $t is reserved for every prozess as time to send und $ \delta $t is the loangest time a prozess needs for transmitting a message
+In the worst case, the synchronization protocol takes \(2n (\delta t + \delta c) + 2\delta c\) time units.
+This assumes that all messages take the maximum possible amount of time to deliver, namely \(n delta t\) overall, with clock skew cancelling itself out across all processes.
+Adding the wait-time for each process \(n(\delta t + 2\delta c)\) yields \(2n (\delta t + \delta c)\).
+The final component is assuming the maximum clock skew between \(P_1 \text{ and } P_n\).
 

### 2019-11-02 assignment1/submission1.tex
Work on assignment 1.3

@@
 \subsection*{a)}
+To enable a resolution the failure model should exclude both crashing and byzantine failures.
+General Omissions, so long as their occurences are bounded, can be resolved by retrying communication sufficiently often.
 \subsection*{b)}
+\par{Phase 1: Determining order of attack}
+The initiating division sends a messenger with it's own size and the time \(t_0\) to both other divisions.
+A division receiving a first message with a division size sends it's division size to the two other divisions.
+If it has received two messages, it calculates the order of attack.
+If it is the largest division it initiates Phase 2.
+\par{Phase 2: Attack coordination}
+The largest division is definitively known at \(t_1 = t_0 + r, r \leq 2d\).
+The largest division attacks at \(t_1\) and sends a message with that time to both other divisions.
+The second largest division attacks at time \(t_2 = t_1 + k\), the smallest division attacks at \(t_3 = t_1 + 2k = t_2 + k\).
+
+Since all clocks are perfectly synchronized we only need to consider message delivery times.
+\(t_1\) is minimal, because it's the first moment that the largest division is known to be the largest.
+Since all divisions must have sent their size \emph{before} \(t_1\) -- otherwise the information would not be available -- all divisions know their position in the order of attack at \(t_1 + d < t_1 + k\), which implies that \(t_2 > t_1 + d\) as required to allow the second division to attack at the specified time.
+And because all attacks must be at least \(k\) apart, \(2k\) is the minimal span of time between the first and last attack.
+
+\begin{tikzpicture}
+	% division identifiers
+	\node[draw, circle] at (0, 5)   (D1) {\(D_1\)};
+	\node[draw, circle] at (0, 2.5) (D2) {\(D_2\)};
+	\node[draw, circle] at (0, 0)   (D3) {\(D_3\)};
+	% division timelines
+	\path[draw,->]
+		(D1) -- (12, 5)
+		(D2) -- (12, 2.5)
+		(D3) -- (12, 0);
+	% message 1 at t_0
+	\node[draw, circle, fill=black] at (1, 5) (M1) {};
+	\node[above=1.2, right=0.2]     at (1, 5) {\(t_0\)};
+	% message 2 (coming from d3)
+	\node[draw, circle, fill=black] at (2, 0) (M2) {};
+	% message 3 (coming from d2)
+	\node[draw, circle, fill=black] at (3, 2.5) (M3) {};
+	\node[above=1]                  at (3, 2.5) {\(t_1\)};
+	% attack message (assuming d1 is largest and last to know)
+	\node[draw, circle, fill=red] at (5, 5) (MA) {};
+	% attack timings
+	\node[draw, circle, fill=red] at (7.5, 2.5) {};
+	\node[draw, circle, fill=red] at (10,  0) {};
+	% message delivery
+	\path[draw, ->]
+		(M1) edge node[above, sloped] {\(\leq d\)} (M3)
+		(M1) edge node[above, sloped] {\(\leq d\)} (M2)
+		(M2) edge node[below, sloped] {\(\leq d\)} (4.5, 2.5)
+		(M2) edge node[below, sloped] {\(\leq d\)} (MA)
+		(M3) edge node[above, sloped] {\(\leq d\)} (4, 5)
+		(M3) edge node[above, sloped] {\(\leq d\)} (4, 0)
+		(MA) edge node[above, sloped] {\(\leq d\)} (6.5, 0)
+		(MA) edge node[above, sloped] {\(\leq d\)} (7, 2.5)
+		;
+	\path[draw, dashed]
+		% t_1 possible area
+		(4, 6) -- (4, -1)
+		(5, 6) -- (5, -1)
+		% t_2
+		(7.5, 6) -- (7.5, -1)
+		% t_3
+		(10, 6) -- (10, -1)
+		;
+	\node at (4, 6) {\parbox{1.8cm}{Range of possible \(t_1\)}};
+	\node at (5, -1.5) {\(t_1\)};
+	\node at (7.5, -1.5) {\(t_2\)};
+	\node at (10, -1.5) {\(t_3\)};
+
+\end{tikzpicture}
 

### 2019-11-02 assignment1/submission1.tex
Correct 1.1, 1.2 and 1.3 + small formatting improvements

@@
 The access is location transparent, because the location is hidden behind an unresolved URL.
-The actual adress of the webservice is automatically resolved through DNS acting as service discovery mechanism.
+The actual address of the webservice is automatically resolved through DNS acting as service discovery mechanism.
+Also the actual song is referenced by title and not by a specific location.
 \subsection*{b)}
-The service is not replication transparent because there is a unique name for every location that has to be checked separatly.
+The service is not replication transparent because there is a unique name for every location that has to be checked separately.
+The client has to know all available replications for this to work.
 \subsection*{c)}
@@
 \subsection*{d)}
-\subsubsection*{i.}
-It is only possible to write all copies at once so all copies are always the same.
-This iplies that independently from when or which copy is read the answer is always up to date.
-\subsubsection*{ii.}
-Write operations are performed on \(n-1\) copies, as required by the locking constraints.
-Read operations are performed on \(2\) different randomly selected copies, the client accepts the "latest available" version.
+\begin{enumerate}[i.]
+	\item It is only possible to write to all copies at once so all copies will always have the same value.
+		This implies that independently from when or which copy is read the answer is always up to date.
+	\item Write operations are performed on \(n-1\) copies, as required by the locking constraints (only locks for these \(n-1\) copies are acquired).
+		A write operation also marks the time when the value was updated.
+		Read operations are performed on \(2\) different randomly selected copies, the client accepts the "newer" version according to the timestamps of the updates.
+		The reading operation only needs to lock the two selected copies.
+\end{enumerate}
 
@@
 \begin{itemize}
+	\item There are no node failures
 	\item No message is lost
-	\item The maximum possible delay is known
+	\item The maximum possible delay has a known upper bound
 	\item There is a limited amount of clock skew between the processes
@@
 \subsection*{b)}
-\subsubsection*{i.}
-Assuming a limited clock skew of \(\delta c\) between the processes, the process \(P_i\) sends a broadcast message \(2 * \delta c + \delta t\) time units after receiving \(i - 1\) messages.
+\begin{enumerate}[i.]
+	\item Assuming an upper limit \(s\) of the clock skew between the processes, the process \(P_i\) sends a broadcast message \(2 \cdot s \cdot \delta t + \delta t\) time units after receiving \(i - 1\) messages.
+		The first process sends immediately.
+
 		This allows for instant delivery to a process skewed by the maximum amount of time ''forwards'' while keeping the order of messages intact when sending the message even with the maximum delay \(\delta t\) to a process skewed by the maximum amount of time ''backwards''.
-\subsubsection*{ii.}
-In the worst case, the synchronization protocol takes \(2n (\delta t + \delta c) + 2\delta c\) time units.
-This assumes that all messages take the maximum possible amount of time to deliver, namely \(n delta t\) overall, with clock skew cancelling itself out across all processes.
-Adding the wait-time for each process \(n(\delta t + 2\delta c)\) yields \(2n (\delta t + \delta c)\).
-The final component is assuming the maximum clock skew between \(P_1 \text{ and } P_n\).
+	\item In the worst case, the synchronization protocol takes \(n \cdot \delta t + (n-1)*(2 \cdot s \cdot \delta t + \delta t)\) time units.
+		The first part \(n \cdot \delta t\) is the maximum possible network delay.
+		The second part \((n-1)*(2 \cdot s \cdot \delta t + \delta t)\) is the waiting time of all processes except the first one that does not have to wait.
+		Clock skew can be ignored here as we do not compare the timestamps of the first send with the last receive.
+\end{enumerate}
 
@@
 To enable a resolution the failure model should exclude both crashing and byzantine failures.
-General Omissions, so long as their occurences are bounded, can be resolved by retrying communication sufficiently often.
+General Omissions, so long as their occurrences are bounded, can be resolved by retrying communication sufficiently often.
+Message delivery also has to have an upper bound on the delay.
 \subsection*{b)}
@@
 \par{Phase 2: Attack coordination}
-The largest division is definitively known at \(t_1 = t_0 + r, r \leq 2d\).
+The largest division receives the second message at \(t_1 = t_0 + r, r \leq 2d\).
 The largest division attacks at \(t_1\) and sends a message with that time to both other divisions.
+The new message (and any outstanding messages from the first phase) will arrive before the other divisions need to attack as \(k > d\).
 The second largest division attacks at time \(t_2 = t_1 + k\), the smallest division attacks at \(t_3 = t_1 + 2k = t_2 + k\).
@@
 Since all clocks are perfectly synchronized we only need to consider message delivery times.
-\(t_1\) is minimal, because it's the first moment that the largest division is known to be the largest.
+\(t_1\) is minimal, because it's the first moment when the largest division knows it is the largest.
 Since all divisions must have sent their size \emph{before} \(t_1\) -- otherwise the information would not be available -- all divisions know their position in the order of attack at \(t_1 + d < t_1 + k\), which implies that \(t_2 > t_1 + d\) as required to allow the second division to attack at the specified time.
+This holds especially for \(t_3 > t_2\).
 And because all attacks must be at least \(k\) apart, \(2k\) is the minimal span of time between the first and last attack.

### 2019-11-04 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/client/CalcSocketClient.java
add ex2-practical

@@ -1,0 +1,63 @@
+package de.unistgt.ipvs.vs.ex1.client;
+
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.OutputStream;
+import java.net.Socket;
+import java.net.UnknownHostException;
+
+/**
+ * Implement the connectTo-, disconnect-, and calculate-method of this class
+ * as necessary to complete the assignment. You may also add some fields or methods.
+ */
+public class CalcSocketClient {
+	private Socket cliSocket;
+	private int    rcvdOKs;		// --> Number of valid message contents
+	private int    rcvdErs;		// --> Number of invalid message contents
+	private int    calcRes;		// --> Calculation result (cf.  'RES')
+	
+	public CalcSocketClient() {
+		this.cliSocket = null;
+		this.rcvdOKs   = 0;
+		this.rcvdErs   = 0;
+		this.calcRes   = 0;
+	}
+	
+	public int getRcvdOKs() {
+		return rcvdOKs;
+	}
+
+	public int getRcvdErs() {
+		return rcvdErs;
+	}
+
+	public int getCalcRes() {
+		return calcRes;
+	}
+
+	public boolean connectTo(String srvIP, int srvPort) {
+               
+		//Solution here
+		
+		return true;
+	}
+
+	public boolean disconnect() {
+               
+	    //Solution here
+               
+		return true;
+	}
+
+	public boolean calculate(String request) {
+               
+		if (cliSocket == null) {
+			System.err.println("Client not connected!");
+			return false;
+		}
+		
+		//Solution here
+		
+		return true;
+	}
+}

### 2019-11-04 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/server/CalcSocketServer.java

@@ -1,0 +1,49 @@
+package de.unistgt.ipvs.vs.ex1.server;
+
+import java.io.IOException;
+import java.net.ServerSocket;
+import java.net.Socket;
+
+/**
+ * Extend the run-method of this class as necessary to complete the assignment.
+ * You may also add some fields, methods, or further classes.
+ */
+public class CalcSocketServer extends Thread {
+	private ServerSocket srvSocket;
+	private int port;
+
+	public CalcSocketServer(int port) {
+		this.srvSocket = null;
+		
+		this.port = port;
+	}
+	
+	@Override
+	public void interrupt() {
+		try {
+			if (srvSocket != null) srvSocket.close();
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
+	}
+
+	@Override
+	public void run() {
+           
+		if (port <= 0) {
+			System.err.println("Wrong number of arguments.\nUsage: SocketServer <listenPort>\n");
+			System.exit(-1);
+		}
+
+		// Start listening server socket ..
+	}
+        
+        public void waitUnitlRunnig(){
+            while(this.srvSocket == null){
+                try {
+                    Thread.sleep(1);
+                } catch (InterruptedException ex) {
+                }
+            }
+        }
+}

### 2019-11-04 assignment2/handout/src/test/java/de/unistgt/ipvs/vs/ex1/TestCalc.java

@@ -1,0 +1,96 @@
+package de.unistgt.ipvs.vs.ex1;
+
+import de.unistgt.ipvs.vs.ex1.client.CalcSocketClient;
+import de.unistgt.ipvs.vs.ex1.server.CalcSocketServer;
+import java.io.IOException;
+import static org.junit.Assert.*;
+import org.junit.Rule;
+
+import org.junit.Test;
+import org.junit.rules.Timeout;
+
+
+public class TestCalc {
+    
+    @Rule
+    public Timeout globalTimeout = Timeout.seconds(30);
+    
+    @Test
+    public void test1() {
+         // Start Client
+        String srvIP   = "localhost"; //"127.0.0.1";
+        int      srvPort = 12345;
+
+        // Start Server
+        CalcSocketServer cSrv = new CalcSocketServer(srvPort);
+        cSrv.start();
+        cSrv.waitUnitlRunnig();
+
+        CalcSocketClient cCli = new CalcSocketClient();
+        cCli.connectTo(srvIP, srvPort);
+
+        //Test 
+        String req1 = "ADD 1 2 3 SUB 3 2 1";
+        cCli.calculate("<" + (req1.length() + 5) + ":" + req1 + ">");
+        cCli.calculate("<08:rEs>");
+
+        assertEquals(0,cCli.getCalcRes());
+        assertEquals(11,cCli.getRcvdOKs());
+
+        cCli.disconnect();
+    }
+
+    @Test
+    public void test2() {
+         // Start Client
+        String srvIP = "localhost"; //"127.0.0.1";
+        int srvPort = 12346;
+
+        // Start Server
+        CalcSocketServer cSrv = new CalcSocketServer(srvPort);
+        cSrv.start();
+        cSrv.waitUnitlRunnig();
+
+        CalcSocketClient cCli = new CalcSocketClient();
+        cCli.connectTo(srvIP, srvPort);
+
+        //Test 
+        String req1 = "Add -3 ASM -2 ABC -1 SUB -1 ASM10 -2 ABC09 -3";
+        cCli.calculate("<" + (req1.length() + 5) + ":" + req1 + ">");
+        cCli.calculate("<08:rEs>");
+
+        assertEquals(0, cCli.getCalcRes());
+
+        cCli.disconnect();
+    }
+
+    @Test
+    public void test3() {
+         // Start Client
+        String srvIP   = "localhost"; //"127.0.0.1";
[…]

### 2019-11-05 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex2/client/CalcRmiClient.java
initial 2.2, but no factory is used

@@ -2,6 +2,7 @@ package de.unistgt.ipvs.vs.ex2.client;
 
 import de.unistgt.ipvs.vs.ex2.common.ICalculation;
 import de.unistgt.ipvs.vs.ex2.common.ICalculationFactory;
+
 import java.net.MalformedURLException;
 import java.rmi.Naming;
 import java.rmi.NotBoundException;
@@ -14,21 +15,54 @@ import java.util.Collection;
  * necessary to complete the assignment. You may also add some fields or methods.
  */
 public class CalcRmiClient {
-	private ICalculation calc = null;
+    private ICalculation calc = null;
 
-	public CalcRmiClient() {
-		this.calc = null;
-	}
+    public CalcRmiClient() {
+        this.calc = null;
+    }
 
-	public int getCalcRes() {
-		return -1;
-	}
+    public int getCalcRes() {
+        try {
+            return calc.getResult();
+        } catch (RemoteException e) {
+            e.printStackTrace();
+        }
+        return -1;
+    }
 
-	public boolean init(String url) {
-		return true;
-	}
+    public boolean init(String url) {
+        try {
+            calc = (ICalculation) Naming.lookup(url);
+        } catch (Exception e) {
+            System.out.println("calc is not set, therefore everything will crash");
+            e.printStackTrace();
+        }
+        return true;
+    }
 
-	public boolean calculate(CalculationMode calcMode, Collection<Integer> numbers) {
-		return true;
-	}
+    public boolean calculate(CalculationMode calcMode, Collection<Integer> numbers) {
+        try {
+            switch (calcMode) {
+                case ADD:
+                    for (int n : numbers) {
+                        calc.add(n);
+                    }
+                    break;
+                case MUL:
+                    for (int n : numbers) {
+                        calc.multiply(n);
+                    }
+                    break;
+                case SUB:
+                    for (int n : numbers) {
+                        calc.subtract(n);
+                    }
+                    break;
+            }
+            return true;
+        } catch (RemoteException e) {
+            e.printStackTrace();
+        }
+        return false;
+    }
 }

### 2019-11-05 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex2/server/CalcRmiServer.java

@@ -1,10 +1,13 @@
 package de.unistgt.ipvs.vs.ex2.server;
 
+import de.unistgt.ipvs.vs.ex2.common.ICalculation;
 import de.unistgt.ipvs.vs.ex2.common.ICalculationFactory;
+
 import java.rmi.Naming;
 import java.rmi.RemoteException;
 import java.net.MalformedURLException;
 import java.rmi.NotBoundException;
+import java.rmi.registry.LocateRegistry;
 import java.util.logging.Level;
 import java.util.logging.Logger;
 
@@ -13,35 +16,47 @@ import java.util.logging.Logger;
  * the assignment. You may also add some fields or methods.
  */
 public class CalcRmiServer extends Thread {
-	private String regHost;
-	private String objName;
-        String url; //Please use this variable to bind the object.
-	
-	public CalcRmiServer(String regHost, String objName) {
-		this.regHost = regHost;
-		this.objName = objName;
-                this.url = "rmi://" + regHost + "/" + objName;
-	}
-	@Override
-	public void run() {
-		if (regHost == null || objName == null) {
-			System.err.println("<registryHost> or <objectName> not set!");
-			return;
-		}
-                
-                //Add solution here
-	}
+    private String regHost;
+    private String objName;
+    String url; //Please use this variable to bind the object.
 
-        public void stopServer(){
-            try {
-                Naming.unbind(url);
-            } catch (RemoteException ex) {
-                Logger.getLogger(CalcRmiServer.class.getName()).log(Level.SEVERE, null, ex);
-            } catch (NotBoundException ex) {
-                Logger.getLogger(CalcRmiServer.class.getName()).log(Level.SEVERE, null, ex);
-            } catch (MalformedURLException ex) {
-                Logger.getLogger(CalcRmiServer.class.getName()).log(Level.SEVERE, null, ex);
-            }
+    public CalcRmiServer(String regHost, String objName) {
+        this.regHost = regHost;
+        this.objName = objName;
+        this.url = "rmi://" + regHost + "/" + objName;
+    }
+
+    @Override
+    public void run() {
+        if (regHost == null || objName == null) {
+            System.err.println("<registryHost> or <objectName> not set!");
+            return;
         }
-        
+        //Add solution here
+        try {
+            LocateRegistry.getRegistry();
+        } catch (RemoteException e) {
+            // should not happen, otherwise port is already used
+        }
+        try {
[…]

### 2019-11-05 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex2/server/CalculationImplFactory.java

@@ -1,15 +1,13 @@
 package de.unistgt.ipvs.vs.ex2.server;
 
 
-
 /**
  * Change this class (implementation/signature/...) as necessary to complete the assignment.
  * You may also add some fields or methods.
  */
 
- public class CalculationImplFactory  {
- 	private static final long serialVersionUID = 8409100566761383094L;
- 
- 	
- 
- }
+public class CalculationImplFactory {
+    private static final long serialVersionUID = 8409100566761383094L;
+
+
+}

### 2019-11-10 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/server/CalculationSession.java
Tippfehler gefixed

@@ -28,7 +28,7 @@ public class CalculationSession implements Runnable {
 			ObjectInputStream in = new ObjectInputStream(this.cliSocket.getInputStream());
 
 			// Send RDY when ready
-			oosOut.writeObject("<08:RDY>");
+			out.writeObject("<08:RDY>");
 
 			ICalculation calc = new CalculationImpl();
 			// TOOD

### 2019-11-11 assignment2/submission2.tex
add assignment2-theory

--- 2019-11-02 assignment1/submission1.tex
@@
 %\usepackage[T1]{fontenc}
-\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
+\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=4cm]{geometry}
 \usepackage[ngerman]{babel}
@@
 \usepackage{enumerate}
+\usepackage{enumitem}
 %\usepackage{minted}
@@
 \newcommand{\gettitle}{Verteilte Systeme I\\Winter Term 2019/20}
-\newcommand{\getsubtitle}{G2T1 – Assignment 1 (theoretical part)}
+\newcommand{\getsubtitle}{G2T1 – Assignment 2 (theoretical part)}
 \newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
@@
 
-\section*{1 Transparency Levels}
+\section*{1 Parameter Passing and RMI}
 \subsection*{a)}
-The access is location transparent, because the location is hidden behind an unresolved URL.
-The actual adress of the webservice is automatically resolved through DNS acting as service discovery mechanism.
-\subsection*{b)}
-The service is not replication transparent because there is a unique name for every location that has to be checked separatly.
-\subsection*{c)}
-The access is replication transparent, because it is not apparent whether Otto checks one or multiple web services.
-It's furthermore unknown which of the two replication locations did provide the file in the end.
-\subsection*{d)}
-\subsubsection*{i.}
-It is only possible to write all copies at once so all copies are always the same.
-This iplies that independently from when or which copy is read the answer is always up to date.
-\subsubsection*{ii.}
-Write operations are performed on \(n-1\) copies, as required by the locking constraints.
-Read operations are performed on \(2\) different randomly selected copies, the client accepts the "latest available" version.
+\begin{enumerate}[label=(\roman*)]
+	\item 
+	\item 
+	\item 
+\end{enumerate}
 
-\section*{2 System Models}
-\subsection*{a)}
-\begin{itemize}
-	\item No message is lost
-	\item The maximum possible delay is known
-	\item There is a limited amount of clock skew between the processes
-\end{itemize}
-\subsection*{b)}
-\subsubsection*{i.}
-Assuming a limited clock skew of \(\delta c\) between the processes, the process \(P_i\) sends a broadcast message \(2 * \delta c + \delta t\) time units after receiving \(i - 1\) messages.
-This allows for instant delivery to a process skewed by the maximum amount of time ''forwards'' while keeping the order of messages intact when sending the message even with the maximum delay \(\delta t\) to a process skewed by the maximum amount of time ''backwards''.
-\subsubsection*{ii.}
-In the worst case, the synchronization protocol takes \(2n (\delta t + \delta c) + 2\delta c\) time units.
-This assumes that all messages take the maximum possible amount of time to deliver, namely \(n delta t\) overall, with clock skew cancelling itself out across all processes.
-Adding the wait-time for each process \(n(\delta t + 2\delta c)\) yields \(2n (\delta t + \delta c)\).
-The final component is assuming the maximum clock skew between \(P_1 \text{ and } P_n\).
-
-\section*{3 Three-Army-Problem}
-\subsection*{a)}
 \subsection*{b)}
 
-\section*{4 System Availability}
+\subsection*{c)}
 
-\begin{enumerate}[a)]
-	\item $ A_x = \dfrac{80t}{100t} = 80\% $
+\section*{2 Chord System}
+\subsection*{a)}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{4}
+\end{table}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{11}
+\end{table}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{16}
+\end{table}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{19}
+\end{table}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{23}
+\end{table}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{26}
+\end{table}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|}
+		\hline
+		1 &  \\ \hline
+		2 &  \\ \hline
+		3 &  \\ \hline
+		4 &  \\ \hline
+		5 &  \\ \hline
+	\end{tabular}
+	\caption{31}
+\end{table}
+\subsection*{b)}
+\subsection*{c)}
+\subsection*{d)}
+\subsection*{e)}
+\subsection*{f)}
 
-		$ A_y = \dfrac{60t}{100t} = 60\%$
+\section*{3 Name Services}
+\subsection*{a)}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|l|}
+		\hline
+		& Iterative resolution & Recursive resolution \\ \hline
+		1. Lookup &  &  \\ \hline
+		2. Lookup &  &  \\ \hline
+		3. Lookup &  &  \\ \hline
+	\end{tabular}
+\end{table}
+\subsection*{b)}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|l|l|l|}
+		\hline
+		& Iterative resolution & Recursive resolution \\ \hline
+		1. Lookup &  &  \\ \hline
+		2. Lookup &  &  \\ \hline
+		3. Lookup &  &  \\ \hline
+	\end{tabular}
+\end{table}
+\subsection*{c)}
 
-	\item Overall statistical availability:
-
-		$ A_S = 1 - P(Y=\text{down} \cup X=\text{down}) = 1 - ((1 - A_x) \cdot (1 - A_y)) = 1 - (0.2 \cdot 0.4) = 1 - 0.08 = 0.92 = 92\%$
-
-		Availability in the concrete example:
-
-		$ A_S = \dfrac{80t}{100t} = 80\% $
-
-	\item $ P(Y=\text{up}|X=\text{up}) = \frac{P(Y=\text{up} \cap X=\text{up})}{P(X=\text{up})} = \frac{0.6}{0.8} = 75\% $
-
-		$ P(Y=\text{up}|X=\text{down}) = \frac{P(Y=\text{up} \cap X=\text{down})}{P(X=\text{down})} = \frac{0}{0.2} = 0\% $
-
-		Node $Y$ seems to depend on Node $X$ to be available.
-
-	\item The availability of the system ($ A_S $) depends on the node $X$.
-
-		$ A_S = P(Y=\text{up} \cup X=\text{up}) = P(X=\text{up}) = A_X = 0.8 $
-
-		This is different from the result in b) ($ A_S = 0.92 $) where the assumption was that the nodes have independent failures.
-		If the nodes are dependent then the overall system can not achieve a higher availability than any one of the nodes.
-
+\begin{enumerate}[label=(\roman*)]
+	\item 
+	\item 
+	\item 
 \end{enumerate}

### 2019-11-11 assignment2/submission2.tex
done 2.3 theoretical and started 2.2

@@
 
-\newcommand{\gettitle}{Verteilte Systeme I\\Winter Term 2019/20}
+\newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
 \newcommand{\getsubtitle}{G2T1 – Assignment 2 (theoretical part)}
@@
 
-\section*{1 Parameter Passing and RMI}
+\section*{1 - Parameter Passing and RMI}
 \subsection*{a)}
@@
 
-\section*{2 Chord System}
+\section*{2 - Chord System}
 \subsection*{a)}
+
+\begin{table}[!htb]
+	\begin{minipage}[t]{.33\linewidth}
+		\centering
+			\begin{tabular}{|l|l|}
+				\hline
+				1 & 11 \\ \hline
+				2 & 11 \\ \hline
+				3 & 11 \\ \hline
+				4 & 16 \\ \hline
+				5 & 23 \\ \hline
+			\end{tabular}
+		\caption{4}
+	\end{minipage}%
+	\hfill%
+	\begin{minipage}[t]{.33\linewidth}
+		\centering
+		\begin{tabular}{|l|l|}
+			\hline
+			1 & 16 \\ \hline
+			2 & 16 \\ \hline
+			3 & 16 \\ \hline
+			4 & 19 \\ \hline
+			5 & 31 \\ \hline
+		\end{tabular}
+		\caption{11}
+	\end{minipage}%
+	\hfill%
+	\begin{minipage}[t]{.33\linewidth}
+		\centering
+		\begin{tabular}{|l|l|}
+			\hline
+			1 & 4 \\ \hline
+			2 & 26 \\ \hline
+			3 & 23 \\ \hline
+			4 & 19 \\ \hline
+			5 & 19 \\ \hline
+		\end{tabular}
+		\caption{16}
+\end{minipage}
+\end{table}
+
+\begin{table}[!htb]
+	\begin{minipage}[t]{.33\linewidth}
+		\centering
+		\begin{tabular}{|l|l|}
+			\hline
+			1 & 23 \\ \hline
+			2 & 23 \\ \hline
+			3 & 23 \\ \hline
+			4 & 31 \\ \hline
+			5 & 4 \\ \hline
+		\end{tabular}
+		\caption{19}
+	\end{minipage}%
+	\hfill%
+	\begin{minipage}[t]{.33\linewidth}
+		\centering
+		\begin{tabular}{|l|l|}
+			\hline
+			1 & 26 \\ \hline
+			2 & 26 \\ \hline
+			3 & 31 \\ \hline
+			4 & 31 \\ \hline
+			5 & 11 \\ \hline
+		\end{tabular}
+		\caption{23}
+	\end{minipage}%
+	\hfill%
+	\begin{minipage}[t]{.33\linewidth}
+		\centering
+		\begin{tabular}{|l|l|}
+			\hline
+			1 & 31 \\ \hline
+			2 & 31 \\ \hline
+			3 & 31 \\ \hline
+			4 & 4 \\ \hline
+			5 & 11 \\ \hline
+		\end{tabular}
+		\caption{26}
+	\end{minipage}
+\end{table}
+
 \begin{table}[!ht]
@@
 		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
+		1 & 4 \\ \hline
+		2 & 4 \\ \hline
+		3 & 4 \\ \hline
+		4 & 11 \\ \hline
+		5 & 16 \\ \hline
 	\end{tabular}
-	\caption{4}
+	\caption{31}
 \end{table}
+\subsection*{b)}
+\begin{enumerate}
+	\item node 4 $ \rightarrow $ node 16
+	\item node 16 $ \rightarrow $ node 19
+	\item node 19 $ \rightarrow $ node 23
+\end{enumerate}
+\subsection*{c)}
 \begin{table}[!ht]
@@
 		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
+		1 & 26 \\ \hline
+		2 & 26 \\ \hline
+		3 & 31 \\ \hline
+		4 & 4 \\ \hline
+		5 & 11 \\ \hline
 	\end{tabular}
-	\caption{11}
+	\caption{24}
 \end{table}
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{|l|l|}
-		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
-	\end{tabular}
-	\caption{16}
-\end{table}
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{|l|l|}
-		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
-	\end{tabular}
-	\caption{19}
-\end{table}
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{|l|l|}
-		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
-	\end{tabular}
-	\caption{23}
-\end{table}
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{|l|l|}
-		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
-	\end{tabular}
-	\caption{26}
-\end{table}
-\begin{table}[!ht]
-	\centering
-	\begin{tabular}{|l|l|}
-		\hline
-		1 &  \\ \hline
-		2 &  \\ \hline
-		3 &  \\ \hline
-		4 &  \\ \hline
-		5 &  \\ \hline
-	\end{tabular}
-	\caption{31}
-\end{table}
-\subsection*{b)}
-\subsection*{c)}
+The finger tables of Node 11, 16, 19 and 23 should be recomputed.\\
+The changes only occur in Node 16 and 23.
 \subsection*{d)}
+Looking at the node 23, there has to be an update on the fifth entry on the finger table.
+When Node 7 and 9 are joining at the exact same time the fifth entry on the finger table of node 23 can become invalid.
+Same applies for Node 31 in the fourth entry.
 \subsection*{e)}
@@
 
-\section*{3 Name Services}
+\section*{3 - Name Services}
 \subsection*{a)}
@@
 	\centering
-	\begin{tabular}{|l|l|l|}
+	\begin{tabular}{|c|c|c|}
 		\hline
 		& Iterative resolution & Recursive resolution \\ \hline
-		1. Lookup &  &  \\ \hline
-		2. Lookup &  &  \\ \hline
-		3. Lookup &  &  \\ \hline
+		1. Lookup & 5 & 5 \\ \hline
+		2. Lookup & 5 & 5 \\ \hline
+		3. Lookup & 5 & 5 \\ \hline
 	\end{tabular}
@@
 	\centering
-	\begin{tabular}{|l|l|l|}
+	\begin{tabular}{|c|c|c|}
 		\hline
 		& Iterative resolution & Recursive resolution \\ \hline
-		1. Lookup &  &  \\ \hline
-		2. Lookup &  &  \\ \hline
-		3. Lookup &  &  \\ \hline
+		1. Lookup & 5 & 5 \\ \hline
+		2. Lookup & 5 & 4 \\ \hline
+		3. Lookup & 5 & 5 \\ \hline
 	\end{tabular}
@@
 
-\begin{enumerate}[label=(\roman*)]
-	\item 
-	\item 
-	\item 
+\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
+\subsubsection*{i.}
+\begin{lstlisting}
+$ dig NS uni-stuttgart.de
+...
+;; ANSWER SECTION:
+uni-stuttgart.de.	2191	IN	NS	minnehaha.rhrk.uni-kl.de.
+uni-stuttgart.de.	2191	IN	NS	dns3.belwue.de.
+uni-stuttgart.de.	2191	IN	NS	dns1.belwue.de.
+uni-stuttgart.de.	2191	IN	NS	dns0.uni-stuttgart.de.
+uni-stuttgart.de.	2191	IN	NS	dns1.uni-stuttgart.edu.
+...
+\end{lstlisting}
+There are 5 server responsible for the DNS entry of 'uni-stuttgart.de':
+\begin{itemize}
+	\item minnehaha.rhrk.uni-kl.de
+	\item dns3.belwue.de
+	\item dns1.belwue.de
+	\item dns0.uni-stuttgart.de
+	\item dns1.uni-stuttgart.edu
+\end{itemize}
+\subsubsection*{ii.}
+\begin{lstlisting}
+$ dig -x 129.69.216.249
+...
+;; ANSWER SECTION:
+249.216.69.129.in-addr.arpa. 86397 IN	PTR	ipvslogin.informatik.uni-stuttgart.de.
+...
+\end{lstlisting}
+PTR is usefull for resolving an IP address to a domain or hostname (reverse DNS lookup).\\
+The given IP resolves to: 'ipvslogin.informatik.uni-stuttgart.de'
+\subsubsection*{iii.}
+\begin{lstlisting}
+$ dig +trace www.uni-stuttgart.de
+...
+.			347153	IN	NS	h.root-servers.net.
+.			347153	IN	NS	l.root-servers.net.
+.			347153	IN	NS	d.root-servers.net.
+.			347153	IN	NS	m.root-servers.net.
+.			347153	IN	NS	e.root-servers.net.
+.			347153	IN	NS	k.root-servers.net.
+.			347153	IN	NS	i.root-servers.net.
+.			347153	IN	NS	b.root-servers.net.
+.			347153	IN	NS	c.root-servers.net.
+.			347153	IN	NS	a.root-servers.net.
+.			347153	IN	NS	f.root-servers.net.
+.			347153	IN	NS	g.root-servers.net.
+.			347153	IN	NS	j.root-servers.net.
+...
+;; Received 553 bytes from 192.168.0.1#53(192.168.0.1) in 28 ms
+
+de.			172800	IN	NS	z.nic.de.
+de.			172800	IN	NS	f.nic.de.
+de.			172800	IN	NS	s.de.net.
+de.			172800	IN	NS	a.nic.de.
+de.			172800	IN	NS	l.de.net.
+de.			172800	IN	NS	n.de.net.
+...
+;; Received 802 bytes from 2001:500:12::d0d#53(g.root-servers.net) in 22 ms
+
+uni-stuttgart.de.	86400	IN	NS	dns0.uni-stuttgart.de.
+uni-stuttgart.de.	86400	IN	NS	dns1.belwue.de.
+uni-stuttgart.de.	86400	IN	NS	dns1.uni-stuttgart.edu.
+uni-stuttgart.de.	86400	IN	NS	dns3.belwue.de.
+uni-stuttgart.de.	86400	IN	NS	minnehaha.rhrk.uni-kl.de.
+...
+;; Received 619 bytes from 2001:67c:1011:1::53#53(n.de.net) in 25 ms
+
+www.uni-stuttgart.de.	3600	IN	A	129.69.8.19
+...
+uni-stuttgart.de.	3600	IN	NS	dns0.uni-stuttgart.de.
+uni-stuttgart.de.	3600	IN	NS	minnehaha.rhrk.uni-kl.de.
+uni-stuttgart.de.	3600	IN	NS	dns1.uni-stuttgart.edu.
+uni-stuttgart.de.	3600	IN	NS	dns3.belwue.de.
+uni-stuttgart.de.	3600	IN	NS	dns1.belwue.de.
+...
+;; Received 1365 bytes from 129.143.2.10#53(dns1.belwue.de) in 25 ms
+
+\end{lstlisting}
+The path the lookup took was:
+\begin{enumerate}
+	\item 192.168.0.1
+	\item g.root-servers.net
+	\item n.de.net
+	\item dns1.belwue.de
 \end{enumerate}

### 2019-11-11 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/client/CalcSocketClient.java
Completely overkill practical exercise assignment 2.5 (ex01)

@@ -1,5 +1,7 @@
 package de.unistgt.ipvs.vs.ex1.client;
 
+import de.unistgt.ipvs.vs.ex1.common.MessageReader;
+
 import java.io.*;
 import java.net.Socket;
 
@@ -13,8 +15,8 @@ public class CalcSocketClient {
 	private int    rcvdErs;		// --> Number of invalid message contents
 	private int    calcRes;		// --> Calculation result (cf.  'RES')
 
-	ObjectOutputStream out;
-	ObjectInputStream in;
+	private OutputStream out;
+	private MessageReader in;
 	
 	public CalcSocketClient() {
 		this.cliSocket = null;
@@ -40,11 +42,11 @@ public class CalcSocketClient {
 		//Solution here
 		try {
 			cliSocket = new Socket(srvIP, srvPort);
-			out = new ObjectOutputStream(cliSocket.getOutputStream());
-			in = new ObjectInputStream(cliSocket.getInputStream());
+			out = cliSocket.getOutputStream();
+			in = new MessageReader(cliSocket.getInputStream());
 
 			// check if RDY was sent, else exit
-			String rdyMsg = (String) in.readObject();
+			String rdyMsg = in.next();
 			if(!rdyMsg.equals("<08:RDY>")) {
 				System.err.println("The server did not send the RDY message.");
 				return false;
@@ -72,8 +74,37 @@ public class CalcSocketClient {
 			System.err.println("Client not connected!");
 			return false;
 		}
-		//Solution here
-		// TODO
+		try {
+			// send the request to the server
+			out.write(request.getBytes());
+			// get the responses and handle them
+			outer: while (true) {
+				String response = in.next();
+				String[] contents = in.contents(response.toUpperCase());
+				for (int i = 0; i < contents.length; i++) {
+					String responsePart = contents[i];
+					switch (responsePart) {
+						case "OK":
+							rcvdOKs++;
+							break;
+						case "ERR":
+							rcvdErs++;
+							break;
+						case "RES":
+							// automatically advance loop counter to skip result for the loop
+							calcRes = Integer.parseInt(contents[++i]);
+							break;
+						case "FIN":
+							//message has been handled, return control flow to caller
+							break outer;
+					}
+				}
+			}
+		} catch (IOException e) {
+			System.err.println("Failed to calculate request with an exception");
+			e.printStackTrace();
+			return false;
+		}
 		return true;
 	}
 }

### 2019-11-11 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/common/MessageReader.java

@@ -1,0 +1,110 @@
+package de.unistgt.ipvs.vs.ex1.common;
+
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.InputStreamReader;
+import java.util.Iterator;
+import java.util.regex.Matcher;
+import java.util.regex.Pattern;
+
+/**
+ * Utility class for handling the reading of messages according to the spec outlined in practical assignment 1.
+ * This is accomplished by wrapping an InputStream and reading it into an internal buffer to simplify parsing.
+ */
+public final class MessageReader implements Iterator<String>, AutoCloseable {
+
+    private static final Pattern messageForm = Pattern.compile("<(?<length>\\d{2}):(?<body>[^>]+?)>");
+
+    private final StringBuilder buffer = new StringBuilder();
+
+    private final InputStreamReader wrappedStream;
+    private boolean eofEncountered = false;
+
+    public MessageReader(InputStream wrappedStream) {
+        this.wrappedStream = new InputStreamReader(wrappedStream);
+    }
+
+    /**
+     * @return true, so long as no EOF has been read
+     */
+    @Override
+    public boolean hasNext() {
+        return !eofEncountered;
+    }
+
+    /**
+     * Obtains the next valid message from the wrapped input stream, blocking to wait for IO if necessary.
+     * @return the next syntactically valid message or an empty string
+     *      if the input has been exhausted or an error occurred during reading.
+     */
+    @Override
+    public String next() {
+        Matcher bufferedMatch = messageForm.matcher(buffer);
+        if (bufferedMatch.find()) {
+            String result = bufferedMatch.group();
+            buffer.delete(0, bufferedMatch.end());
+            return result;
+        }
+        // read input into buffer until we find something cool
+        // because the buffer is larger than 100 characters, any message conforming to the spec
+        // should be readable in a single read operationo
+        char[] readBuffer = new char[128];
+        while (!messageForm.matcher(buffer).find()) {
+            final int count;
+            try {
+                // read method is blocking and waits for input if necessary
+                count = wrappedStream.read(readBuffer);
+            } catch (IOException e) {
+                System.err.println("Could not read input for message");
+                e.printStackTrace();
+                break;
+            }
+            if (count == -1) {
+                eofEncountered = true;
+                break;
+            }
+            buffer.append(readBuffer, 0, count);
+        }
+        bufferedMatch = messageForm.matcher(buffer);
+        if (bufferedMatch.find()) {
[…]

### 2019-11-11 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/common/MessageWriter.java

@@ -1,0 +1,76 @@
+package de.unistgt.ipvs.vs.ex1.common;
+
+
+import java.io.IOException;
+import java.io.ObjectOutputStream;
+import java.io.OutputStream;
+
+/**
+ * Adds support for batching message capabilities while automatically handling message length requirements
+ * of the protocol outlined in Practical Assignment 1.
+ */
+public class MessageWriter implements AutoCloseable {
+    // message contents may be 94 long before overflowing the size field
+    // this comes from the fact that 5 chars are necessary for the protocol overhead
+    // <dd:[..]>
+    private static final int SAFE_LENGTH = 94;
+    // %02d -> zero-padded two digit number
+    private static final String MESSAGE_FORMAT = "<%02d:%s>";
+    private static final String PART_SEPARATOR = " ";
+    // storage for the message parts
+    private final StringBuilder buffer = new StringBuilder(100);
+    private final OutputStream writeTarget;
+
+    public MessageWriter(OutputStream writeTarget) {
+        this.writeTarget = writeTarget;
+    }
+
+    /**
+     * Forcibly sends the given part to the encapsulated output stream, flushing the buffer in the process.
+     * @param part The message part to be sent. This is not checked for validity.
+     */
+    public void forceSend(String part) throws IOException {
+        safeAppend(part);
+        send();
+    }
+
+    /**
+     * Adds a part to the message buffer to be sent at a later date.
+     * If the resulting buffer would be too large to send in a single message, the current buffer is written before sending.
+     * @param part The message part to be sent. This is not checked for validity.
+     */
+    public void enqueue(String part) throws IOException {
+        safeAppend(part);
+    }
+
+    /**
+     * Handle message size requirements and part delimiter when adding to the message buffer
+     */
+    private void safeAppend(String part) throws IOException {
+        // adding message to the buffer would overflow message size
+        if (buffer.length() + part.length() + PART_SEPARATOR.length() >= SAFE_LENGTH) {
+            // therefore flush the buffer to output
+            send();
+        }
+        // if buffer isn't empty, separate the parts
+        if (buffer.length() != 0) {
+            buffer.append(PART_SEPARATOR);
+        }
+        buffer.append(part);
+    }
+
+    /**
+     * Writes the current buffer and subsequently clears it.
+     */
+    private void send() throws IOException {
+        writeTarget.write(String.format(MESSAGE_FORMAT, buffer.length() + 5, buffer.toString()).getBytes());
[…]

### 2019-11-11 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/server/CalcSocketServer.java

@@ -13,6 +13,7 @@ public class CalcSocketServer extends Thread {
 	private int port;
 
 	public CalcSocketServer(int port) {
+		super("Calculation Server Accept Thread");
 		this.srvSocket = null;
 		this.port = port;
 	}
@@ -37,6 +38,7 @@ public class CalcSocketServer extends Thread {
 		try {
 			srvSocket = new ServerSocket(port);
 			while(true) {
+				// spawn a new session for each client that connects
 				new Thread(new CalculationSession(srvSocket.accept())).start();
 			}
 		} catch (IOException e) {
@@ -44,7 +46,7 @@ public class CalcSocketServer extends Thread {
 		}
 	}
 
-	public void waitUnitlRunnig(){
+	public void waitUntilRunning(){
 		while(this.srvSocket == null){
 			try {
 				Thread.sleep(1);

### 2019-11-11 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/server/CalculationSession.java

@@ -1,17 +1,19 @@
 package de.unistgt.ipvs.vs.ex1.server;
 
 import de.unistgt.ipvs.vs.ex1.common.ICalculation;
+import de.unistgt.ipvs.vs.ex1.common.MessageReader;
+import de.unistgt.ipvs.vs.ex1.common.MessageWriter;
 
-import java.io.ObjectInputStream;
-import java.io.ObjectOutputStream;
+import java.io.IOException;
 import java.net.Socket;
+import java.rmi.RemoteException;
 
 /**
  * Add fields and methods to this class as necessary to fulfill the assignment.
  */
 public class CalculationSession implements Runnable {
 	Socket cliSocket;
-
+	private Operation currentOp;
 	enum Operation {
 		Add, Subtract, Multiply
 	}
@@ -22,18 +24,82 @@ public class CalculationSession implements Runnable {
 
 	@Override
 	public void run() {
-		// System.out.println("New session thread");
-		try {
-			ObjectOutputStream out = new ObjectOutputStream(this.cliSocket.getOutputStream());
-			ObjectInputStream in = new ObjectInputStream(this.cliSocket.getInputStream());
-
+		try (final MessageReader in = new MessageReader(this.cliSocket.getInputStream());
+			final MessageWriter out = new MessageWriter(this.cliSocket.getOutputStream())) {
 			// Send RDY when ready
-			out.writeObject("<08:RDY>");
-
-			ICalculation calc = new CalculationImpl();
-			// TOOD
+			out.forceSend("RDY");
+			final ICalculation calc = new CalculationImpl();
+			while (in.hasNext()) {
+				// in.next is a blocking call waiting for a complete message if necessary
+				String message = in.next();
+				if (message.isEmpty()) {
+					// empty message can only happen if the input is exhausted
+					// therefore skip the whole message handling
+					break;
+				}
+				// acknowledge the message
+				out.forceSend("OK");
+				for (String part : in.contents(message.toUpperCase())) {
+					processPart(part, calc, out);
+				}
+				// notify the client of the finished processing for the current message
+				out.forceSend("FIN");
+			}
 		} catch (Exception e) {
 			e.printStackTrace();
+		}
+	}
+
+	private void processPart(final String part, final ICalculation calc, final MessageWriter out) throws IOException {
+		switch (part) {
+			case "ADD":
+				currentOp = Operation.Add;
+				out.enqueue("OK " + part);
+				break;
+			case "SUB":
[…]

### 2019-11-11 assignment2/handout/src/test/java/de/unistgt/ipvs/vs/ex1/TestCalc.java

@@ -2,7 +2,7 @@ package de.unistgt.ipvs.vs.ex1;
 
 import de.unistgt.ipvs.vs.ex1.client.CalcSocketClient;
 import de.unistgt.ipvs.vs.ex1.server.CalcSocketServer;
-import java.io.IOException;
+
 import static org.junit.Assert.*;
 import org.junit.Rule;
 
@@ -24,7 +24,7 @@ public class TestCalc {
         // Start Server
         CalcSocketServer cSrv = new CalcSocketServer(srvPort);
         cSrv.start();
-        cSrv.waitUnitlRunnig();
+        cSrv.waitUntilRunning();
 
         CalcSocketClient cCli = new CalcSocketClient();
         cCli.connectTo(srvIP, srvPort);
@@ -49,7 +49,7 @@ public class TestCalc {
         // Start Server
         CalcSocketServer cSrv = new CalcSocketServer(srvPort);
         cSrv.start();
-        cSrv.waitUnitlRunnig();
+        cSrv.waitUntilRunning();
 
         CalcSocketClient cCli = new CalcSocketClient();
         cCli.connectTo(srvIP, srvPort);
@@ -73,7 +73,7 @@ public class TestCalc {
         // Start Server
         CalcSocketServer cSrv = new CalcSocketServer(srvPort);
         cSrv.start();
-        cSrv.waitUnitlRunnig();
+        cSrv.waitUntilRunning();
 
         CalcSocketClient cCli = new CalcSocketClient();
         cCli.connectTo(srvIP, srvPort);
@@ -91,6 +91,5 @@ public class TestCalc {
         
         cCli.disconnect();
     }
-        
 
 }

### 2019-11-14 assignment2/submission2.tex
Fix errors and formatting in submission2

@@
 
-\begin{table}[!htb]
-	\begin{minipage}[t]{.33\linewidth}
+	\begin{minipage}[t]{.25\linewidth}
         \centering
+        Node 4:\\
 			\begin{tabular}{|l|l|}
@@
 			\end{tabular}
-		\caption{4}
 	\end{minipage}%
-	\hfill%
-	\begin{minipage}[t]{.33\linewidth}
+	\begin{minipage}[t]{.25\linewidth}
         \centering
+        Node 11:\\
 		\begin{tabular}{|l|l|}
@@
 		\end{tabular}
-		\caption{11}
 	\end{minipage}%
-	\hfill%
-	\begin{minipage}[t]{.33\linewidth}
+	\begin{minipage}[t]{.25\linewidth}
         \centering
+        Node 16:\\
 		\begin{tabular}{|l|l|}
 			\hline
-			1 & 4 \\ \hline
-			2 & 26 \\ \hline
+			1 & 19 \\ \hline
+			2 & 19 \\ \hline
 			3 & 23 \\ \hline
-			4 & 19 \\ \hline
-			5 & 19 \\ \hline
+			4 & 26 \\ \hline
+			5 & 4 \\ \hline
 		\end{tabular}
-		\caption{16}
     \end{minipage}
-\end{table}
-
-\begin{table}[!htb]
-	\begin{minipage}[t]{.33\linewidth}
+    \begin{minipage}[t]{.25\linewidth}
         \centering
+        Node 19:\\
         \begin{tabular}{|l|l|}
@@
         \end{tabular}
-		\caption{19}
     \end{minipage}%
-	\hfill%
-	\begin{minipage}[t]{.33\linewidth}
+
+	\begin{minipage}[t]{.25\linewidth}
 		\centering
+        Node 23:\\
 		\begin{tabular}{|l|l|}
@@
 		\end{tabular}
-		\caption{23}
 	\end{minipage}%
-	\hfill%
-	\begin{minipage}[t]{.33\linewidth}
+	\begin{minipage}[t]{.25\linewidth}
 		\centering
+        Node 26:\\
 		\begin{tabular}{|l|l|}
@@
 		\end{tabular}
-		\caption{26}
 	\end{minipage}
-\end{table}
-
-\begin{table}[!ht]
+	\begin{minipage}[t]{.25\linewidth}
 		\centering
+        Node 31:\\
 		\begin{tabular}{|l|l|}
@@
 		\end{tabular}
-	\caption{31}
-\end{table}
+	\end{minipage}
+
 \subsection*{b)}
@@
 \end{enumerate}
+
 \subsection*{c)}
-\begin{table}[!ht]
+\begin{minipage}[t]{.25\linewidth}
 	\centering
+    Node 24:\\
 	\begin{tabular}{|l|l|}
@@
 	\end{tabular}
-	\caption{24}
-\end{table}
-The finger tables of Node 11, 16, 19 and 23 should be recomputed.\\
-The changes only occur in Node 16 and 23.
+\end{minipage}
+
+The finger table of node 16 will change eventually while the one for node 23 will change immediately.
+
 \subsection*{d)}
-Looking at the node 23, there has to be an update on the fifth entry on the finger table.
-When Node 7 and 9 are joining at the exact same time the fifth entry on the finger table of node 23 can become invalid.
-Same applies for Node 31 in the fourth entry.
+If both nodes join simultaneously they are between the same two nodes (4 and 11).
+If they now try to each update their predecessor and successor then we have a race condition while updating a linked list.
+This can leave the list (or in this case ring) in an inconsistent state that makes routing impossible!
+
 \subsection*{e)}
@@
 \section*{3 - Name Services}
+
 \subsection*{a)}
-\begin{table}[!ht]
-	\centering
+\begin{center}
 	\begin{tabular}{|c|c|c|}
@@
 		& Iterative resolution & Recursive resolution \\ \hline
-		1. Lookup & 5 & 5 \\ \hline
-		2. Lookup & 5 & 5 \\ \hline
-		3. Lookup & 5 & 5 \\ \hline
+		1. Lookup & 10 & 10 \\ \hline
+		2. Lookup & 10 & 10 \\ \hline
+		3. Lookup & 8 & 8 \\ \hline
 	\end{tabular}
-\end{table}
+\end{center}
+
 \subsection*{b)}
-\begin{table}[!ht]
-	\centering
+\begin{center}
 	\begin{tabular}{|c|c|c|}
@@
 		& Iterative resolution & Recursive resolution \\ \hline
-		1. Lookup & 5 & 5 \\ \hline
-		2. Lookup & 5 & 4 \\ \hline
-		3. Lookup & 5 & 5 \\ \hline
+		1. Lookup & 10 & 10 (8 fast, 2 slow) \\ \hline
+		2. Lookup & 10 & 8 (6 fast, 2 slow) \\ \hline
+		3. Lookup & 8 & 8 (6 fast, 2 slow) \\ \hline
 	\end{tabular}
-\end{table}
+\end{center}
+
 \subsection*{c)}

### 2019-11-15 assignment2/submission2.tex
Exercise 1 solution added

@@
 \begin{enumerate}[label=(\roman*)]
-	\item 
-	\item 
-	\item 
+	\item [1, 2, 3, 4, 5]
+	\item [5, 4, 2, 4, 5]
+	\item [1, 2, 3, 4, 5] or [5, 4, 3, 2, 1] depending on the order in which the arrays are copied back.
 \end{enumerate}
@@
 \subsection*{b)}
-
+All objects that extend UnicastRemoteObject are send as reference, all other objects are send as value.
+In the frist and last two lines the parameter is a StringBuffer object wich does not implement UnicastRemoteObject and send as value.
+In the third line the parameter is a StringServer object wich does implement UnicastRemoteObject and because of this is send as reference.
 \subsection*{c)}
+Call-by-Value has a huge overhead if a large structure is given of wich only a small part is writen/read.
+Call-by-reference has a huge overhead if it is often writen/read.
 

### 2019-11-15 assignment2/submission2.tex
Expand and clarify solution to Problem 2.1

--- 2019-11-14 assignment2/submission2.tex
@@
 \begin{enumerate}[label=(\roman*)]
-	\item 
-	\item 
-	\item 
+	% no return value, so a is never changed
+	\item \texttt{[1, 2, 3, 4, 5]}
+	% the code given does not perform a swap, it only overwrites the lower indices with values from the upper ones.
+	% because of reference semantics that overwrites the second input as well
+	\item \texttt{[5, 4, 3, 4, 5]}
+	\item \texttt{[5, 4, 3, 2, 1]} \emph{or} \texttt{[1, 2, 3, 4, 5]} \emph{or} \texttt{[5, 4, 3, 4, 5]}, depending on the exact semantic behaviour of the copy and restore operations.
+	Given that the remote server passes the arguments by reference, the first question is whether \textit{x} and \textit{y} have separate copies of \textit{a}. 
+	If they do, the first or second option is possible. 
+	If there are no separate copies of \textit{a}, the procedure behaves as if the parameters were passed by reference. 
+	Otherwise it depends on which value is restored first.
 \end{enumerate}
@@
 
+The result printed to the system console is:
+\begin{verbatim}
+$ java StringServer
+1: B
+2: BRS
+3: ABSBRS
+\end{verbatim}
+
+We can deduce that all objects that extend UnicastRemoteObject are sent as reference, all other objects are sent as value.
+In the first and last two lines the parameter is a StringBuffer object which does not implement UnicastRemoteObject and is therefore sent as value. 
+In the third line the parameter is a StringServer object which does implement UnicastRemoteObject and because of this is sent as reference.
+
 \subsection*{c)}
+
+The first line shows that even "complex java objects" are passed by value so long as they are not remote objects. 
+This implies that even very large non-remote objects would be serialized and passed by value (e.g. Arrays with hundreds of thousands of elements).
+Such a behaviour incurs large overheads, especially if only parts of the passed value are accessed by the remote procedure.
+
+The lines two and three demonstrate the reference-passing behaviour within java's RMI implementation. 
+Notably, remote objects are passed by reference. 
+Calls to the passed remote objects are routed back to the JVM the actual object resides in. 
+This can lead to large overheads if a lot of calls are made to the referenced object.
 

### 2019-11-15 assignment2/check/RemoteBuffer.java

@@ -1,0 +1,5 @@
+public interface RemoteBuffer extends java.rmi.Remote {
+  public StringBuffer buffer() throws java.rmi.RemoteException;
+  public void append_R(RemoteBuffer r) throws java.rmi.RemoteException;
+  public void append_S(StringBuffer s) throws java.rmi.RemoteException;
+}

### 2019-11-15 assignment2/check/StringServer.java

@@ -1,0 +1,39 @@
+public class StringServer extends java.rmi.server.UnicastRemoteObject implements RemoteBuffer {
+  private StringBuffer buffer;
+  public StringServer(String s) throws java.rmi.RemoteException {
+    super();
+    buffer = new StringBuffer(s);
+  }
+  @Override
+  public StringBuffer buffer() {
+    return buffer;
+  }
+
+  @Override
+  public void append_R(RemoteBuffer r) throws java.rmi.RemoteException {
+    r.append_S(new StringBuffer("R"));
+    buffer.append(r.buffer());
+  }
+
+  @Override
+  public void append_S(StringBuffer s) throws java.rmi.RemoteException {
+    s.append("S");
+    buffer.append(s);
+  }
+
+  public static void main(String[] args) throws Exception {
+    StringServer s1 = new StringServer("A");
+    StringServer s2 = new StringServer("B");
+    java.rmi.Naming.bind("rmi://localhost/MyBuffer", s1);
+    RemoteBuffer rb = (RemoteBuffer) java.rmi.Naming.lookup("rmi://localhost/MyBuffer");
+    
+    rb.append_S(s2.buffer());
+    System.out.println("1: " + s2.buffer());
+    rb.append_R(s2);
+    System.out.println("2: " + s2.buffer());
+    System.out.println("3: " + rb.buffer());
+    java.rmi.Naming.unbind("rmi://localhost/MyBuffer");
+  }
+}
+
+

### 2019-11-15 assignment2/submission2.tex
Add algorithm for solving 2.2e

@@
 \usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 \usepackage{parskip}
@@
 \subsection*{e)}
+
+This code updates the FT entries of \textit{p} for the former successor's predecessor which has the id \textit{q}.
+
+\begin{algorithm}
+\begin{algorithmic}
+	\For{$i \gets 0 \textbf{ to } \lfloor log_2(q - p) \rfloor$}
+	   \State $FT_p[i] \gets q$
+	\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+This code assumes that only \textit{q} has been inserted between \textit{p} and the former successor.
+The obvious issue arising from that (namely that multiple insertions are not correctly handled), can be easily remedied by repeating the consistency check after updating the finger table.
+As presented the code is already eventually consistent, though.
+
 \subsection*{f)}

### 2019-11-15 assignment2/submission2.tex
Add argumentation based on forum entries for 2.2f

@@
 
+As outlined in the forum, it's assumed that the node ids within the system are at least vaguely evenly distributed under the given hash function.
+Given that assumption and the assumption that there are at least three nodes in the system, it's reasonable to assume that no node is responsible for an adress space larger than \(2^m-1\).
+Carrying that assumption onwards, any entity will be stored on two nodes, because the smallest \(p \leq ID_1(e)\) will resolve to different nodes in the separate chord rings.
+
+This property holds true under the assumption that no node is responsible for an adress space larger than the largest lookup shortcut, which is the offset for the second chord ring.
+If that property is not given, the insert is not guaranteed to store an entity's copy on two different nodes.
+That failure can be seen when considering \(ID_1(n_1) + 2 = ID_1(n_2) + 1 = ID_1(n_3)\) for three nodes \(\{n_1, n_2, n_3\}\).
+It's likely that an inserted entity on a chord-ring structure with 8 entries will be inserted into \(n_3\), violating the preferred property.
+
 \section*{3 - Name Services}

### 2019-11-15 assignment2/submission2.tex
Fix solution for 2.1aiii

@@
 	\item \texttt{[5, 4, 3, 4, 5]}
-	\item \texttt{[5, 4, 3, 2, 1]} \emph{or} \texttt{[1, 2, 3, 4, 5]} \emph{or} \texttt{[5, 4, 3, 4, 5]}, depending on the exact semantic behaviour of the copy and restore operations.
-	Given that the remote server passes the arguments by reference, the first question is whether \textit{x} and \textit{y} have separate copies of \textit{a}. 
-	If they do, the first or second option is possible. 
-	If there are no separate copies of \textit{a}, the procedure behaves as if the parameters were passed by reference. 
-	Otherwise it depends on which value is restored first.
+	\item \texttt{[5, 4, 3, 2, 1]} \emph{or} \texttt{[1, 2, 3, 4, 5]} depending on the order of restore operations.
+	The first option appears when \textit{y} is restored before \textit{x}.
+	The second option appears when the restoration order is reversed.
 \end{enumerate}

### 2019-11-16 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/common/MessageWriter.java
Add working solution for practical exercise 2.6

@@ -2,7 +2,6 @@ package de.unistgt.ipvs.vs.ex1.common;
 
 
 import java.io.IOException;
-import java.io.ObjectOutputStream;
 import java.io.OutputStream;
 
 /**

### 2019-11-16 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex1/server/CalcSocketServer.java

@@ -2,7 +2,6 @@ package de.unistgt.ipvs.vs.ex1.server;
 
 import java.io.IOException;
 import java.net.ServerSocket;
-import java.net.Socket;
 
 /**
  * Extend the run-method of this class as necessary to complete the assignment.

### 2019-11-16 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex2/client/CalcRmiClient.java

@@ -3,9 +3,7 @@ package de.unistgt.ipvs.vs.ex2.client;
 import de.unistgt.ipvs.vs.ex2.common.ICalculation;
 import de.unistgt.ipvs.vs.ex2.common.ICalculationFactory;
 
-import java.net.MalformedURLException;
 import java.rmi.Naming;
-import java.rmi.NotBoundException;
 import java.rmi.RemoteException;
 import java.util.Collection;
 
@@ -32,10 +30,11 @@ public class CalcRmiClient {
 
     public boolean init(String url) {
         try {
-            calc = (ICalculation) Naming.lookup(url);
+            calc = ((ICalculationFactory) Naming.lookup(url)).getSession();
         } catch (Exception e) {
             System.out.println("calc is not set, therefore everything will crash");
             e.printStackTrace();
+            return false;
         }
         return true;
     }

### 2019-11-16 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex2/server/CalcRmiServer.java

@@ -1,13 +1,11 @@
 package de.unistgt.ipvs.vs.ex2.server;
 
-import de.unistgt.ipvs.vs.ex2.common.ICalculation;
 import de.unistgt.ipvs.vs.ex2.common.ICalculationFactory;
 
 import java.rmi.Naming;
 import java.rmi.RemoteException;
 import java.net.MalformedURLException;
 import java.rmi.NotBoundException;
-import java.rmi.registry.LocateRegistry;
 import java.util.logging.Level;
 import java.util.logging.Logger;
 
@@ -23,7 +21,8 @@ public class CalcRmiServer extends Thread {
     public CalcRmiServer(String regHost, String objName) {
         this.regHost = regHost;
         this.objName = objName;
-        this.url = "rmi://" + regHost + "/" + objName;
+        // do not specify scheme, because the unit-tests don't specify it either.
+        this.url = "//" + regHost + "/" + objName;
     }
 
     @Override
@@ -32,15 +31,9 @@ public class CalcRmiServer extends Thread {
             System.err.println("<registryHost> or <objectName> not set!");
             return;
         }
-        //Add solution here
         try {
-            LocateRegistry.getRegistry();
-        } catch (RemoteException e) {
-            // should not happen, otherwise port is already used
-        }
-        try {
-            ICalculation srv = new CalculationImpl();
-            Naming.rebind(this.url, srv);
+            ICalculationFactory srv = new CalculationImplFactory();
+            Naming.bind(url, srv);
             System.out.println("Server binding successful");
         } catch (Exception e) {
             e.printStackTrace();

### 2019-11-16 assignment2/handout/src/main/java/de/unistgt/ipvs/vs/ex2/server/CalculationImplFactory.java

@@ -1,13 +1,32 @@
 package de.unistgt.ipvs.vs.ex2.server;
 
+import java.rmi.RemoteException;
+import java.rmi.server.UnicastRemoteObject;
+
+import de.unistgt.ipvs.vs.ex2.common.ICalculation;
+import de.unistgt.ipvs.vs.ex2.common.ICalculationFactory;
 
 /**
- * Change this class (implementation/signature/...) as necessary to complete the assignment.
- * You may also add some fields or methods.
+ * Change this class (implementation/signature/...) as necessary to complete the
+ * assignment. You may also add some fields or methods.
  */
 
-public class CalculationImplFactory {
+public class CalculationImplFactory extends UnicastRemoteObject implements ICalculationFactory {
+    
     private static final long serialVersionUID = 8409100566761383094L;
-
-
+    
+    /**
+     * 0-args constructor for inheritance purposes. Auto-generated 0-args ctor would not handle RemoteException
+     * @throws RemoteException
+     */
+    protected CalculationImplFactory() throws RemoteException {
+        super();
+    }
+    
+    @Override
+    public ICalculation getSession() throws RemoteException {
+        // we can return a CalculationImpl directly, because it inherits from Remote.
+        // it is therefore marshalled across RMI boundaries by reference.
+        return new CalculationImpl();
+    }
 }

### 2019-11-16 assignment2/submission2.tex
Fix small mistakes in submission2

@@
 \usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=4cm]{geometry}
-\usepackage[ngerman]{babel}
+\usepackage[english]{babel}
 \usepackage{amsmath, amsthm, amssymb, amstext}
@@
 As outlined in the forum, it's assumed that the node ids within the system are at least vaguely evenly distributed under the given hash function.
-Given that assumption and the assumption that there are at least three nodes in the system, it's reasonable to assume that no node is responsible for an adress space larger than \(2^m-1\).
-Carrying that assumption onwards, any entity will be stored on two nodes, because the smallest \(p \leq ID_1(e)\) will resolve to different nodes in the separate chord rings.
+Given that assumption and the assumption that there are at least three nodes in the system, it's reasonable to assume that no node is responsible for an address space larger than \(2^m-1\).
+As \(2^m-1\) is also the offset between the two hash ids it is obvious that between the two ids must be at least one node.
+And if a node is between the two ids different nodes must be responsible for the ids.
 
@@
 That failure can be seen when considering \(ID_1(n_1) + 2 = ID_1(n_2) + 1 = ID_1(n_3)\) for three nodes \(\{n_1, n_2, n_3\}\).
-It's likely that an inserted entity on a chord-ring structure with 8 entries will be inserted into \(n_3\), violating the preferred property.
+It's likely that an inserted entity on a chord-ring structure with 8 entries will be inserted into \(n_1\), violating the preferred property.
 

### 2019-11-25 assignment3/submission3.tex
add ex3 files

--- 2019-11-16 assignment2/submission2.tex
@@
 \newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
-\newcommand{\getsubtitle}{G2T1 – Assignment 2 (theoretical part)}
+\newcommand{\getsubtitle}{G2T1 – Assignment 3 (theoretical part)}
 \newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
@@
 
-\section*{1 - Parameter Passing and RMI}
+\section*{1 - Physical Clocks}
 \subsection*{a)}
-\begin{enumerate}[label=(\roman*)]
-	% no return value, so a is never changed
-	\item \texttt{[1, 2, 3, 4, 5]}
-	% the code given does not perform a swap, it only overwrites the lower indices with values from the upper ones.
-	% because of reference semantics that overwrites the second input as well
-	\item \texttt{[5, 4, 3, 4, 5]}
-	\item \texttt{[5, 4, 3, 2, 1]} \emph{or} \texttt{[1, 2, 3, 4, 5]} depending on the order of restore operations.
-	The first option appears when \textit{y} is restored before \textit{x}.
-	The second option appears when the restoration order is reversed.
-\end{enumerate}
 
@@
 
-The result printed to the system console is:
-\begin{verbatim}
-$ java StringServer
-1: B
-2: BRS
-3: ABSBRS
-\end{verbatim}
+\section*{2 - Logical Clocks}
+\subsection*{a)}
 
-We can deduce that all objects that extend UnicastRemoteObject are sent as reference, all other objects are sent as value.
-In the first and last two lines the parameter is a StringBuffer object which does not implement UnicastRemoteObject and is therefore sent as value. 
-In the third line the parameter is a StringServer object which does implement UnicastRemoteObject and because of this is sent as reference.
+\subsection*{b)}
+
+\section*{3 - Global State}
+\subsection*{a)}
+
+\subsection*{b)}
 
@@
 
-The first line shows that even "complex java objects" are passed by value so long as they are not remote objects. 
-This implies that even very large non-remote objects would be serialized and passed by value (e.g. Arrays with hundreds of thousands of elements).
-Such a behaviour incurs large overheads, especially if only parts of the passed value are accessed by the remote procedure.
-
-The lines two and three demonstrate the reference-passing behaviour within java's RMI implementation. 
-Notably, remote objects are passed by reference. 
-Calls to the passed remote objects are routed back to the JVM the actual object resides in. 
-This can lead to large overheads if a lot of calls are made to the referenced object.
-
-\section*{2 - Chord System}
+\section*{4 - Snapshot Algorithm}
 \subsection*{a)}
 
-	\begin{minipage}[t]{.25\linewidth}
-        \centering
-        Node 4:\\
-			\begin{tabular}{|l|l|}
-				\hline
-				1 & 11 \\ \hline
-				2 & 11 \\ \hline
-				3 & 11 \\ \hline
-				4 & 16 \\ \hline
-				5 & 23 \\ \hline
-			\end{tabular}
-	\end{minipage}%
-	\begin{minipage}[t]{.25\linewidth}
-        \centering
-        Node 11:\\
-		\begin{tabular}{|l|l|}
-			\hline
-			1 & 16 \\ \hline
-			2 & 16 \\ \hline
-			3 & 16 \\ \hline
-			4 & 19 \\ \hline
-			5 & 31 \\ \hline
-		\end{tabular}
-	\end{minipage}%
-	\begin{minipage}[t]{.25\linewidth}
-        \centering
-        Node 16:\\
-		\begin{tabular}{|l|l|}
-			\hline
-			1 & 19 \\ \hline
-			2 & 19 \\ \hline
-			3 & 23 \\ \hline
-			4 & 26 \\ \hline
-			5 & 4 \\ \hline
-		\end{tabular}
-    \end{minipage}
-    \begin{minipage}[t]{.25\linewidth}
-        \centering
-        Node 19:\\
-        \begin{tabular}{|l|l|}
-            \hline
-            1 & 23 \\ \hline
-            2 & 23 \\ \hline
-            3 & 23 \\ \hline
-            4 & 31 \\ \hline
-            5 & 4 \\ \hline
-        \end{tabular}
-    \end{minipage}%
-
-	\begin{minipage}[t]{.25\linewidth}
-		\centering
-        Node 23:\\
-		\begin{tabular}{|l|l|}
-			\hline
-			1 & 26 \\ \hline
-			2 & 26 \\ \hline
-			3 & 31 \\ \hline
-			4 & 31 \\ \hline
-			5 & 11 \\ \hline
-		\end{tabular}
-	\end{minipage}%
-	\begin{minipage}[t]{.25\linewidth}
-		\centering
-        Node 26:\\
-		\begin{tabular}{|l|l|}
-			\hline
-			1 & 31 \\ \hline
-			2 & 31 \\ \hline
-			3 & 31 \\ \hline
-			4 & 4 \\ \hline
-			5 & 11 \\ \hline
-		\end{tabular}
-	\end{minipage}
-	\begin{minipage}[t]{.25\linewidth}
-		\centering
-        Node 31:\\
-		\begin{tabular}{|l|l|}
-			\hline
-			1 & 4 \\ \hline
-			2 & 4 \\ \hline
-			3 & 4 \\ \hline
-			4 & 11 \\ \hline
-			5 & 16 \\ \hline
-		\end{tabular}
-	\end{minipage}
-
 \subsection*{b)}
-\begin{enumerate}
-	\item node 4 $ \rightarrow $ node 16
-	\item node 16 $ \rightarrow $ node 19
-	\item node 19 $ \rightarrow $ node 23
-\end{enumerate}
-
-\subsection*{c)}
-\begin{minipage}[t]{.25\linewidth}
-	\centering
-    Node 24:\\
-	\begin{tabular}{|l|l|}
-		\hline
-		1 & 26 \\ \hline
-		2 & 26 \\ \hline
-		3 & 31 \\ \hline
-		4 & 4 \\ \hline
-		5 & 11 \\ \hline
-	\end{tabular}
-\end{minipage}
-
-The finger table of node 16 will change eventually while the one for node 23 will change immediately.
-
-\subsection*{d)}
-If both nodes join simultaneously they are between the same two nodes (4 and 11). 
-If they now try to each update their predecessor and successor then we have a race condition while updating a linked list. 
-This can leave the list (or in this case ring) in an inconsistent state that makes routing impossible!
-
-\subsection*{e)}
-
-This code updates the FT entries of \textit{p} for the former successor's predecessor which has the id \textit{q}.
-
-\begin{algorithm}
-\begin{algorithmic}
-	\For{$i \gets 0 \textbf{ to } \lfloor log_2(q - p) \rfloor$}
-	   \State $FT_p[i] \gets q$
-	\EndFor
-\end{algorithmic}
-\end{algorithm}
-
-This code assumes that only \textit{q} has been inserted between \textit{p} and the former successor.
-The obvious issue arising from that (namely that multiple insertions are not correctly handled), can be easily remedied by repeating the consistency check after updating the finger table.
-As presented the code is already eventually consistent, though.
-
-\subsection*{f)}
-
-As outlined in the forum, it's assumed that the node ids within the system are at least vaguely evenly distributed under the given hash function.
-Given that assumption and the assumption that there are at least three nodes in the system, it's reasonable to assume that no node is responsible for an address space larger than \(2^m-1\).
-As \(2^m-1\) is also the offset between the two hash ids it is obvious that between the two ids must be at least one node.
-And if a node is between the two ids different nodes must be responsible for the ids.
-
-This property holds true under the assumption that no node is responsible for an adress space larger than the largest lookup shortcut, which is the offset for the second chord ring.
-If that property is not given, the insert is not guaranteed to store an entity's copy on two different nodes.
-That failure can be seen when considering \(ID_1(n_1) + 2 = ID_1(n_2) + 1 = ID_1(n_3)\) for three nodes \(\{n_1, n_2, n_3\}\).
-It's likely that an inserted entity on a chord-ring structure with 8 entries will be inserted into \(n_1\), violating the preferred property.
-
-\section*{3 - Name Services}
-
-\subsection*{a)}
-\begin{center}
-	\begin{tabular}{|c|c|c|}
-		\hline
-		& Iterative resolution & Recursive resolution \\ \hline
-		1. Lookup & 10 & 10 \\ \hline
-		2. Lookup & 10 & 10 \\ \hline
-		3. Lookup & 8 & 8 \\ \hline
-	\end{tabular}
-\end{center}
-
-\subsection*{b)}
-\begin{center}
-	\begin{tabular}{|c|c|c|}
-		\hline
-		& Iterative resolution & Recursive resolution \\ \hline
-		1. Lookup & 10 & 10 (8 fast, 2 slow) \\ \hline
-		2. Lookup & 10 & 8 (6 fast, 2 slow) \\ \hline
-		3. Lookup & 8 & 8 (6 fast, 2 slow) \\ \hline
-	\end{tabular}
-\end{center}
-
-\subsection*{c)}
-
-\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
-\subsubsection*{i.}
-\begin{lstlisting}
-$ dig NS uni-stuttgart.de
-...
-;; ANSWER SECTION:
-uni-stuttgart.de.	2191	IN	NS	minnehaha.rhrk.uni-kl.de.
-uni-stuttgart.de.	2191	IN	NS	dns3.belwue.de.
-uni-stuttgart.de.	2191	IN	NS	dns1.belwue.de.
-uni-stuttgart.de.	2191	IN	NS	dns0.uni-stuttgart.de.
-uni-stuttgart.de.	2191	IN	NS	dns1.uni-stuttgart.edu.
-...
-\end{lstlisting}
-There are 5 server responsible for the DNS entry of 'uni-stuttgart.de':
-\begin{itemize}
-	\item minnehaha.rhrk.uni-kl.de
-	\item dns3.belwue.de
-	\item dns1.belwue.de
-	\item dns0.uni-stuttgart.de
-	\item dns1.uni-stuttgart.edu
-\end{itemize}
-\subsubsection*{ii.}
-\begin{lstlisting}
-$ dig -x 129.69.216.249
-...
-;; ANSWER SECTION:
-249.216.69.129.in-addr.arpa. 86397 IN	PTR	ipvslogin.informatik.uni-stuttgart.de.
-...
-\end{lstlisting}
-PTR is usefull for resolving an IP address to a domain or hostname (reverse DNS lookup).\\
-The given IP resolves to: 'ipvslogin.informatik.uni-stuttgart.de'
-\subsubsection*{iii.}
-\begin{lstlisting}
-$ dig +trace www.uni-stuttgart.de
-...
-.			347153	IN	NS	h.root-servers.net.
-.			347153	IN	NS	l.root-servers.net.
-.			347153	IN	NS	d.root-servers.net.
-.			347153	IN	NS	m.root-servers.net.
-.			347153	IN	NS	e.root-servers.net.
-.			347153	IN	NS	k.root-servers.net.
-.			347153	IN	NS	i.root-servers.net.
-.			347153	IN	NS	b.root-servers.net.
-.			347153	IN	NS	c.root-servers.net.
-.			347153	IN	NS	a.root-servers.net.
-.			347153	IN	NS	f.root-servers.net.
-.			347153	IN	NS	g.root-servers.net.
-.			347153	IN	NS	j.root-servers.net.
-...
-;; Received 553 bytes from 192.168.0.1#53(192.168.0.1) in 28 ms
-
-de.			172800	IN	NS	z.nic.de.
-de.			172800	IN	NS	f.nic.de.
-de.			172800	IN	NS	s.de.net.
-de.			172800	IN	NS	a.nic.de.
-de.			172800	IN	NS	l.de.net.
-de.			172800	IN	NS	n.de.net.
-...
-;; Received 802 bytes from 2001:500:12::d0d#53(g.root-servers.net) in 22 ms
-
-uni-stuttgart.de.	86400	IN	NS	dns0.uni-stuttgart.de.
-uni-stuttgart.de.	86400	IN	NS	dns1.belwue.de.
-uni-stuttgart.de.	86400	IN	NS	dns1.uni-stuttgart.edu.
-uni-stuttgart.de.	86400	IN	NS	dns3.belwue.de.
-uni-stuttgart.de.	86400	IN	NS	minnehaha.rhrk.uni-kl.de.
-...
-;; Received 619 bytes from 2001:67c:1011:1::53#53(n.de.net) in 25 ms
-
-www.uni-stuttgart.de.	3600	IN	A	129.69.8.19
-...
-uni-stuttgart.de.	3600	IN	NS	dns0.uni-stuttgart.de.
-uni-stuttgart.de.	3600	IN	NS	minnehaha.rhrk.uni-kl.de.
-uni-stuttgart.de.	3600	IN	NS	dns1.uni-stuttgart.edu.
-uni-stuttgart.de.	3600	IN	NS	dns3.belwue.de.
-uni-stuttgart.de.	3600	IN	NS	dns1.belwue.de.
-...
-;; Received 1365 bytes from 129.143.2.10#53(dns1.belwue.de) in 25 ms
-
-\end{lstlisting}
-The path the lookup took was:
-\begin{enumerate}
-	\item 192.168.0.1
-	\item g.root-servers.net
-	\item n.de.net
-	\item dns1.belwue.de
-\end{enumerate}
 

### 2019-11-27 assignment3/submission3.tex
Solution for Task 2 added

@@
 \subsection*{a)}
+\subsection*{a)}
+\begin{enumerate}[label=(\roman*)]
+	\item 
+	C(e$^1_1$) = 2\\
+	C(e$^2_1$) = 3\\
+	C(e$^3_1$) = 4\\
+	C(e$^4_1$) = 5\\
 
+	C(e$^1_2$) = 5\\
+	C(e$^2_2$) = 6\\
+	C(e$^3_2$) = 7\\
+	C(e$^4_2$) = 8\\
+
+	C(e$^1_3$) = 1\\
+	C(e$^2_3$) = 2\\
+	C(e$^3_3$) = 3\\
+	C(e$^4_3$) = 9\\
+	\item 
+	VC(e$^1_1$) = \texttt{[1 0 1]}\\
+	VC(e$^2_1$) = \texttt{[2 0 1]}\\
+	VC(e$^3_1$) = \texttt{[3 0 1]}\\
+	VC(e$^4_1$) = \texttt{[4 0 1]}\\
+
+	VC(e$^1_2$) = \texttt{[3 1 1]}\\
+	VC(e$^2_2$) = \texttt{[3 2 2]}\\
+	VC(e$^3_2$) = \texttt{[3 3 2]}\\
+	VC(e$^4_2$) = \texttt{[3 4 2]}\\
+
+	VC(e$^1_3$) = \texttt{[0 0 1]}\\
+	VC(e$^2_3$) = \texttt{[0 0 2]}\\
+	VC(e$^3_3$) = \texttt{[0 0 3]}\\
+	VC(e$^4_3$) = \texttt{[3 4 4]}\\
+	\item
+	P$_1$: e$^1_1$ e$^2_1$ e$^3_1$\\
+	P$_2$: e$^1_2$ e$^2_2$\\
+	P$_3$: e$^1_3$ e$^2_3$\\
+	The vector event time of those events is smaller than the vector event time of e$^3_2$. A smaller vector time means events are causally related.
+\end{enumerate}
 \subsection*{b)}
-
+\begin{enumerate}[label=(\roman*)]
+	\item 
+	P$_1$: e$_6$ e$_5$ e$_9$ e$_4$ e$_8$ e$_{10}$\\
+	P$_2$: e$_{12}$ e$_3$ e$_{13}$\\
+	P$_3$: e$_{11}$ e$_7$ e$_1$ e$_2$\\
+	\item 
+	Local event: e$_5$ e$_9$ e$_3$ e$_1$ e$_{11}$\\
+	Send event: e$_6$ e$_4$ e$_{10}$ e$_{12}$\\
+	Recieve event:  e$_8$ e$_{13}$ e$_7$ e$_2$\\
+\end{enumerate}
 \section*{3 - Global State}

### 2019-11-27 assignment3/submission3.tex
done 6.3
Solution to exercise 3.1a1

@@
 \usepackage {tikz}
-\usetikzlibrary {positioning}
+\usetikzlibrary{positioning}
 
@@
 
+\begin{figure}[ht!]
+\begin{tikzpicture}[scale=0.15]
+\foreach \x in {1,...,97} {
+	\draw (\x,0) -- (\x, 2);
+	\draw (\x, -5) -- (\x, -7);
+};
+\draw (0,0) -- (97,0);
+\draw (0,-5) -- (97,-5);
+\node at(0, -1) {\tiny Master};
+\node at(0, -4) {\tiny Slave};
+\foreach \x/\n in {3/7,9/8,15/9,21/10,27/11,33/12,39/13,45/14,51/15,57/16,63/17,69/18,75/19,81/20,87/21,93/22} {
+	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,1) (a\n) {}
+		node[below=7mm, above of=a\n] {\tiny \n};
+};
+\foreach \x/\n in {3/3,9/4,15/5,21/6,27/7,33/8,37/9,41/10,45/11,49/12,53/13,57/14,61/15,65/16,69/17,73/18,77/19,81/20,87/21,93/22} {
+	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,-6) (b\n) {}
+		node[above=7mm, below of=b\n] {\tiny \n};
+};
+\path[draw,->] 
+	(b5.north) edge (a10.south)
+	(a11.south) edge (b8.north);
+\end{tikzpicture}
+\end{figure}
+\begin{align*}
+O = \frac12((t_2 - t_1) - (t_4 - t_3)) = \frac12((10 - 5) - (8 - 11)) = \frac12(5 + 3) = 4  &\\
+d = 1 & \\
+\Delta = 12, C_s(t_4 + \Delta) = 20 & \\
+\frac{dC_S}{dt} = \frac{12 + 4}{12} = \frac{4}{3} &
+\end{align*}
+
 \subsection*{b)}
@@
 \section*{2 - Logical Clocks}
-\subsection*{a)}
 \subsection*{a)}
@@
 \end{enumerate}
+
 \section*{3 - Global State}
 \subsection*{a)}
+$ \{(e^{1}_{1}, e^{1}_{2}), (e^{2}_{1}, e^{2}_{2}), (e^{2}_{2}, e^{3}_{1}), (e^{4}_{1}, e^{3}_{2}), (e^{4}_{1}, e^{4}_{2}), (e^{5}_{1}, e^{5}_{2}), (e^{1}_{1}, e^{2}_{2}), (e^{2}_{2}, e^{4}_{1}), (e^{4}_{1}, e^{5}_{2})\} $
 
 \subsection*{b)}
+\begin{enumerate}
+	\item yes, as it is a valid linearization.
+	\item no, because $ e_{2}^{3} $ can not happen before $ e_{1}^{3} $
+\end{enumerate}
 
 \subsection*{c)}
+\begin{figure}[!ht]
+	\centering
+	\begin{tikzpicture}[auto,node distance=1.5cm]
+		\node[draw] (s00) {$ S_{00} $};
+		\node[draw] (s10) [below left = of s00] {$ S_{10} $};
+		\node[draw] (s01) [below right = of s00] {$ S_{01} $};
+		\node[draw] (s11) [below right = of s10] {$ S_{11} $};
+		\node[draw] (s02) [below right = of s01] {$ S_{02} $};
 
+		\node[draw] (s21) [below left = of s11] {$ S_{21} $};
+		\node[draw] (s12) [below left = of s02] {$ S_{12} $};
+
+		\node[draw] (s31) [below left = of s21] {$ S_{31} $};
+		\node[draw] (s22) [below left = of s12] {$ S_{22} $};
+
+		\node[draw] (s41) [below left = of s31] {$ S_{41} $};
+		\node[draw] (s32) [below left = of s22] {$ S_{32} $};
+
+		\node[draw] (s42) [below right = of s41] {$ S_{42} $};
+		\node[draw] (s33) [below right = of s32] {$ S_{33} $};
+
+		\node[draw] (s43) [below right = of s42] {$ S_{43} $};
+		\node[draw] (s34) [below right = of s33] {$ S_{34} $};
+
+		\node[draw] (s44) [below right = of s43] {$ S_{44} $};
+		\node[draw] (s35) [below right = of s34] {$ S_{35} $};
+
+		\node[draw] (s54) [below left = of s44] {$ S_{54} $};
+		\node[draw] (s45) [below left = of s35] {$ S_{45} $};
+
+		\node[draw] (s55) [below right = of s54] {$ S_{55} $};
+
+		\path (s10) edge node {$ e^{1}_{1} $} (s00);
+		\path (s11) edge node {$ e^{1}_{1} $} (s01);
+		\path (s12) edge node {$ e^{1}_{1} $} (s02);
+
+		\path (s00) edge node {$ e^{1}_{2} $} (s01);
+		\path (s10) edge node {$ e^{1}_{2} $} (s11);
+
+		\path (s21) edge node {$ e^{2}_{1} $} (s11);
+		\path (s22) edge node {$ e^{2}_{1} $} (s12);
+
+		\path (s41) edge node {$ e^{2}_{2} $} (s42);
+		\path (s31) edge node {$ e^{2}_{2} $} (s32);
+		\path (s21) edge node {$ e^{2}_{2} $} (s22);
+		\path (s11) edge node {$ e^{2}_{2} $} (s12);
+		\path (s01) edge node {$ e^{2}_{2} $} (s02);
+
+		\path (s31) edge node {$ e^{3}_{1} $} (s21);
+		\path (s32) edge node {$ e^{3}_{1} $} (s22);
+
+		\path (s32) edge node {$ e^{3}_{2} $} (s33);
+		\path (s42) edge node {$ e^{3}_{2} $} (s43);
+
+		\path (s41) edge node {$ e^{4}_{1} $} (s31);
+		\path (s42) edge node {$ e^{4}_{1} $} (s32);
+		\path (s43) edge node {$ e^{4}_{1} $} (s33);
+		\path (s44) edge node {$ e^{4}_{1} $} (s34);
+		\path (s45) edge node {$ e^{4}_{1} $} (s35);
+
+		\path (s33) edge node {$ e^{4}_{2} $} (s34);
+		\path (s43) edge node {$ e^{4}_{2} $} (s44);
+
+		\path (s34) edge node {$ e^{5}_{2} $} (s35);
+		\path (s44) edge node {$ e^{5}_{2} $} (s45);
+		\path (s54) edge node {$ e^{5}_{2} $} (s55);
+
+		\path (s54) edge node {$ e^{5}_{1} $} (s44);
+		\path (s55) edge node {$ e^{5}_{1} $} (s45);
+	\end{tikzpicture}
+	\caption{solution for 3.c)}
+\end{figure}
+
+\newpage
 \section*{4 - Snapshot Algorithm}

### 2019-12-01 assignment3/submission3.tex
Solution to Exercise 3.1a2
Add solution to exercise 4

@@
 \end{tikzpicture}
+\caption{Solution for Figure 1}
 \end{figure}
@@
 \frac{dC_S}{dt} = \frac{12 + 4}{12} = \frac{4}{3} &
+\end{align*}
+
+\begin{figure}[ht!]
+	\begin{tikzpicture}[scale=0.15]
+	\foreach \x in {1,...,97} {
+		\draw (\x,0) -- (\x, 2);
+		\draw (\x, -5) -- (\x, -7);
+	};
+	\draw (0,0) -- (97,0);
+	\draw (0,-5) -- (97,-5);
+	\node at(0, -1) {\tiny Master};
+	\node at(0, -4) {\tiny Slave};
+	\foreach \x/\n in {3/1,9/2,15/3,21/4,27/5,33/6,39/7,45/8,51/9,57/10,63/11,69/12,75/13,81/14,87/15,93/16} {
+		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,1) (a\n) {}
+			node[below=7mm, above of=a\n] {\tiny \n};
+	};
+	\foreach \x/\n in {3/6,9/7,15/8,21/9,27/10,33/11,49/12,65/13,81/14,87/15,93/16} {
+		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,-6) (b\n) {}
+			node[above=7mm, below of=b\n] {\tiny \n};
+	};
+	\path[draw,->]
+		(b8.north) edge (a4.south)
+		(a5.south) edge (b11.north);
+\end{tikzpicture}
+\caption{Solution for Figure 2}
+\end{figure}
+\begin{align*}
+O = \frac12((t_2 - t_1) - (t_4 - t_3)) = \frac12((4 - 8) - (11 - 5)) = \frac12(-4 - 6) = -5  &\\
+d = 1 & \\
+\Delta = 3, C_s(t_4 + \Delta) = 14 & \\
+\frac{dC_S}{dt} = \frac{3}{3 +5} = \frac{3}{8} &
 \end{align*}
@@
 
+See Figure 4
+
+\begin{figure}[!ht]
+    \centering
+    \begin{tikzpicture}[auto,node distance=1.5cm,event/.style={fill=black,circle, inner sep=0pt, minimum size=5pt},state/.style={draw,rectangle, red, color=red, inner sep=0pt, minimum size=5pt, minimum width=10pt}, msg/.style={-latex,thick}, marker/.style={-latex,thick,red}]
+        \node[] (p1s) {\(P_1\)};
+        \node[] (p2s) [below = of p1s] {\(P_2\)};
+        \node[] (p3s) [below = of p2s] {\(P_3\)};
+
+        \node[] (p1e) [right = 13cm of p1s] {};
+        \node[] (p2e) [right = 13cm of p2s] {};
+        \node[] (p3e) [right = 13cm of p3s] {};
+
+        \path (p1s.east) edge (p1e);
+        \path (p2s.east) edge (p2e);
+        \path (p3s.east) edge (p3e);
+
+        \node[event,label={\(e_1^1\)}] (e11) [right = 1.3cm of p1s] {};
+        \node[event,label={\(e_2^1\)}] (e21) [right = 1cm of p2s] {};
+        \node[event,label={\(e_3^1\)}] (e31) [right = 2cm of p3s] {};
+
+        \node[event,label={\(e_1^2\)}] (e12) [right = 4.2cm of p1s] {};
+        \node[event,label={\(e_2^2\)}] (e22) [right = 2.1cm of p2s] {};
+        \node[event,label={\(e_3^2\)}] (e32) [right = 4cm of p3s] {};
+
+        \node[event,label={\(e_1^3\)}] (e13) [right = 6cm of p1s] {};
+        \node[event,label={\(e_2^3\)}] (e23) [right = 4.5cm of p2s] {};
+        \node[event,label={\(e_3^3\)}] (e33) [right = 10cm of p3s] {};
+
+        \node[event,label={\(e_1^4\)}] (e14) [right = 7cm of p1s] {};
+        \node[event,label=below:{\(e_2^4\)}] (e24) [right = 8.5cm of p2s] {};
+
+        \node[event,label={\(e_1^5\)}] (e15) [right = 10.5cm of p1s] {};
+        \node[event,label={\(e_2^5\)}] (e25) [right = 9cm of p2s] {};
+
+        \node[event,label={\(e_2^6\)}] (e26) [right = 12.5cm of p2s] {};
+
+        \path (e21) edge[msg] node {\(m_1\)} (e31);
+        \path (e11) edge[msg] node[left] {\(m_2\)} (e23);
+        \path (e22) edge[msg] node[right] {\(m_3\)} (e12);
+        \path (e32) edge[msg] node[below right] {\(m_4\)} (e14);
+        \path (e13) edge[msg] node[right] {\(m_5\)} (e25);
+        \path (e15) edge[msg] node {\(m_6\)} (e26);
+        \path (e24) edge[msg] node {\(m_7\)} (e33);
+
+        \node[state,label={[red]:{\(s_3\)}}] (s3) [right = 3cm of p3s] {};
+        \node[state,label={[red]:{\(s_1,c_4\)}}] (s1) [right = 5cm of p1s] {};
+        \node[state,label={[red]below:{\(s_2,c_1\)}}] (s2) [right = 6.5cm of p2s] {};
+
+        \node[state,label={[red]:{\(c_2\)}}] (c2) [right = 8cm of p1s] {};
+        \node[state,label={[red]:{\(c_3\)}}] (c3) [right = 8.5cm of p3s] {};
+
+        \path (s3.north east) edge[marker] node[below] {\(M\)} (s1.south west);
+        \path (s1.south east) edge[marker] node[left] {\(M'\)} (s2.north west);
+        \path (s2.north east) edge[marker] node[below left] {\(M''\)} (c2.south west);
+        \path (s2.south east) edge[marker] node[] {\(M''\)} (c3.north west);
+    \end{tikzpicture}
+
+    The red boxes represent where a process saves some state.
+A process state save is denoted by \(s_i\) and a channel state save by \(c_j\) where \(i\) is the process number and \(j\) is the channel number.
+    \caption{solution for 4.a)}
+\end{figure}
+
 \subsection*{b)}
 
+No channel received a message between the process starting recording and the first/second marker arriving. All channel states are empty sets.
+
+\begin{align*}
+    c_1 &= \{\}\\
+    c_2 &= \{\}\\
+    c_3 &= \{\}\\
+    c_4 &= \{\}\\
+\end{align*}
+
 \end{document}

### 2019-12-01 assignment3/submission3.tex
Correct solutions to Exercise 1

@@
 \begin{figure}[ht!]
-\begin{tikzpicture}[scale=0.15]
-\foreach \x in {1,...,97} {
+\begin{tikzpicture}[scale=0.2]
+\foreach \x in {1,...,81} {
 	\draw (\x,0) -- (\x, 2);
@@
 };
-\draw (0,0) -- (97,0);
-\draw (0,-5) -- (97,-5);
+\draw (0,0) -- (81,0);
+\draw (0,-5) -- (81,-5);
 \node at(0, -1) {\tiny Master};
 \node at(0, -4) {\tiny Slave};
-\foreach \x/\n in {3/7,9/8,15/9,21/10,27/11,33/12,39/13,45/14,51/15,57/16,63/17,69/18,75/19,81/20,87/21,93/22} {
-	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,1) (a\n) {}
+\foreach \x/\n in {3/7,8/8,13/9,18/10,23/11,28/12,33/13,38/14,43/15,48/16,53/17,58/18,63/19,68/20,73/21,78/22} {
+	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,1) (a\n) {}
 		node[below=7mm, above of=a\n] {\tiny \n};
 };
-\foreach \x/\n in {3/3,9/4,15/5,21/6,27/7,33/8,37/9,41/10,45/11,49/12,53/13,57/14,61/15,65/16,69/17,73/18,77/19,81/20,87/21,93/22} {
-	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,-6) (b\n) {}
+\foreach \x/\n in {3/3,8/4,13/5,18/6,23/7,28/8,31/9,34/10,37/11,40/12,43/13,46/14,49/15,52/16,55/17,58/18,63/19,68/20,73/21,78/22} {
+	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,-6) (b\n) {}
 		node[above=7mm, below of=b\n] {\tiny \n};
@@
 d = 1 & \\
-\Delta = 12, C_s(t_4 + \Delta) = 20 & \\
-\frac{dC_S}{dt} = \frac{12 + 4}{12} = \frac{4}{3} &
+\Delta = 6, C_s(t_4 + \Delta) = 18 & \\
+\frac{dC_S}{dt} = \frac{6 + 4}{6} = \frac{5}{3} &
 \end{align*}
@@
 \begin{figure}[ht!]
-	\begin{tikzpicture}[scale=0.15]
-	\foreach \x in {1,...,97} {
+	\begin{tikzpicture}[scale=0.2]
+	\foreach \x in {1,...,81} {
 		\draw (\x,0) -- (\x, 2);
@@
 	};
-	\draw (0,0) -- (97,0);
-	\draw (0,-5) -- (97,-5);
+	\draw (0,0) -- (81,0);
+	\draw (0,-5) -- (81,-5);
 	\node at(0, -1) {\tiny Master};
 	\node at(0, -4) {\tiny Slave};
-	\foreach \x/\n in {3/1,9/2,15/3,21/4,27/5,33/6,39/7,45/8,51/9,57/10,63/11,69/12,75/13,81/14,87/15,93/16} {
-		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,1) (a\n) {}
+	\foreach \x/\n in {3/1,8/2,13/3,18/4,23/5,28/6,33/7,38/8,43/9,48/10,53/11,58/12,63/13,68/14,73/15,78/16} {
+		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,1) (a\n) {}
 			node[below=7mm, above of=a\n] {\tiny \n};
 	};
-	\foreach \x/\n in {3/6,9/7,15/8,21/9,27/10,33/11,49/12,65/13,81/14,87/15,93/16} {
-		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=3mm] at(\x + .5,-6) (b\n) {}
+	\foreach \x/\n in {3/6,8/7,13/8,18/9,23/10,28/11,41/12,55/13,68/14,73/15,78/16} {
+		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,-6) (b\n) {}
 			node[above=7mm, below of=b\n] {\tiny \n};

### 2019-12-01 assignment3/submission3.tex
Add solution for Exercise 3.1b

@@
 d = 1 & \\
-\Delta = 3, C_s(t_4 + \Delta) = 14 & \\
-\frac{dC_S}{dt} = \frac{3}{3 +5} = \frac{3}{8} &
+\Delta = 8, C_s(t_4 + \Delta) = 14 & \\
+\frac{dC_S}{dt} = \frac{8 - 5}{8} = \frac{3}{8} &
 \end{align*}
@@
 \subsection*{b)}
+\begin{enumerate}[label=(\roman*)]
+
+	\item We can use the difference in offsets to determine the clockrate of the master node in relation to the subservient node.
+	For that purpose we calculate \(O_1 = \frac12((t_6 - t_5) - (t_8 - t_7))\) and \(O_0 = \frac12((t_2 - t_1) - (t_4 - t_3))\).
+
+	It's clear that \(\frac{O_1 - O_0}{t_6 - t_2}\) is the rate at which the clocks are desynchronizing in reference to the master clock.
+	If \(O_1 - O_0\) is negative, the subservient clock is faster than the master, and vice versa.
+	This means that for negative \(O_1 - O_0\), the frequency ratio is \(f = \frac{t_6 - t_2}{O_0 - O_1}\).
+	For a faster master clock the frequency ratio is the desynchronization rate from before.
+
+	\item It makes no sense to synchronize the clock rates before the clock values, because to enable a ''stable'' value synchronization  -- i.e.
+monotonically increasing without large gaps in tick-time or clockvalues -- the clock frequency needs to be adjusted again.
+	As such, the synchronized clock frequency would need to be changed again to synchronize clock values, making it an unnecessary step.
+
+	\item Synchronization of the depicted clocks.
+	\begin{align*}
+		O_0 = \frac12((5-6)-(9-6)) = 1 & \\
+		O_1 = \frac12((8-11) - (14-9)) = -4 & \\
+		d = \frac12((8-11) + (14-9)) = 1 & \\
+		f = \frac{(8 - 5)}{1 - (-4)} = \frac35 &
+	\end{align*}
+
+	We choose \(\Delta = 8, C_s(t_8 + \Delta) = 18\) modifying the clock frequency of \(C_s\) to \(f * \frac{\Delta + O_1}{\Delta}\) of the original frequency.
+	When \(t_8 + \Delta\), we again modify the frequency by \(\frac{\Delta}{\Delta - O_1}\) to return to the synchronized clock freqency.
+
+	This means \(\frac{dC_s}{dt} = f * \frac{8 - 4}{8} = f * \frac12 = \frac{3}{10}\)
+
+	\item Completing the figure.
+	\begin{figure}[ht!]
+		\begin{tikzpicture}[scale=0.2]
+		\foreach \x in {1,...,81} {
+			\draw (\x,0) -- (\x, 2);
+			\draw (\x, -5) -- (\x, -7);
+		};
+		\draw (0,0) -- (81,0);
+		\draw (0,-5) -- (81,-5);
+		\node at(0, -1) {\tiny Master};
+		\node at(0, -4) {\tiny Slave};
+		\foreach \x/\n in {3/4,8/5,13/6,18/7,23/8,28/9,33/10,38/11,43/12,48/13,53/14,58/15,63/16,68/17,73/18,78/19} {
+			\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,1) (a\n) {}
+				node[below=7mm, above of=a\n] {\tiny \n};
+		};
+		\foreach \x/\n in {3/5,6/6,9/7,12/8,15/9,18/10,21/11,24/12,27/13,30/14,40/15,50/16,60/17,70/18,75/19,80/20} {
+			\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,-6) (b\n) {}
+				node[above=7mm, below of=b\n] {\tiny \n};
+		};
+		\path[draw,->]
+			(b6.north) edge (a5.south)
+			(a6.south) edge (b9.north)
+			(b11.north) edge (a8.south)
+			(a9.south) edge (b14.north);
+	\end{tikzpicture}
+	\caption{Solution for Figure 2}
+	\end{figure}
+
+	Note that the delay is not used in the timing schema proposed in (i).
+	It has been included in the calculations for (iii) for completeness' sake.
+\end{enumerate}
 

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/VectorClock.java
done 4.1 practical a)

@@ -21,25 +21,25 @@ public class VectorClock {
 	}
 
 	public void increment() {
-		// TODO
 		/*
 		 * Complete a code to increment the local clock component
 		 */
-
+		vectorClock[processId] += 1;
 	}
 
 	public int[] get() {
-		// TODO
 		// Complete a code to return the vectorClock value
-		return null;
+		return vectorClock;
 	}
 
 	public void update(VectorClock other) {
-		// TODO
 		/*
 		 * Implement Supermum operation
 		 */
-
+		this.increment();
+		for (int i = 0; i < vectorClock.length; i++) {
+			vectorClock[i] = Math.max(vectorClock[i], other.get()[i]);
+		}
 	}
 
 	public boolean checkConsistency(int otherProcessId, VectorClock other) {
@@ -48,9 +48,6 @@ public class VectorClock {
 		 * Implement a code to check if a state is consist regarding two vector clocks (i.e. this and other).
 		 * See slide 41 from global state lecture.
 		 */
-
-		return false;
-
+		return vectorClock[otherProcessId] == other.get()[otherProcessId];
 	}
-
 }

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Monitor.java
started 4.1 practical b)

@@ -182,13 +182,39 @@ public class Monitor implements Runnable {
 	 * @return list of all reachable states
 	 */
 	private LinkedList<State> findReachableStates(State state) {
-		// TODO
 		/*
 		 * Given a state, implement a code that find all reachable states. The
 		 * function should return a list of all reachable states
-		 *
 		 */
-		return null;
+		LinkedList<State> reachable = new LinkedList<State>();
+		// current vector clocks
+		ArrayList<VectorClock> vClocks = new ArrayList<VectorClock>();
+		for (int i = 0; i < numberOfProcesses; i++) {
+			VectorClock curr_vc = processesMessages.get(i).get(state.getProcessMessageCurrentIndex(i)).getVectorClock();
+			vClocks.add(i, curr_vc);
+		}
+		// for each process
+		outer: for (int i = 0; i < numberOfProcesses; i++) {
+			int next = state.getProcessMessageCurrentIndex(i) + 1;
+			// check if process has no more lines
+			if (processesMessages.get(i).size() <= next) {
+				continue;
+			}
+			VectorClock next_vc = processesMessages.get(i).get(next).getVectorClock();
+			for (int j = 0; j < numberOfProcesses; j++) {
+				if (i == j) {
+					continue;
+				} else if (!next_vc.checkConsistency(j, vClocks.get(j))) {
+					// skip if not consistent
+					continue outer;
+				}
+			}
+			// adds reachable state
+			int[] new_indexes = state.getProcessesMessagesCurrentIndex().clone();
+			new_indexes[i] += 1;
+			reachable.add(new State(new_indexes));
+		}
+		return reachable;
 	}
 
 	/**
@@ -213,8 +239,6 @@ public class Monitor implements Runnable {
 		 * Predicate.predicate0(process_i_Message, process_j_Message); break;
 		 * case 1: ... }
 		 */
-
 		 return false;
 	}
-
 }

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Predicate.java

@@ -11,20 +11,17 @@ public class Predicate {
 	}
 
 	static public boolean predicate1(Message process_i_Message, Message process_j_Message) {
-		// TODO
 		/*
 		 * Add you code here to implement predicate2 x1 - x2= 15
 		 */
-	 	return false;
+	 	return (process_i_Message.getLocalVariable() - process_j_Message.getLocalVariable() == 15);
 	}
 
 	static public boolean predicate2(Message process_i_Message, Message process_j_Message) {
-		// TODO
 		/*
 		 * Add you code here to implement predicate2 x1 + x2= 30
 		 */
-
-		return false;
+		return (process_i_Message.getLocalVariable() + process_j_Message.getLocalVariable() == 30);
 	}
 
 	static public boolean predicate3(Message process_i_Message, Message process_j_Message) {

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Predicate.java
done 4.1 practical c)

@@ -25,12 +25,10 @@ public class Predicate {
 	}
 
 	static public boolean predicate3(Message process_i_Message, Message process_j_Message) {
-		// TODO
 		/*
 		 * Add you code here to implement predicate1 x1- x3=8
 		 */
-
-		return false;
+		return (process_i_Message.getLocalVariable() - process_j_Message.getLocalVariable() == 8);
 	}
 
 }

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Process1.java

@@ -49,8 +49,19 @@ public class Process1 extends AbstractProcess {
 		message = new Message(new VectorClock(vectorClock), this.localVariable);
 		monitor.receiveMessage(this.Id, message);
 
-		// TODO add part (c) changes here!
-		
+		// line 6
+		send(2, message); // send to process 1
+
+		// line 7
+		// receive
+		receivedMessage = receive(1); // receive from process 1
+		this.vectorClock.update(receivedMessage.getVectorClock());
+		this.localVariable = receivedMessage.getLocalVariable() - this.localVariable;
+		this.vectorClock.increment();
+
+		// notify the monitor
+		message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
 
 		// send terminate signal
 		monitor.processTerminated(this.Id);

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Process3.java

@@ -14,7 +14,51 @@ public class Process3 extends AbstractProcess {
 	public void run() {
 
 		//TODO Implement processes3 (Listing 4) code here!
+		// send the initial state to Monitor
+		Message message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
 
+		// line 1: x3 := 4
+		this.localVariable = 4;
+		this.vectorClock.increment();
+		// notify the monitor
+		message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
+
+		// line 2: x3 := x3 * 2
+		this.localVariable = this.localVariable * 2;
+		this.vectorClock.increment();
+		// notify the monitor
+		message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
+
+		//line 3: x3 := receive () - x3
+		// receive
+		Message receivedMessage = receive(0); // receive from process 0
+		this.vectorClock.update(receivedMessage.getVectorClock());
+		this.localVariable = receivedMessage.getLocalVariable() - this.localVariable;
+		this.vectorClock.increment();
+		// notify the monitor
+		message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
+
+		// line 4: x3 := x3 - 2
+		this.localVariable = this.localVariable - 2;
+		this.vectorClock.increment();
+		// notify the monitor
+		message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
+
+		// line 5: x3 := x3 + 11
+		this.localVariable = this.localVariable + 11;
+		this.vectorClock.increment();
+		// notify the monitor
+		message = new Message(new VectorClock(vectorClock), this.localVariable);
+		monitor.receiveMessage(this.Id, message);
+
+		// line 5: send ( P1 , x3 )
+		send(0, message);
+
+		System.out.printf("process:%d , the local variable= %d\n", this.Id, this.localVariable);
 	}
-
 }

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Process1.java
fix practical 4.1 c)

@@ -49,19 +49,21 @@ public class Process1 extends AbstractProcess {
 		message = new Message(new VectorClock(vectorClock), this.localVariable);
 		monitor.receiveMessage(this.Id, message);
 
-		// line 6
-		send(2, message); // send to process 1
+		if (numberOfProcesses > 2) {
+			// line 6
+			send(2, message); // send to process 1
 
-		// line 7
-		// receive
-		receivedMessage = receive(1); // receive from process 1
-		this.vectorClock.update(receivedMessage.getVectorClock());
-		this.localVariable = receivedMessage.getLocalVariable() - this.localVariable;
-		this.vectorClock.increment();
+			// line 7
+			// receive
+			receivedMessage = receive(2); // receive from process 2
+			this.vectorClock.update(receivedMessage.getVectorClock());
+			this.localVariable = receivedMessage.getLocalVariable() - this.localVariable;
+			this.vectorClock.increment();
 
-		// notify the monitor
-		message = new Message(new VectorClock(vectorClock), this.localVariable);
-		monitor.receiveMessage(this.Id, message);
+			// notify the monitor
+			message = new Message(new VectorClock(vectorClock), this.localVariable);
+			monitor.receiveMessage(this.Id, message);
+		}
 
 		// send terminate signal
 		monitor.processTerminated(this.Id);

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Process3.java

@@ -13,7 +13,6 @@ public class Process3 extends AbstractProcess {
 	@Override
 	public void run() {
 
-		//TODO Implement processes3 (Listing 4) code here!
 		// send the initial state to Monitor
 		Message message = new Message(new VectorClock(vectorClock), this.localVariable);
 		monitor.receiveMessage(this.Id, message);
@@ -59,6 +58,8 @@ public class Process3 extends AbstractProcess {
 		// line 5: send ( P1 , x3 )
 		send(0, message);
 
+		// send terminate signal
+		monitor.processTerminated(this.Id);
 		System.out.printf("process:%d , the local variable= %d\n", this.Id, this.localVariable);
 	}
 }

### 2019-12-02 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Monitor.java
continued 6.1.b) but not done

@@ -1,244 +1,316 @@
 package de.unistgt.ipvs.vs.ex4.distributed_debugging_algorithm;
 
-import java.util.ArrayList;
-import java.util.Iterator;
-import java.util.LinkedList;
-import java.util.List;
+import java.util.*;
 import java.util.concurrent.atomic.AtomicInteger;
 
 //you are not allowed to change this class structure. However, you can add local functions!
 public class Monitor implements Runnable {
 
-	/**
-	 * The state consists on vector timestamp and local variables of each
-	 * process. In this class, a state is represented by messages (events)
-	 * indices of each process. The message contains a local variable and vector
-	 * timestamp, see Message class. E.g. if state.processesMessagesCurrentIndex
-	 * contains {1, 2}, it means that the state contains the second message
-	 * (event) from process1 and the third message (event) from process2
-	 */
-	private class State {
-		// Message indices of each process
-		private int[] processesMessagesCurrentIndex;
+    /**
+     * The state consists on vector timestamp and local variables of each
+     * process. In this class, a state is represented by messages (events)
+     * indices of each process. The message contains a local variable and vector
+     * timestamp, see Message class. E.g. if state.processesMessagesCurrentIndex
+     * contains {1, 2}, it means that the state contains the second message
+     * (event) from process1 and the third message (event) from process2
+     */
+    private class State {
+        // Message indices of each process
+        private int[] processesMessagesCurrentIndex;
 
-		public State(int numberOfProcesses) {
-			processesMessagesCurrentIndex = new int[numberOfProcesses];
-		}
+        public State(int numberOfProcesses) {
+            processesMessagesCurrentIndex = new int[numberOfProcesses];
+        }
 
-		public State(int[] processesMessagesCurrentIndex) {
-			this.processesMessagesCurrentIndex = processesMessagesCurrentIndex;
-		}
+        public State(int[] processesMessagesCurrentIndex) {
+            this.processesMessagesCurrentIndex = processesMessagesCurrentIndex;
+        }
 
-		{
-			processesMessagesCurrentIndex = new int[numberOfProcesses];
-		}
+        {
+            processesMessagesCurrentIndex = new int[numberOfProcesses];
+        }
 
-		public int[] getProcessesMessagesCurrentIndex() {
-			return processesMessagesCurrentIndex;
-		}
+        public int[] getProcessesMessagesCurrentIndex() {
[…]

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Message.java
Add debugger displays to existing classes

@@ -17,4 +17,10 @@ public class Message {
 	public VectorClock getVectorClock() {
 		return vectorClock;
 	}
+
+	@Override
+	public String toString() {
+		// Debugger display
+		return localVariable + "@" + vectorClock.toString();
+	}
 }

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Monitor.java

@@ -49,6 +49,12 @@ public class Monitor implements Runnable {
 
             return true;
         }
+
+        @Override
+        public String toString() {
+            // debugger display
+            return "[" + String.join(", ", Arrays.stream(processesMessagesCurrentIndex).mapToObj(String::valueOf).toArray(String[]::new)) + "]";
+        }
     }
 
     private int numberOfProcesses;

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/VectorClock.java

@@ -48,6 +48,14 @@ public class VectorClock {
 		 * Implement a code to check if a state is consist regarding two vector clocks (i.e. this and other).
 		 * See slide 41 from global state lecture.
 		 */
-		return vectorClock[otherProcessId] == other.get()[otherProcessId];
+               return vectorClock[otherProcessId] == other.get()[otherProcessId];
+	}
+
+	@Override
+	public String toString() {
+		// debugger display
+		String[] parts = Arrays.stream(vectorClock).mapToObj(String::valueOf).toArray(String[]::new);
+		parts[processId] = "(" + parts[processId] + ")";
+		return "[" + String.join(", ", parts) + "]";
 	}
 }

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Monitor.java
Simplify initializations and make formatting consistent

@@ -16,18 +16,15 @@ public class Monitor implements Runnable {
      */
     private class State {
         // Message indices of each process
-        private int[] processesMessagesCurrentIndex;
+        private final int[] processesMessagesCurrentIndex;
 
         public State(int numberOfProcesses) {
-            processesMessagesCurrentIndex = new int[numberOfProcesses];
+            // new array automatically initializes clocks to zero
+            this(new int[numberOfProcesses]);
         }
 
         public State(int[] processesMessagesCurrentIndex) {
             this.processesMessagesCurrentIndex = processesMessagesCurrentIndex;
-        }
-
-        {
-            processesMessagesCurrentIndex = new int[numberOfProcesses];
         }
 
         public int[] getProcessesMessagesCurrentIndex() {
@@ -35,7 +32,7 @@ public class Monitor implements Runnable {
         }
 
         public int getProcessMessageCurrentIndex(int processId) {
-            return this.processesMessagesCurrentIndex[processId];
+            return processesMessagesCurrentIndex[processId];
         }
 
         @Override
@@ -43,10 +40,11 @@ public class Monitor implements Runnable {
             State otherState = (State) other;
 
             // Iterate over processesMessagesCurrentIndex array
-            for (int i = 0; i < numberOfProcesses; i++)
-                if (this.processesMessagesCurrentIndex[i] != otherState.processesMessagesCurrentIndex[i])
+            for (int i = 0; i < numberOfProcesses; i++) {
+                if (this.processesMessagesCurrentIndex[i] != otherState.processesMessagesCurrentIndex[i]) {
                     return false;
-
+                }
+            }
             return true;
         }
 
@@ -90,16 +88,11 @@ public class Monitor implements Runnable {
 
         states = new LinkedList<>();
 
-        possiblyTruePredicatesIndex = new boolean[numberOfPredicates];// there
-        // are
-        // three
-        // predicates
-        for (int i = 0; i < numberOfPredicates; i++)
-            possiblyTruePredicatesIndex[i] = false;
-
+        possiblyTruePredicatesIndex = new boolean[numberOfPredicates];
+        Arrays.fill(possiblyTruePredicatesIndex, false);
+        
         definitelyTruePredicatesIndex = new boolean[numberOfPredicates];
-        for (int i = 0; i < numberOfPredicates; i++)
-            definitelyTruePredicatesIndex[i] = false;
[…]

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Process2.java

@@ -23,7 +23,6 @@ public class Process2 extends AbstractProcess {
 		this.vectorClock.update(receivedMessage.getVectorClock());
 		this.localVariable = receivedMessage.getLocalVariable();
 		this.vectorClock.increment();
-
 		// notify the monitor
 		message = new Message(new VectorClock(vectorClock), this.localVariable);
 		monitor.receiveMessage(this.Id, message);

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/VectorClock.java

@@ -1,23 +1,21 @@
 package de.unistgt.ipvs.vs.ex4.distributed_debugging_algorithm;
 
+import java.util.Arrays;
+
 //you are not allowed to change this class structure
 public class VectorClock {
 
 	protected int[] vectorClock;
 	private int processId;
-	private int numberOfProcesses;
 
 	public VectorClock(int processId, int numberOfProcesses) {
 		vectorClock = new int[numberOfProcesses];
-		this.numberOfProcesses = numberOfProcesses;
 		this.processId = processId;
 	}
 
 	VectorClock(VectorClock other) {
 		vectorClock = other.vectorClock.clone();
 		processId = other.processId;
-		numberOfProcesses = other.numberOfProcesses;
-
 	}
 
 	public void increment() {

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/test/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_test/DebugTest.java

@@ -15,125 +15,125 @@ import org.junit.rules.Timeout;
 
 public class DebugTest {
 
-	@Rule
-        public Timeout globalTimeout = Timeout.seconds(30);
-	
-        @Test
-	public void testPartb() throws InterruptedException {
+    @Rule
+    public Timeout globalTimeout = Timeout.seconds(30);
+    
+    @Test
+    public void testPartb() throws InterruptedException {
 
-		System.out.println("Check only part (b) from the question (4)");
+        System.out.println("Check only part (b) from the question (4)");
 
-		// instantiation and run threads
-		int numberOfProcesses = 2;
-		Monitor monitor = new Monitor(numberOfProcesses);
-		AbstractProcess[] processes = new AbstractProcess[numberOfProcesses];
+        // instantiation and run threads
+        int numberOfProcesses = 2;
+        Monitor monitor = new Monitor(numberOfProcesses);
+        AbstractProcess[] processes = new AbstractProcess[numberOfProcesses];
 
-		Process1 process1 = new Process1(monitor, processes, 0);
-		processes[0] = process1;
+        Process1 process1 = new Process1(monitor, processes, 0);
+        processes[0] = process1;
 
-		Process2 process2 = new Process2(monitor, processes, 1);
-		processes[1] = process2;
+        Process2 process2 = new Process2(monitor, processes, 1);
+        processes[1] = process2;
 
-		Thread process1Thread = new Thread(process1);
-		Thread process2Thread = new Thread(process2);
-		Thread monitorThread = new Thread(monitor);
+        Thread process1Thread = new Thread(process1);
+        Thread process2Thread = new Thread(process2);
+        Thread monitorThread = new Thread(monitor);
 
-		process1Thread.start();
-		process2Thread.start();
+        process1Thread.start();
+        process2Thread.start();
 
-		monitorThread.start();
+        monitorThread.start();
 
-		process1Thread.join();
-		process2Thread.join();
+        process1Thread.join();
+        process2Thread.join();
 
-		monitorThread.join();
+        monitorThread.join();
 
-		// checking ---------------
+        // checking ---------------
 
-		// check local variables
-		assertEquals(10,processes[0].getLocalVariable());
-		assertEquals(25,processes[1].getLocalVariable());
+        // check local variables
+        assertEquals(10, processes[0].getLocalVariable());
[…]

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/Monitor.java
Implement linearization flattenening and Vector Clock consistency
co-authored-by: DerEinherjer <steffen.wonner@web.de>

@@ -159,7 +159,6 @@ public class Monitor implements Runnable {
     }
 
     public void buildLattice(int predicateNo, int process_i_id, int process_j_id) {
-        // TODO check if works
         /*
          * - implement this function to build the lattice of consistent states.
          * - The goal of building the lattice is to check a predicate if it is
@@ -270,32 +269,50 @@ public class Monitor implements Runnable {
          * Predicate.predicate0(process_i_Message, process_j_Message); break;
          * case 1: ... }
          */
-		// Check if possibly true
-		for (State s : states) {
-			// get messages
-			int index_i = s.getProcessMessageCurrentIndex(process_i_id);
-			int index_j = s.getProcessMessageCurrentIndex(process_j_id);
-			Message msg_i = processesMessages.get(process_i_id).get(index_i);
-			Message msg_j = processesMessages.get(process_j_id).get(index_j);
-			// selects predicate
-			if (getPredicateNo(predicateNo, msg_i, msg_j)) {
-				possiblyTruePredicatesIndex[predicateNo] = true;
-				break;
-			}
-		}
-
-		// TODO
-		// Check if definitely true
-		boolean predicate = false;
-		// states in one level
-		LinkedList<State> states_l = new LinkedList<State>();
-		states_l.add(states.getFirst());
-
-		/*while (!states_l.isEmpty()) {
-			getPredicateNo(predicateNo, msg_i, msg_j);
-		}*/
-
+        boolean predicate = true;
+        List<List<State>> linearizations = determineLinearizations(states.getFirst());
+        for (Iterable<State> linearization : linearizations) {
+            boolean ok = false;
+            for (State state : linearization) {            // get messages
+                int index_i = state.getProcessMessageCurrentIndex(process_i_id);
+                int index_j = state.getProcessMessageCurrentIndex(process_j_id);
+                Message msg_i = processesMessages.get(process_i_id).get(index_i);
+                Message msg_j = processesMessages.get(process_j_id).get(index_j);
+                if (getPredicateNo(predicateNo, msg_i, msg_j)) {
+                    // found at least one linearization that fulfils the predicate,
+                    // therefore the predicate is possibly true
+                    possiblyTruePredicatesIndex[predicateNo] = true;
+                    ok = true;
+                    break;
+                }
+            }
+            if (!ok) {
+                predicate = false;
+                break;
+            }
+        }
         return predicate;
+    }
+
+    /**
+     * Recursively determines all possible linearizations when starting from the given state.
[…]

### 2019-12-03 assignment4/distributed_debugging_algorithm-task/src/main/java/de/unistgt/ipvs/vs/ex4/distributed_debugging_algorithm/VectorClock.java

@@ -32,21 +32,23 @@ public class VectorClock {
 
 	public void update(VectorClock other) {
 		/*
-		 * Implement Supermum operation
+		 * Implement Supremum operation
 		 */
-		this.increment();
+		increment();
 		for (int i = 0; i < vectorClock.length; i++) {
-			vectorClock[i] = Math.max(vectorClock[i], other.get()[i]);
+			vectorClock[i] = Math.max(vectorClock[i], other.vectorClock[i]);
 		}
 	}
 
 	public boolean checkConsistency(int otherProcessId, VectorClock other) {
-		//TODO
 		/*
 		 * Implement a code to check if a state is consist regarding two vector clocks (i.e. this and other).
 		 * See slide 41 from global state lecture.
 		 */
-               return vectorClock[otherProcessId] == other.get()[otherProcessId];
+		// check that this clock does not count an event that the other clock hasn't seen yet in it's local process
+		return vectorClock[otherProcessId] <= other.vectorClock[otherProcessId]
+		// and check that the other clock has not seen an event for this process that hasn't happened yet
+			&& vectorClock[processId] >= other.vectorClock[processId];
 	}
 
 	@Override

### 2019-12-09 assignment4/submission4.tex
add ex4-theoretical files

--- 2019-12-01 assignment3/submission3.tex
@@
 \newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
-\newcommand{\getsubtitle}{G2T1 – Assignment 3 (theoretical part)}
+\newcommand{\getsubtitle}{G2T1 – Assignment 4 (theoretical part)}
 \newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
@@
 
-\section*{1 - Physical Clocks}
+\section*{1 - Global State}
 \subsection*{a)}
+\subsection*{b)}
+\subsection*{c)}
+\subsection*{d)}
 
-\begin{figure}[ht!]
-\begin{tikzpicture}[scale=0.2]
-\foreach \x in {1,...,81} {
-	\draw (\x,0) -- (\x, 2);
-	\draw (\x, -5) -- (\x, -7);
-};
-\draw (0,0) -- (81,0);
-\draw (0,-5) -- (81,-5);
-\node at(0, -1) {\tiny Master};
-\node at(0, -4) {\tiny Slave};
-\foreach \x/\n in {3/7,8/8,13/9,18/10,23/11,28/12,33/13,38/14,43/15,48/16,53/17,58/18,63/19,68/20,73/21,78/22} {
-	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,1) (a\n) {}
-		node[below=7mm, above of=a\n] {\tiny \n};
-};
-\foreach \x/\n in {3/3,8/4,13/5,18/6,23/7,28/8,31/9,34/10,37/11,40/12,43/13,46/14,49/15,52/16,55/17,58/18,63/19,68/20,73/21,78/22} {
-	\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,-6) (b\n) {}
-		node[above=7mm, below of=b\n] {\tiny \n};
-};
-\path[draw,->] 
-	(b5.north) edge (a10.south)
-	(a11.south) edge (b8.north);
-\end{tikzpicture}
-\caption{Solution for Figure 1}
-\end{figure}
-\begin{align*}
-O = \frac12((t_2 - t_1) - (t_4 - t_3)) = \frac12((10 - 5) - (8 - 11)) = \frac12(5 + 3) = 4 &\\
-d = 1 & \\
-\Delta = 6, C_s(t_4 + \Delta) = 18 & \\
-\frac{dC_S}{dt} = \frac{6 + 4}{6} = \frac{5}{3} &
-\end{align*}
-
-\begin{figure}[ht!]
-	\begin{tikzpicture}[scale=0.2]
-	\foreach \x in {1,...,81} {
-		\draw (\x,0) -- (\x, 2);
-		\draw (\x, -5) -- (\x, -7);
-	};
-	\draw (0,0) -- (81,0);
-	\draw (0,-5) -- (81,-5);
-	\node at(0, -1) {\tiny Master};
-	\node at(0, -4) {\tiny Slave};
-	\foreach \x/\n in {3/1,8/2,13/3,18/4,23/5,28/6,33/7,38/8,43/9,48/10,53/11,58/12,63/13,68/14,73/15,78/16} {
-		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,1) (a\n) {}
-			node[below=7mm, above of=a\n] {\tiny \n};
-	};
-	\foreach \x/\n in {3/6,8/7,13/8,18/9,23/10,28/11,41/12,55/13,68/14,73/15,78/16} {
-		\draw node[rectangle, fill=black, text width=1mm, inner sep=0, minimum height=4mm] at(\x,-6) (b\n) {}
-			node[above=7mm, below of=b\n] {\tiny \n};
-	};
-	\path[draw,->]
-		(b8.north) edge (a4.south)
-		(a5.south) edge (b11.north);
-\end{tikzpicture}
-\caption{Solution for Figure 2}
-\end{figure}
-\begin{align*}
-O = \frac12((t_2 - t_1) - (t_4 - t_3)) = \frac12((4 - 8) - (11 - 5)) = \frac12(-4 - 6) = -5  &\\
-d = 1 & \\
-\Delta = 3, C_s(t_4 + \Delta) = 14 & \\
-\frac{dC_S}{dt} = \frac{3}{3 +5} = \frac{3}{8} &
-\end{align*}
-
+\section*{2 - Transaction Processing}
+\subsection*{a)}
 \subsection*{b)}
 
-\section*{2 - Logical Clocks}
-\subsection*{a)}
-\begin{enumerate}[label=(\roman*)]
-	\item 
-	C(e$^1_1$) = 2\\
-	C(e$^2_1$) = 3\\
-	C(e$^3_1$) = 4\\
-	C(e$^4_1$) = 5\\
-
-	C(e$^1_2$) = 5\\
-	C(e$^2_2$) = 6\\
-	C(e$^3_2$) = 7\\
-	C(e$^4_2$) = 8\\
-
-	C(e$^1_3$) = 1\\
-	C(e$^2_3$) = 2\\
-	C(e$^3_3$) = 3\\
-	C(e$^4_3$) = 9\\
-	\item 
-	VC(e$^1_1$) = \texttt{[1 0 1]}\\
-	VC(e$^2_1$) = \texttt{[2 0 1]}\\
-	VC(e$^3_1$) = \texttt{[3 0 1]}\\
-	VC(e$^4_1$) = \texttt{[4 0 1]}\\
-
-	VC(e$^1_2$) = \texttt{[3 1 1]}\\
-	VC(e$^2_2$) = \texttt{[3 2 2]}\\
-	VC(e$^3_2$) = \texttt{[3 3 2]}\\
-	VC(e$^4_2$) = \texttt{[3 4 2]}\\
-
-	VC(e$^1_3$) = \texttt{[0 0 1]}\\
-	VC(e$^2_3$) = \texttt{[0 0 2]}\\
-	VC(e$^3_3$) = \texttt{[0 0 3]}\\
-	VC(e$^4_3$) = \texttt{[3 4 4]}\\
-	\item
-	P$_1$: e$^1_1$ e$^2_1$ e$^3_1$\\
-	P$_2$: e$^1_2$ e$^2_2$\\
-	P$_3$: e$^1_3$ e$^2_3$\\
-	The vector event time of those events is smaller than the vector event time of e$^3_2$. A smaller vector time means events are causally related.
-\end{enumerate}
-\subsection*{b)}
-\begin{enumerate}[label=(\roman*)]
-	\item 
-	P$_1$: e$_6$ e$_5$ e$_9$ e$_4$ e$_8$ e$_{10}$\\
-	P$_2$: e$_{12}$ e$_3$ e$_{13}$\\
-	P$_3$: e$_{11}$ e$_7$ e$_1$ e$_2$\\
-	\item 
-	Local event: e$_5$ e$_9$ e$_3$ e$_1$ e$_{11}$\\
-	Send event: e$_6$ e$_4$ e$_{10}$ e$_{12}$\\
-	Recieve event:  e$_8$ e$_{13}$ e$_7$ e$_2$\\
-\end{enumerate}
-
-\section*{3 - Global State}
-\subsection*{a)}
-$ \{(e^{1}_{1}, e^{1}_{2}), (e^{2}_{1}, e^{2}_{2}), (e^{2}_{2}, e^{3}_{1}), (e^{4}_{1}, e^{3}_{2}), (e^{4}_{1}, e^{4}_{2}), (e^{5}_{1}, e^{5}_{2}), (e^{1}_{1}, e^{2}_{2}), (e^{2}_{2}, e^{4}_{1}), (e^{4}_{1}, e^{5}_{2})\} $
-
-\subsection*{b)}
-\begin{enumerate}
-	\item yes, as it is a valid linearization.
-	\item no, because $ e_{2}^{3} $ can not happen before $ e_{1}^{3} $
-\end{enumerate}
-
-\subsection*{c)}
-\begin{figure}[!ht]
-	\centering
-	\begin{tikzpicture}[auto,node distance=1.5cm]
-		\node[draw] (s00) {$ S_{00} $};
-		\node[draw] (s10) [below left = of s00] {$ S_{10} $};
-		\node[draw] (s01) [below right = of s00] {$ S_{01} $};
-		\node[draw] (s11) [below right = of s10] {$ S_{11} $};
-		\node[draw] (s02) [below right = of s01] {$ S_{02} $};
-
-		\node[draw] (s21) [below left = of s11] {$ S_{21} $};
-		\node[draw] (s12) [below left = of s02] {$ S_{12} $};
-
-		\node[draw] (s31) [below left = of s21] {$ S_{31} $};
-		\node[draw] (s22) [below left = of s12] {$ S_{22} $};
-
-		\node[draw] (s41) [below left = of s31] {$ S_{41} $};
-		\node[draw] (s32) [below left = of s22] {$ S_{32} $};
-
-		\node[draw] (s42) [below right = of s41] {$ S_{42} $};
-		\node[draw] (s33) [below right = of s32] {$ S_{33} $};
-
-		\node[draw] (s43) [below right = of s42] {$ S_{43} $};
-		\node[draw] (s34) [below right = of s33] {$ S_{34} $};
-
-		\node[draw] (s44) [below right = of s43] {$ S_{44} $};
-		\node[draw] (s35) [below right = of s34] {$ S_{35} $};
-
-		\node[draw] (s54) [below left = of s44] {$ S_{54} $};
-		\node[draw] (s45) [below left = of s35] {$ S_{45} $};
-
-		\node[draw] (s55) [below right = of s54] {$ S_{55} $};
-
-		\path (s10) edge node {$ e^{1}_{1} $} (s00);
-		\path (s11) edge node {$ e^{1}_{1} $} (s01);
-		\path (s12) edge node {$ e^{1}_{1} $} (s02);
-
-		\path (s00) edge node {$ e^{1}_{2} $} (s01);
-		\path (s10) edge node {$ e^{1}_{2} $} (s11);
-
-		\path (s21) edge node {$ e^{2}_{1} $} (s11);
-		\path (s22) edge node {$ e^{2}_{1} $} (s12);
-
-		\path (s41) edge node {$ e^{2}_{2} $} (s42);
-		\path (s31) edge node {$ e^{2}_{2} $} (s32);
-		\path (s21) edge node {$ e^{2}_{2} $} (s22);
-		\path (s11) edge node {$ e^{2}_{2} $} (s12);
-		\path (s01) edge node {$ e^{2}_{2} $} (s02);
-
-		\path (s31) edge node {$ e^{3}_{1} $} (s21);
-		\path (s32) edge node {$ e^{3}_{1} $} (s22);
-
-		\path (s32) edge node {$ e^{3}_{2} $} (s33);
-		\path (s42) edge node {$ e^{3}_{2} $} (s43);
-
-		\path (s41) edge node {$ e^{4}_{1} $} (s31);
-		\path (s42) edge node {$ e^{4}_{1} $} (s32);
-		\path (s43) edge node {$ e^{4}_{1} $} (s33);
-		\path (s44) edge node {$ e^{4}_{1} $} (s34);
-		\path (s45) edge node {$ e^{4}_{1} $} (s35);
-
-		\path (s33) edge node {$ e^{4}_{2} $} (s34);
-		\path (s43) edge node {$ e^{4}_{2} $} (s44);
-
-		\path (s34) edge node {$ e^{5}_{2} $} (s35);
-		\path (s44) edge node {$ e^{5}_{2} $} (s45);
-		\path (s54) edge node {$ e^{5}_{2} $} (s55);
-
-		\path (s54) edge node {$ e^{5}_{1} $} (s44);
-		\path (s55) edge node {$ e^{5}_{1} $} (s45);
-	\end{tikzpicture}
-	\caption{solution for 3.c)}
-\end{figure}
-
-\newpage
-\section*{4 - Snapshot Algorithm}
-\subsection*{a)}
-
-See Figure 4
-
-\begin{figure}[!ht]
-    \centering
-    \begin{tikzpicture}[auto,node distance=1.5cm,event/.style={fill=black,circle, inner sep=0pt, minimum size=5pt},state/.style={draw,rectangle, red, color=red, inner sep=0pt, minimum size=5pt, minimum width=10pt}, msg/.style={-latex,thick}, marker/.style={-latex,thick,red}]
-        \node[] (p1s) {\(P_1\)};
-        \node[] (p2s) [below = of p1s] {\(P_2\)};
-        \node[] (p3s) [below = of p2s] {\(P_3\)};
-
-        \node[] (p1e) [right = 13cm of p1s] {};
-        \node[] (p2e) [right = 13cm of p2s] {};
-        \node[] (p3e) [right = 13cm of p3s] {};
-
-        \path (p1s.east) edge (p1e);
-        \path (p2s.east) edge (p2e);
-        \path (p3s.east) edge (p3e);
-
-        \node[event,label={\(e_1^1\)}] (e11) [right = 1.3cm of p1s] {};
-        \node[event,label={\(e_2^1\)}] (e21) [right = 1cm of p2s] {};
-        \node[event,label={\(e_3^1\)}] (e31) [right = 2cm of p3s] {};
-
-        \node[event,label={\(e_1^2\)}] (e12) [right = 4.2cm of p1s] {};
-        \node[event,label={\(e_2^2\)}] (e22) [right = 2.1cm of p2s] {};
-        \node[event,label={\(e_3^2\)}] (e32) [right = 4cm of p3s] {};
-
-        \node[event,label={\(e_1^3\)}] (e13) [right = 6cm of p1s] {};
-        \node[event,label={\(e_2^3\)}] (e23) [right = 4.5cm of p2s] {};
-        \node[event,label={\(e_3^3\)}] (e33) [right = 10cm of p3s] {};
-
-        \node[event,label={\(e_1^4\)}] (e14) [right = 7cm of p1s] {};
-        \node[event,label=below:{\(e_2^4\)}] (e24) [right = 8.5cm of p2s] {};
-
-        \node[event,label={\(e_1^5\)}] (e15) [right = 10.5cm of p1s] {};
-        \node[event,label={\(e_2^5\)}] (e25) [right = 9cm of p2s] {};
-
-        \node[event,label={\(e_2^6\)}] (e26) [right = 12.5cm of p2s] {};
-
-        \path (e21) edge[msg] node {\(m_1\)} (e31);
-        \path (e11) edge[msg] node[left] {\(m_2\)} (e23);
-        \path (e22) edge[msg] node[right] {\(m_3\)} (e12);
-        \path (e32) edge[msg] node[below right] {\(m_4\)} (e14);
-        \path (e13) edge[msg] node[right] {\(m_5\)} (e25);
-        \path (e15) edge[msg] node {\(m_6\)} (e26);
-        \path (e24) edge[msg] node {\(m_7\)} (e33);
-
-        \node[state,label={[red]:{\(s_3\)}}] (s3) [right = 3cm of p3s] {};
-        \node[state,label={[red]:{\(s_1,c_4\)}}] (s1) [right = 5cm of p1s] {};
-        \node[state,label={[red]below:{\(s_2,c_1\)}}] (s2) [right = 6.5cm of p2s] {};
-
-        \node[state,label={[red]:{\(c_2\)}}] (c2) [right = 8cm of p1s] {};
-        \node[state,label={[red]:{\(c_3\)}}] (c3) [right = 8.5cm of p3s] {};
-
-        \path (s3.north east) edge[marker] node[below] {\(M\)} (s1.south west);
-        \path (s1.south east) edge[marker] node[left] {\(M'\)} (s2.north west);
-        \path (s2.north east) edge[marker] node[below left] {\(M''\)} (c2.south west);
-        \path (s2.south east) edge[marker] node[] {\(M''\)} (c3.north west);
-    \end{tikzpicture}
-
-    The red boxes represent where a process saves some state.
-A process state save is denoted by \(s_i\) and a channel state save by \(c_j\) where \(i\) is the process number and \(j\) is the channel number.
-    \caption{solution for 4.a)}
-\end{figure}
-
-\subsection*{b)}
-
-No channel received a message between the process starting recording and the first/second marker arriving. All channel states are empty sets.
-
-\begin{align*}
-    c_1 &= \{\}\\
-    c_2 &= \{\}\\
-    c_3 &= \{\}\\
-    c_4 &= \{\}\\
-\end{align*}
+\section*{3 - Two-Phase Locking}
 

### 2019-12-10 assignment4/submission4.tex
Add solution for exercise 4.2 a) and b)

@@
 \subsection*{a)}
+The conflicting operation pairs are: 
+\[\{(w_1[u], w_3[u]), (r_1[y], w_2[y]), (r_3[y], w_2[y]), (r_2[x], w_3[x])\}\]
+
 \subsection*{b)}
+
+\begin{figure}[ht!]
+\begin{tikzpicture}
+    \node[draw, circle] at (0, 0)    (T1) {\(T_1\)};
+    \node[draw, circle] at (5, 0)    (T2) {\(T_2\)};
+    \node[draw, circle] at (2.5,2.5) (T3) {\(T_3\)};
+
+    \draw[->]
+        (T1) edge node[above,sloped] {\(w_1[y] \to r_2[y]\)} (T2)
+        (T2) edge node[above,sloped] {\(w_2[z] \to w_3[z]\)} (T3)
+        (T3) edge node[above,sloped] {\(w_3[y] \to r_1[y]\)} (T1);
+\end{tikzpicture}
+\caption{Serialization Graph for History \(H_1\)}
+\label{fig:sg_h1}
+\end{figure}
+
+It's clear that the serialization graph shown in Figure \ref{fig:sg_h1} contains a cycle, as such \(H_1\) is not serializable.\\~\\
+
+\begin{figure}[ht!]
+\begin{tikzpicture}
+    \node[draw, circle] at (0, 0)    (T1) {\(T_1\)};
+    \node[draw, circle] at (5, 0)    (T2) {\(T_2\)};
+    \node[draw, circle] at (2.5,2.5) (T3) {\(T_3\)};
+
+    \draw[->]
+        (T2) edge node[above,sloped] {\(r_2[y] \to w_1[y]\)} (T1)
+        (T3) edge node[above,sloped] {\(r_3[x] \to w_1[x]\)} (T1)
+        (T3) edge node[above,sloped] {\parbox[c]{2.5cm}{\(r_3[x] \to w_2[x] \\ w_3[u] \to w_2[u]\)}} (T2);
+\end{tikzpicture}
+\caption{Serialization Graph for History \(H_2\)}
+\label{fig:sg_h2}
+\end{figure}
+
+The serialization graph shown in Figure \ref{fig:sg_h2} does not contain a cycle, which would imply that \(H_2\) is serializable.\\
+Unfortunately \(H_2\) is not a correct history as it does not specify the order or the conflicting operations \(w_1[x], w_2[x]\).
 

### 2019-12-10 assignment4/submission4.tex
done 4.1 c) and d)

--- 2019-12-09 assignment4/submission4.tex
@@
 \subsection*{c)}
+
+\tikzstyle{nofill}=[rectangle,draw, align=center]
+\tikzstyle{green}=[rectangle,draw,fill=green!50,align=center]
+\tikzstyle{blue}=[rectangle,draw,fill=blue!50,align=center]
+\tikzstyle{both}=[rectangle,draw,fill=red!50!,align=center]
+
+\begin{figure}[!ht]
+\centering
+\begin{tikzpicture}[auto,node distance=1.5cm]
+\node[green] (s00) {$ S_{00} $ \\ $ x_1 = 0 $ \\ $ x_2 = 0 $};
+
+\node[both] (s10) [below left = of s00] {$ S_{10} $ \\ $ x_1 = 10 $ \\ $ x_2 = 0 $};
+
+\node[both] (s20) [below left = of s10] {$ S_{20} $ \\ $ x_1 = 5 $ \\ $ x_2 = 0 $};
+\node[green] (s11) [below right = of s10] {$ S_{11} $ \\ $ x_1 = 10 $ \\ $ x_2 = 20 $};
+
+\node[both] (s30) [below left = of s20] {$ S_{30} $ \\ $ x_1 = 15 $ \\ $ x_2 = 0 $};
+\node[green] (s21) [below left = of s11] {$ S_{21} $ \\ $ x_1 = 5 $ \\ $ x_2 = 20 $};
+\node[green] (s12) [below right = of s11] {$ S_{12} $ \\ $ x_1 = 10 $ \\ $ x_2 = 30 $};
+
+\node[nofill] (s31) [below left = of s21] {$ S_{31} $ \\ $ x_1 = 15 $ \\ $ x_2 = 20 $};
+\node[green] (s22) [below left = of s12] {$ S_{22} $ \\ $ x_1 = 5 $ \\ $ x_2 = 30 $};
+
+\node[green] (s32) [below left = of s22] {$ S_{32} $ \\ $ x_1 = 15 $ \\ $ x_2 = 30 $};
+
+\node[both] (s42) [below left = of s32] {$ S_{42} $ \\ $ x_1 = 55 $ \\ $ x_2 = 30 $};
+
+\node[both] (s52) [below left = of s42] {$ S_{52} $ \\ $ x_1 = 60 $ \\ $ x_2 = 30 $};
+
+\node[both] (s53) [below right = of s52] {$ S_{53} $ \\ $ x_1 = 60 $ \\ $ x_2 = 13 $};
+
+\path (s10) edge node {} (s00);
+
+\path (s10) edge node {} (s11);
+\path (s10) edge node {} (s20);
+
+\path (s11) edge node {} (s21);
+\path (s20) edge node {} (s21);
+\path (s20) edge node {} (s30);
+\path (s11) edge node {} (s12);
+
+\path (s30) edge node {} (s31);
+\path (s21) edge node {} (s31);
+\path (s21) edge node {} (s22);
+\path (s12) edge node {} (s22);
+
+\path (s31) edge node {} (s32);
+\path (s22) edge node {} (s32);
+
+\path (s32) edge node {} (s42);
+
+\path (s42) edge node {} (s52);
+
+\path (s52) edge node {} (s53);
+\end{tikzpicture}
+\caption{solution for 1.c)}
+\end{figure}
+
 \subsection*{d)}
+\begin{itemize}
+	\item $ \phi_{1} = (x_{2} - x_{1}) = 5 $\\
+	$ \Rightarrow (\neg\phi_{1}) $ marked in \textcolor{blue}{blue}
+	\item $ \phi_{2} = (x_{2} - x_{1}) \leq 5 $ and $ (x_{2} - x_{1}) > 0 $\\
+	$ \Rightarrow (\phi_{2}) $ marked in \textcolor{green}{green}
+	\item if they fulfill $ (\neg\phi_{1}) $ and $ (\phi_{2}) $ marked in \textcolor{red}{red}
+\end{itemize}
 
+\begin{enumerate}
+	\item safety condition $ \neg\phi_{1} $ is not fulfilled in the state $ S_{31} $and thereby not fulfilled.
+	\item liveness condition $ \phi_{2} $ is fulfilled in the end and thereby fulfilled.
+\end{enumerate}
 \section*{2 - Transaction Processing}

### 2019-12-11 assignment4/submission4.tex
fix typo
Added 1 a and b

@@
 \subsection*{a)}
+$C_{1}$: $e_{11}$ $e_{21}$ $e_{31}$ $e_{12}$ $e_{22}$\\
+$C_{2}$: $e_{11}$ $e_{21}$ $e_{31}$ $e_{12}$ $e_{22}$ $e_{41}$
+
 \subsection*{b)}
+\begin{enumerate}[label=(\roman*)]
+\item
+The inequation give true for all states that can follow the first one given independant on how many states lie between them.
+To make sure that the other state lies on level L+1 you can either check if any other state lies between them or check the timestamps.
+The stamp of the prosess with tha last event has to be 1 higher than in the other state.
+\item
+$S_{30}$ -> $S_{31}$ = true
+$S_{30}$ -> $S_{32}$ = true
+$S_{30}$ -> $S_{42}$ = true
+
+$S_{31}$ -> $S_{32}$ = true
+$S_{32}$ -> $S_{42}$ = true
+So $S_{32}$ and $S_{42}$ are not on level L+1, because $S_{31}$ lies between them and $S_{30}$.\\
+You either have to check the timestamps (not given) or check all other states if they lie between $S_{30}$ and $S_{31}$. Non of them do so $S_{31}$ is on level l+1 of $S_{30}$.
+\end{enumerate}
 \subsection*{c)}
@@
 \begin{enumerate}
-	\item safety condition $ \neg\phi_{1} $ is not fulfilled in the state $ S_{31} $and thereby not fulfilled.
+	\item safety condition $ \neg\phi_{1} $ is not fulfilled in the state $ S_{31} $ and thereby not fulfilled.
 	\item liveness condition $ \phi_{2} $ is fulfilled in the end and thereby fulfilled.

### 2019-12-12 assignment4/submission4.tex
Added exercise 4.3

@@
 
+\begin{itemize}
+    \item[\(H_1\):] could be generated by 2-phase-locking. The shrinking phase for each transaction begins right after the last operation before the commit.
+
+        Execution split for each resource:
+
+        \begin{itemize}
+            \item[\(y\):] \(rL_1 \rightarrow r_1 \rightarrow wL_1 \rightarrow w_1 \rightarrow wU_1 \rightarrow rL_2 \rightarrow r_2 \rightarrow c_1 \rightarrow rU_2 \rightarrow c_2 \rightarrow c_3\)
+            \item[\(x\):] only read operations $\Rightarrow$ no conflicting locks possible
+            \item[\(z\):] only read operations $\Rightarrow$ no conflicting locks possible
+            \item[\(w\):] \(c_1 \rightarrow wL_2 \rightarrow w_2 \rightarrow wU_2 \rightarrow c_2 \rightarrow wL_3 \rightarrow w_3 \rightarrow wU_3 \rightarrow c_3\)
+        \end{itemize}
+
+    \item[\(H_2\):] is not possible with 2-phase-locking!
+
+        Conflicting excerpt: \(r_1[x] \rightarrow w_2[x] \rightarrow r_1[y]\)
+
+        Either \(T_2\) cannot aquire the write lock for \(w_2[x]\) as \(T_1\) still holds the read lock or \(T_1\) is already in the lock shrinking phase when it wants to execute \(r_1[y]\) (this is the first operation of \(T_1\)on \(y\)) but both transactions commit.
+
+    \item[\(H_3\):] is not possible with 2-phase-locking!
+
+        Conflicting excerpt: \(r_1[y] \rightarrow w_3[y] \rightarrow w_1[z]\)
+
+        Analog situation to \(H_2\). Shrinking for \(T_1\) can only happen after \(w_1[z]\) but \(r_1[y]\) and \(w_3[y]\) are conflicting locks.
+
+\end{itemize}
+
 \end{document}

### 2019-12-14 assignment4/submission4.tex
Fix figure placement and wrong arrows

@@
 \usepackage{enumitem}
+\usepackage{float}
 %\usepackage{minted}
@@
 \item
-$S_{30}$ -> $S_{31}$ = true
-$S_{30}$ -> $S_{32}$ = true
-$S_{30}$ -> $S_{42}$ = true
+$S_{30} \rightarrow S_{31}$ = true\\
+$S_{30} \rightarrow S_{32}$ = true\\
+$S_{30} \rightarrow S_{42}$ = true
 
-$S_{31}$ -> $S_{32}$ = true
-$S_{32}$ -> $S_{42}$ = true
+$S_{31} \rightarrow S_{32}$ = true\\
+$S_{32} \rightarrow S_{42}$ = true\\
+
 So $S_{32}$ and $S_{42}$ are not on level L+1, because $S_{31}$ lies between them and $S_{30}$.\\
@@
 \end{enumerate}
-\subsection*{c)}
+\subsection*{c)} See Figure 1
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[]
 \centering
@@
 \end{enumerate}
+
 \section*{2 - Transaction Processing}
+
 \subsection*{a)}
@@
 
-\begin{figure}[ht!]
+\begin{figure}[H]
+\centering
 \begin{tikzpicture}

### 2019-12-15 assignment4/submission4.tex
Correct typos in exercise 4.1, add linebreaks for simplified diffing

@@
 \item
-The inequation give true for all states that can follow the first one given independant on how many states lie between them.
-To make sure that the other state lies on level L+1 you can either check if any other state lies between them or check the timestamps.
-The stamp of the prosess with tha last event has to be 1 higher than in the other state.
+The inequation gives true for all states that can follow the first one given, independent of how many states lie between them.
+To make sure that the other state lies on level \(L+1\) you can either check if any other state lies between them or check the timestamps.
+The timestamp of the process with that last event has to be 1 higher than in the other state.
 \item
@@
 
-So $S_{32}$ and $S_{42}$ are not on level L+1, because $S_{31}$ lies between them and $S_{30}$.\\
-You either have to check the timestamps (not given) or check all other states if they lie between $S_{30}$ and $S_{31}$. Non of them do so $S_{31}$ is on level l+1 of $S_{30}$.
+So $S_{32}$ and $S_{42}$ are not on level \(L+1\), because $S_{31}$ lies between them and $S_{30}$.\\
+You either have to check the timestamps (not given) or check all other states if they lie between $S_{30}$ and $S_{31}$. 
+None of them do, so $S_{31}$ is on level \(L+1\) of $S_{30}$.
 \end{enumerate}
@@
 \begin{itemize}
-    \item[\(H_1\):] could be generated by 2-phase-locking. The shrinking phase for each transaction begins right after the last operation before the commit.
+    \item[\(H_1\):] could be generated by 2-phase-locking. 
+    The shrinking phase for each transaction begins right after the last operation before the commit.
 
@@
 
-        Analog situation to \(H_2\). Shrinking for \(T_1\) can only happen after \(w_1[z]\) but \(r_1[y]\) and \(w_3[y]\) are conflicting locks.
+        Analogous situation to \(H_2\).
+        Shrinking for \(T_1\) can only happen after \(w_1[z]\) but \(r_1[y]\) and \(w_3[y]\) are conflicting locks.
 

### 2019-12-16 assignment5/A5Q5-Replication-Task/src/main/java/de/uni_stuttgart/ipvs/ids/communication/NonBlockingReceiver.java
add practical ex5

@@ -1,0 +1,44 @@
+package de.uni_stuttgart.ipvs.ids.communication;
+
+import java.io.ByteArrayInputStream;
+import java.io.IOException;
+import java.io.ObjectInputStream;
+import java.net.DatagramPacket;
+import java.net.DatagramSocket;
+import java.net.SocketAddress;
+import java.net.SocketException;
+import java.net.SocketTimeoutException;
+import java.util.Collection;
+import java.util.Vector;
+
+/**
+ * Part b) Extend the method receiveMessages to return all DatagramPackets that
+ * were received during the given timeout.
+ * 
+ * Also implement unpack() to conveniently convert a Collection of
+ * DatagramPackets containing ValueResponseMessages to a collection of
+ * VersionedValueWithSource objects.
+ * 
+ */
+public class NonBlockingReceiver {
+
+	protected DatagramSocket socket;
+
+	public NonBlockingReceiver(DatagramSocket socket) {
+		this.socket = socket;
+	}
+
+	public Vector<DatagramPacket> receiveMessages(int timeoutMillis, int expectedMessages)
+			throws IOException {
+		// TODO: Impelement me!	
+		return null;
+	}
+
+	public static <T> Collection<MessageWithSource<T>> unpack(
+			Collection<DatagramPacket> packetCollection) throws IOException,
+			ClassNotFoundException {
+		// TODO: Impelement me!	
+		return null;
+	}
+	
+}

### 2019-12-16 assignment5/A5Q5-Replication-Task/src/main/java/de/uni_stuttgart/ipvs/ids/replication/MajorityConsensus.java

@@ -1,0 +1,119 @@
+package de.uni_stuttgart.ipvs.ids.replication;
+
+import java.io.ByteArrayOutputStream;
+import java.io.IOException;
+import java.io.ObjectOutputStream;
+import java.net.DatagramPacket;
+import java.net.DatagramSocket;
+import java.net.InetSocketAddress;
+import java.net.SocketAddress;
+import java.net.SocketException;
+import java.net.UnknownHostException;
+import java.util.ArrayList;
+import java.util.Collection;
+
+import de.uni_stuttgart.ipvs.ids.communication.MessageWithSource;
+import de.uni_stuttgart.ipvs.ids.communication.NonBlockingReceiver;
+import de.uni_stuttgart.ipvs.ids.communication.ReadRequestMessage;
+import de.uni_stuttgart.ipvs.ids.communication.ReleaseReadLock;
+import de.uni_stuttgart.ipvs.ids.communication.ReleaseWriteLock;
+import de.uni_stuttgart.ipvs.ids.communication.RequestReadVote;
+import de.uni_stuttgart.ipvs.ids.communication.RequestWriteVote;
+import de.uni_stuttgart.ipvs.ids.communication.ValueResponseMessage;
+import de.uni_stuttgart.ipvs.ids.communication.Vote;
+import de.uni_stuttgart.ipvs.ids.communication.Vote.State;
+import de.uni_stuttgart.ipvs.ids.communication.WriteRequestMessage;
+
+public class MajorityConsensus<T> {
+
+	protected Collection<SocketAddress> replicas;
+
+	protected DatagramSocket socket;
+	protected NonBlockingReceiver nbio;
+
+	final static int TIMEOUT = 1000;
+
+	public MajorityConsensus(Collection<SocketAddress> replicas, int port)
+			throws SocketException {
+		this.replicas = replicas;
+		SocketAddress address = new InetSocketAddress("127.0.0.1", port);
+		this.socket = new DatagramSocket(address);
+		this.nbio = new NonBlockingReceiver(socket);
+	}
+
+	/**
+	 * Part c) Implement this method.
+	 */
+	protected Collection<MessageWithSource<Vote>> requestReadVote() throws QuorumNotReachedException {
+		// TODO: Implement me!
+		return null;
+	}
+	
+	/**
+	 * Part c) Implement this method.
+	 */
+	protected void releaseReadLock(Collection<SocketAddress> lockedReplicas) {
[…]

### 2019-12-16 assignment5/A5Q5-Replication-Task/src/test/java/de/uni_stuttgart/ipvs/ids/test/TestCases.java

@@ -1,0 +1,133 @@
+package de.uni_stuttgart.ipvs.ids.test;
+
+
+import de.uni_stuttgart.ipvs.ids.replication.MajorityConsensus;
+import de.uni_stuttgart.ipvs.ids.replication.QuorumNotReachedException;
+import de.uni_stuttgart.ipvs.ids.replication.Replica;
+import java.io.IOException;
+import java.net.SocketAddress;
+import java.net.SocketException;
+import java.util.ArrayList;
+import java.util.List;
+import org.junit.AfterClass;
+import static org.junit.Assert.*;
+import org.junit.Rule;
+
+import org.junit.Test;
+import org.junit.rules.Timeout;
+
+
+public class TestCases {
+    
+    @Rule
+    public Timeout globalTimeout = Timeout.seconds(30);
+    
+    
+    @Test
+    public void test1() throws SocketException {
+        
+        int replicaPort = 11000;
+        int noReplicas = 10;
+        double prob = 1; // Success Probability
+        double value = 2.0;
+
+        List<Replica<Double>> replicas = new ArrayList<Replica<Double>>(noReplicas);
+        List<SocketAddress> replicaAddrs = new ArrayList<SocketAddress>(replicas.size());
+
+        for (int i = 0; i < noReplicas; i++) {
+            Replica<Double> r = new Replica<Double>(i, replicaPort + i, prob, value);
+            r.start();
+            replicas.add(r);
+            replicaAddrs.add(r.getSocketAddress());
+        }
+
+        MajorityConsensus<Double> mc = new MajorityConsensus<Double>(replicaAddrs, 3404);
+
+        try {
+            double y = mc.get().getValue();
+
+            assertEquals(2.0, y, 0.0);
+            mc.set(3.0);
+            y = mc.get().getValue();
+            assertEquals(3.0, y, 0.0);
+      
+        } catch (QuorumNotReachedException e) {
+            assertTrue(false);
+        }
+    }
+    
+    @Test
+    public void test2() throws SocketException {
+        
+        int replicaPort = 9000;
+        int noReplicas = 10;
+        double prob = 0.001; // Success Probability
+        double value = 2.0;
+
+        List<Replica<Double>> replicas = new ArrayList<Replica<Double>>(noReplicas);
+        List<SocketAddress> replicaAddrs = new ArrayList<SocketAddress>(replicas.size());
+
+        for (int i = 0; i < noReplicas; i++) {
[…]

### 2020-01-05 assignment5/A5Q5-Replication-Task/src/main/java/de/uni_stuttgart/ipvs/ids/replication/Replica.java
Fix replica implementation

@@ -29,9 +29,9 @@ public class Replica<T> extends Thread {
 	private VersionedValue<T> value;
 
 	protected DatagramSocket socket = null;
-	
+
 	protected LockType lock;
-	
+
 	/**
 	 * This address holds the addres of the client holding the lock. This
 	 * variable should be set to NULL every time the lock is set to UNLOCKED.
@@ -47,14 +47,14 @@ public class Replica<T> extends Thread {
 		this.value = new VersionedValue<T>(0, initialValue);
 		this.lock = LockType.UNLOCKED;
 	}
-	
+
 
 	/**
 	 * Part a) Implement this run method to receive and process request
 	 * messages. To simulate a replica that is sometimes unavailable, it should
 	 * randomly discard requests as long as it is not locked.
 	 * The probability for discarding a request is (1 - availability).
-	 * 
+	 *
 	 * For each request received, it must also be checked whether the request is valid.
 	 * For example:
 	 * - Does the requesting client hold the correct lock?
@@ -80,13 +80,14 @@ public class Replica<T> extends Thread {
 
 				//Simulating crashed Nodes
 				//Only happend if node is unlocked
-				if(lock == LockType.UNLOCKED && random.nextDouble()>availability)
+				if(lock == LockType.UNLOCKED && random.nextDouble()>availability) {
 					continue;
+				}
 
 				//Take action based on object type
 				if(msg instanceof RequestReadVote)
 				{
-					if(lock != LockType.WRITELOCK)
+					if(lock == LockType.UNLOCKED)
 					{
 						lock = LockType.READLOCK;
 						lockHolder = packet.getSocketAddress();
@@ -130,7 +131,6 @@ public class Replica<T> extends Thread {
 
 						byte[] byteMsg = bStream.toByteArray();
 
-						DatagramSocket socket = new DatagramSocket();
 						socket.send(new DatagramPacket(byteMsg, byteMsg.length, packet.getSocketAddress()));
 						continue;
 					}
@@ -158,7 +158,7 @@ public class Replica<T> extends Thread {
 					}
 				}
 
-				if(msg instanceof RequestWriteVote)
+				if(msg instanceof ReleaseWriteLock)
 				{
 					if(lock == LockType.WRITELOCK && lockHolder.equals(packet.getSocketAddress()))
 					{
@@ -178,7 +178,7 @@ public class Replica<T> extends Thread {
[…]

### 2020-01-05 assignment5/A5Q5-Replication-Task/src/main/java/de/uni_stuttgart/ipvs/ids/communication/NonBlockingReceiver.java
Fix NonBlockingReceiver implementation

@@ -14,11 +14,11 @@ import java.util.Vector;
 /**
  * Part b) Extend the method receiveMessages to return all DatagramPackets that
  * were received during the given timeout.
- * 
+ *
  * Also implement unpack() to conveniently convert a Collection of
  * DatagramPackets containing ValueResponseMessages to a collection of
  * VersionedValueWithSource objects.
- * 
+ *
  */
 public class NonBlockingReceiver {
 
@@ -29,39 +29,45 @@ public class NonBlockingReceiver {
 	}
 
 	public Vector<DatagramPacket> receiveMessages(int timeoutMillis, int expectedMessages)
-			throws IOException {
+			 {
 
 		Vector<DatagramPacket> ret = new Vector<>();
 
 		//Get the start time
 		//The remaining time for the current timeout will be calculated based on this
 		long start = System.currentTimeMillis();
+		long end = start + timeoutMillis;
 
 		//Repeat while timeout is not reached
-		while(start + timeoutMillis > System.currentTimeMillis())
+		for (long currentTimeout = timeoutMillis; currentTimeout > 0 && ret.size() < expectedMessages; currentTimeout = end - System.currentTimeMillis())
 		{
 			try
 			{
 				byte[] buffer = new byte[2048];
 				DatagramPacket packet = new DatagramPacket(buffer, buffer.length);
-				socket.setSoTimeout((int) (System.currentTimeMillis() - start));
+				socket.setSoTimeout((int) currentTimeout);
 				socket.receive(packet);
 
 				ret.add(packet);
-
-				//All messages received
-				if(ret.size() == expectedMessages)
-					break;
-
 			}
 			catch (SocketTimeoutException e)
 			{
 				//Timeout reached
 				break;
 			}
+			catch (IOException e)
+			{
+				// something went wrong with receiving a packet
+				e.printStackTrace();
+				continue;
+			}
 		}
 		//Set socket back to none timeout mode
-		socket.setSoTimeout(0);
+		try {
+			socket.setSoTimeout(0);
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
 		return ret;
 	}
 
@@ -73,12 +79,12 @@ public class NonBlockingReceiver {
 		for(DatagramPacket packet : packetCollection)
 		{
 			ObjectInputStream iStream = new ObjectInputStream(new ByteArrayInputStream(packet.getData()));
[…]

### 2020-01-05 assignment5/A5Q5-Replication-Task/src/main/java/de/uni_stuttgart/ipvs/ids/replication/MajorityConsensus.java
Add implementation for c)

@@ -1,7 +1,10 @@
 package de.uni_stuttgart.ipvs.ids.replication;
 
+import java.io.ByteArrayInputStream;
 import java.io.ByteArrayOutputStream;
 import java.io.IOException;
+import java.io.ObjectInputStream;
+import java.io.ObjectOutput;
 import java.io.ObjectOutputStream;
 import java.net.DatagramPacket;
 import java.net.DatagramSocket;
@@ -11,6 +14,10 @@ import java.net.SocketException;
 import java.net.UnknownHostException;
 import java.util.ArrayList;
 import java.util.Collection;
+import java.util.HashSet;
+import java.util.Vector;
+import java.util.stream.Collectors;
+import java.util.stream.Stream;
 
 import de.uni_stuttgart.ipvs.ids.communication.MessageWithSource;
 import de.uni_stuttgart.ipvs.ids.communication.NonBlockingReceiver;
@@ -45,17 +52,43 @@ public class MajorityConsensus<T> {
 	 * Part c) Implement this method.
 	 */
 	protected Collection<MessageWithSource<Vote>> requestReadVote() throws QuorumNotReachedException {
-		// TODO: Implement me!
-		return null;
+		RequestReadVote message = new RequestReadVote();
+		try {
+			byte[] byteMsg = this.serializeMessage(message);
+			for (SocketAddress replica: this.replicas) {
+				this.sendMessage(replica, byteMsg);
+			}
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
+		int required = (int) (this.replicas.size() / 2) + 1; // TODO check quorum condition...
+		Vector<DatagramPacket> rawMessages = this.nbio.receiveMessages(5000, this.replicas.size()); // TODO adjust timeout
+		try {
+			Collection<MessageWithSource<Vote>> messages = NonBlockingReceiver.unpack(rawMessages);
+			return this.checkQuorum(messages);
+		} catch (IOException|ClassNotFoundException e) {
+			e.printStackTrace();
+		}
+		throw new QuorumNotReachedException(required, new HashSet<SocketAddress>());
 	}
-	
+
 	/**
 	 * Part c) Implement this method.
 	 */
 	protected void releaseReadLock(Collection<SocketAddress> lockedReplicas) {
-		// TODO: Implement me!
+		ReleaseReadLock message = new ReleaseReadLock();
+		try {
+			byte[] byteMsg = this.serializeMessage(message);
+			for (SocketAddress replica: lockedReplicas) {
[…]

### 2020-01-05 assignment5/A5Q5-Replication-Task/src/main/java/de/uni_stuttgart/ipvs/ids/replication/MajorityConsensus.java
done ex5part2-d)

@@ -33,194 +33,253 @@ import de.uni_stuttgart.ipvs.ids.communication.WriteRequestMessage;
 
 public class MajorityConsensus<T> {
 
-	protected Collection<SocketAddress> replicas;
+    protected Collection<SocketAddress> replicas;
 
-	protected DatagramSocket socket;
-	protected NonBlockingReceiver nbio;
+    protected DatagramSocket socket;
+    protected NonBlockingReceiver nbio;
 
-	final static int TIMEOUT = 1000;
+    final static int TIMEOUT = 1000;
 
-	public MajorityConsensus(Collection<SocketAddress> replicas, int port)
-			throws SocketException {
-		this.replicas = replicas;
-		SocketAddress address = new InetSocketAddress("127.0.0.1", port);
-		this.socket = new DatagramSocket(address);
-		this.nbio = new NonBlockingReceiver(socket);
-	}
+    public MajorityConsensus(Collection<SocketAddress> replicas, int port)
+            throws SocketException {
+        this.replicas = replicas;
+        SocketAddress address = new InetSocketAddress("127.0.0.1", port);
+        this.socket = new DatagramSocket(address);
+        this.nbio = new NonBlockingReceiver(socket);
+    }
 
-	/**
-	 * Part c) Implement this method.
-	 */
-	protected Collection<MessageWithSource<Vote>> requestReadVote() throws QuorumNotReachedException {
-		RequestReadVote message = new RequestReadVote();
-		try {
-			byte[] byteMsg = this.serializeMessage(message);
-			for (SocketAddress replica: this.replicas) {
-				this.sendMessage(replica, byteMsg);
-			}
-		} catch (IOException e) {
-			e.printStackTrace();
-		}
-		int required = (int) (this.replicas.size() / 2) + 1; // TODO check quorum condition...
-		Vector<DatagramPacket> rawMessages = this.nbio.receiveMessages(5000, this.replicas.size()); // TODO adjust timeout
-		try {
-			Collection<MessageWithSource<Vote>> messages = NonBlockingReceiver.unpack(rawMessages);
-			return this.checkQuorum(messages);
-		} catch (IOException|ClassNotFoundException e) {
-			e.printStackTrace();
-		}
-		throw new QuorumNotReachedException(required, new HashSet<SocketAddress>());
-	}
+    /**
+     * Part c) Implement this method.
+     */
+    protected Collection<MessageWithSource<Vote>> requestReadVote() throws QuorumNotReachedException {
[…]

### 2020-01-20 assignment6/submission6.tex
add ex6

--- 2019-12-12 assignment4/submission4.tex
@@
 \usetikzlibrary{positioning}
+\usepackage{marvosym}
 
@@
 \newcommand{\gettitle}{Distributed systems I\\Winter Term 2019/20}
-\newcommand{\getsubtitle}{G2T1 – Assignment 4 (theoretical part)}
+\newcommand{\getsubtitle}{G2T1 – Assignment 6 (theoretical part)}
 \newcommand{\getauthors}{Felix Bühler \and Clemens Lieb \and Steffen Wonner \and Fabian Bühler}
@@
 
-\section*{1 - Global State}
-\subsection*{a)}
-$C_{1}$: $e_{11}$ $e_{21}$ $e_{31}$ $e_{12}$ $e_{22}$\\
-$C_{2}$: $e_{11}$ $e_{21}$ $e_{31}$ $e_{12}$ $e_{22}$ $e_{41}$
+\section*{1 - Replication with Atomic Multicast}
 
-\subsection*{b)}
-\begin{enumerate}[label=(\roman*)]
-\item
-The inequation give true for all states that can follow the first one given independant on how many states lie between them.
-To make sure that the other state lies on level L+1 you can either check if any other state lies between them or check the timestamps.
-The stamp of the prosess with tha last event has to be 1 higher than in the other state.
-\item
-$S_{30}$ -> $S_{31}$ = true
-$S_{30}$ -> $S_{32}$ = true
-$S_{30}$ -> $S_{42}$ = true
+\section*{2 - Multicast Semantics}
 
-$S_{31}$ -> $S_{32}$ = true
-$S_{32}$ -> $S_{42}$ = true
-So $S_{32}$ and $S_{42}$ are not on level L+1, because $S_{31}$ lies between them and $S_{30}$.\\
-You either have to check the timestamps (not given) or check all other states if they lie between $S_{30}$ and $S_{31}$. Non of them do so $S_{31}$ is on level l+1 of $S_{30}$.
-\end{enumerate}
-\subsection*{c)}
-
-\tikzstyle{nofill}=[rectangle,draw, align=center]
-\tikzstyle{green}=[rectangle,draw,fill=green!50,align=center]
-\tikzstyle{blue}=[rectangle,draw,fill=blue!50,align=center]
-\tikzstyle{both}=[rectangle,draw,fill=red!50!,align=center]
-
-\begin{figure}[!ht]
-\centering
-\begin{tikzpicture}[auto,node distance=1.5cm]
-\node[green] (s00) {$ S_{00} $ \\ $ x_1 = 0 $ \\ $ x_2 = 0 $};
-
-\node[both] (s10) [below left = of s00] {$ S_{10} $ \\ $ x_1 = 10 $ \\ $ x_2 = 0 $};
-
-\node[both] (s20) [below left = of s10] {$ S_{20} $ \\ $ x_1 = 5 $ \\ $ x_2 = 0 $};
-\node[green] (s11) [below right = of s10] {$ S_{11} $ \\ $ x_1 = 10 $ \\ $ x_2 = 20 $};
-
-\node[both] (s30) [below left = of s20] {$ S_{30} $ \\ $ x_1 = 15 $ \\ $ x_2 = 0 $};
-\node[green] (s21) [below left = of s11] {$ S_{21} $ \\ $ x_1 = 5 $ \\ $ x_2 = 20 $};
-\node[green] (s12) [below right = of s11] {$ S_{12} $ \\ $ x_1 = 10 $ \\ $ x_2 = 30 $};
-
-\node[nofill] (s31) [below left = of s21] {$ S_{31} $ \\ $ x_1 = 15 $ \\ $ x_2 = 20 $};
-\node[green] (s22) [below left = of s12] {$ S_{22} $ \\ $ x_1 = 5 $ \\ $ x_2 = 30 $};
-
-\node[green] (s32) [below left = of s22] {$ S_{32} $ \\ $ x_1 = 15 $ \\ $ x_2 = 30 $};
-
-\node[both] (s42) [below left = of s32] {$ S_{42} $ \\ $ x_1 = 55 $ \\ $ x_2 = 30 $};
-
-\node[both] (s52) [below left = of s42] {$ S_{52} $ \\ $ x_1 = 60 $ \\ $ x_2 = 30 $};
-
-\node[both] (s53) [below right = of s52] {$ S_{53} $ \\ $ x_1 = 60 $ \\ $ x_2 = 13 $};
-
-\path (s10) edge node {} (s00);
-
-\path (s10) edge node {} (s11);
-\path (s10) edge node {} (s20);
-
-\path (s11) edge node {} (s21);
-\path (s20) edge node {} (s21);
-\path (s20) edge node {} (s30);
-\path (s11) edge node {} (s12);
-
-\path (s30) edge node {} (s31);
-\path (s21) edge node {} (s31);
-\path (s21) edge node {} (s22);
-\path (s12) edge node {} (s22);
-
-\path (s31) edge node {} (s32);
-\path (s22) edge node {} (s32);
-
-\path (s32) edge node {} (s42);
-
-\path (s42) edge node {} (s52);
-
-\path (s52) edge node {} (s53);
-\end{tikzpicture}
-\caption{solution for 1.c)}
-\end{figure}
-
-\subsection*{d)}
-\begin{itemize}
-	\item $ \phi_{1} = (x_{2} - x_{1}) = 5 $\\
-	$ \Rightarrow (\neg\phi_{1}) $ marked in \textcolor{blue}{blue}
-	\item $ \phi_{2} = (x_{2} - x_{1}) \leq 5 $ and $ (x_{2} - x_{1}) > 0 $\\
-	$ \Rightarrow (\phi_{2}) $ marked in \textcolor{green}{green}
-	\item if they fulfill $ (\neg\phi_{1}) $ and $ (\phi_{2}) $ marked in \textcolor{red}{red}
-\end{itemize}
-
-\begin{enumerate}
-	\item safety condition $ \neg\phi_{1} $ is not fulfilled in the state $ S_{31} $ and thereby not fulfilled.
-	\item liveness condition $ \phi_{2} $ is fulfilled in the end and thereby fulfilled.
-\end{enumerate}
-\section*{2 - Transaction Processing}
-\subsection*{a)}
-The conflicting operation pairs are:
-\[\{(w_1[u], w_3[u]), (r_1[y], w_2[y]), (r_3[y], w_2[y]), (r_2[x], w_3[x])\}\]
-
-\subsection*{b)}
-
-\begin{figure}[ht!]
-\begin{tikzpicture}
-    \node[draw, circle] at (0, 0)    (T1) {\(T_1\)};
-    \node[draw, circle] at (5, 0)    (T2) {\(T_2\)};
-    \node[draw, circle] at (2.5,2.5) (T3) {\(T_3\)};
-
-    \draw[->]
-        (T1) edge node[above,sloped] {\(w_1[y] \to r_2[y]\)} (T2)
-        (T2) edge node[above,sloped] {\(w_2[z] \to w_3[z]\)} (T3)
-        (T3) edge node[above,sloped] {\(w_3[y] \to r_1[y]\)} (T1);
-\end{tikzpicture}
-\caption{Serialization Graph for History \(H_1\)}
-\label{fig:sg_h1}
-\end{figure}
-
-It's clear that the serialization graph shown in Figure \ref{fig:sg_h1} contains a cycle, as such \(H_1\) is not serializable.\\~\\
-
-\begin{figure}[ht!]
-\begin{tikzpicture}
-    \node[draw, circle] at (0, 0)    (T1) {\(T_1\)};
-    \node[draw, circle] at (5, 0)    (T2) {\(T_2\)};
-    \node[draw, circle] at (2.5,2.5) (T3) {\(T_3\)};
-
-    \draw[->]
-        (T2) edge node[above,sloped] {\(r_2[y] \to w_1[y]\)} (T1)
-        (T3) edge node[above,sloped] {\(r_3[x] \to w_1[x]\)} (T1)
-        (T3) edge node[above,sloped] {\parbox[c]{2.5cm}{\(r_3[x] \to w_2[x] \\ w_3[u] \to w_2[u]\)}} (T2);
-\end{tikzpicture}
-\caption{Serialization Graph for History \(H_2\)}
-\label{fig:sg_h2}
-\end{figure}
-
-The serialization graph shown in Figure \ref{fig:sg_h2} does not contain a cycle, which would imply that \(H_2\) is serializable.\\
-Unfortunately \(H_2\) is not a correct history as it does not specify the order or the conflicting operations \(w_1[x], w_2[x]\).
-
-\section*{3 - Two-Phase Locking}
-
-\begin{itemize}
-    \item[\(H_1\):] could be generated by 2-phase-locking. The shrinking phase for each transaction begins right after the last operation before the commit.
-
-        Execution split for each resource:
-
-        \begin{itemize}
-            \item[\(y\):] \(rL_1 \rightarrow r_1 \rightarrow wL_1 \rightarrow w_1 \rightarrow wU_1 \rightarrow rL_2 \rightarrow r_2 \rightarrow c_1 \rightarrow rU_2 \rightarrow c_2 \rightarrow c_3\)
-            \item[\(x\):] only read operations $\Rightarrow$ no conflicting locks possible
-            \item[\(z\):] only read operations $\Rightarrow$ no conflicting locks possible
-            \item[\(w\):] \(c_1 \rightarrow wL_2 \rightarrow w_2 \rightarrow wU_2 \rightarrow c_2 \rightarrow wL_3 \rightarrow w_3 \rightarrow wU_3 \rightarrow c_3\)
-        \end{itemize}
-
-    \item[\(H_2\):] is not possible with 2-phase-locking!
-
-        Conflicting excerpt: \(r_1[x] \rightarrow w_2[x] \rightarrow r_1[y]\)
-
-        Either \(T_2\) cannot aquire the write lock for \(w_2[x]\) as \(T_1\) still holds the read lock or \(T_1\) is already in the lock shrinking phase when it wants to execute \(r_1[y]\) (this is the first operation of \(T_1\)on \(y\)) but both transactions commit.
-
-    \item[\(H_3\):] is not possible with 2-phase-locking!
-
-        Conflicting excerpt: \(r_1[y] \rightarrow w_3[y] \rightarrow w_1[z]\)
-
-        Analog situation to \(H_2\). Shrinking for \(T_1\) can only happen after \(w_1[z]\) but \(r_1[y]\) and \(w_3[y]\) are conflicting locks.
-
-\end{itemize}
+\section*{3 - Causal Mulitcast}
 

### 2020-01-23 assignment6/submission6.tex
done 6.3a)
correct 6.3.a)
add drawing for 5.3.b)
Add 6.1 a) b) and c)

@@
 
+\subsection*{a)}
+
+As shown in figure \ref{fig:1a} message M1 is delivered by R1 but not R2.
+This can happen if the sender is faulty as the validity property is only enforced for correct senders.
+If the agreement property is not enforced R2 may never receive the message M1.
+This means that x is not replicated and R2 has a stale value even though it is still correct.
+
+R2 will however not deliver any other message since it still expects a message with the sequence number of M1 to arrive before delivering any message with a higher sequence number.
+
+\begin{figure}[!ht]
+    \centering
+    \begin{tikzpicture}[deliver/.style={fill=black,circle, inner sep=0pt, minimum size=5pt}, label/.style={font=\scriptsize}, msg/.style={->}]
+    \node (C1) {C1};
+    \node[below=1cm of C1] (R1) {R1};
+    \node[below=1cm of R1] (R2) {R2};
+
+    \node[right=6cm of C1] (C1end) {};
+    \node[right=6cm of R1] (R1end) {};
+    \node[right=6cm of R2] (R2end) {};
+
+    \path[draw, ->]
+        (C1) edge (C1end)
+        (R1) edge (R1end)
+        (R2) edge (R2end)
+        ;
+
+    \node(m1s) [right=0.5cm of C1] {};
+    \node[deliver] (m1r1) [right=2cm of R1] {};
+    \node[] (m1r2) [below=0.5cm of m1r1] {\Lightning};
+    \node[] (fault) [right=0.5cm of m1s] {\Lightning};
+
+    \path (m1s.center) edge[msg] node[label] {M1 = write(x)} (m1r1);
+    \path (m1s.center) edge[msg] (m1r2);
+
+    \end{tikzpicture}
+    \caption{Atomic Multicast without Agreement}
+    \label{fig:1a}
+\end{figure}
+
+\subsection*{b)}
+
+Assuming the atomicity property is the total order property from the lecture.
+
+As seen in figure \ref{fig:1b} if the total order of messages is not enforced the actual order messages are delivered at each replica may be different.
+Now both replicas have a different value for x.
+
+\begin{figure}[!ht]
+    \centering
+    \begin{tikzpicture}[deliver/.style={fill=black,circle, inner sep=0pt, minimum size=5pt}, label/.style={font=\scriptsize}, msg/.style={->}]
+    \node (C1) {C1};
+    \node[below=1cm of C1] (C2) {C2};
+    \node[below=1cm of C2] (R1) {R1};
+    \node[below=1cm of R1] (R2) {R2};
+
+    \node[right=6cm of C1] (C1end) {};
+    \node[right=6cm of C2] (C2end) {};
+    \node[right=6cm of R1] (R1end) {};
+    \node[right=6cm of R2] (R2end) {};
+
+    \path[draw, ->]
+        (C1) edge (C1end)
+        (C2) edge (C2end)
+        (R1) edge (R1end)
+        (R2) edge (R2end)
+        ;
+
+    \node(m1s) [right=0.5cm of C1] {};
+    \node[label] (M1) [below=0.5cm of m1s.east] {M1 = write(x)};
+    \node[deliver] (m1r1) [right=4cm of R1] {};
+    \node[deliver] (m1r2) [right=1.5cm of R2] {};
+
+    \path (m1s.center) edge[msg] (m1r1);
+    \path (m1s.center) edge[msg] (m1r2);
+
+    \node(m2s) [right=1.3cm of C2] {};
+    \node[label] (M2) [below right=0.5cm and -0.3cm of m2s] {M2 = write(x)};
+    \node[deliver] (m2r1) [right=2.5cm of R1] {};
+    \node[deliver] (m2r2) [right=5cm of R2] {};
+
+    \path (m2s.center) edge[msg] (m2r1);
+    \path (m2s.center) edge[msg] (m2r2);
+
+    \end{tikzpicture}
+    \caption{Atomic Multicast without Atomicity/Total Order}
+    \label{fig:1b}
+\end{figure}
+
+\subsection*{c)}
+
+\subsubsection*{i.}
+As seen in figure \ref{fig:1c} if the client does not wait for an acknowledgement for its messages a newer message may get overwritten by an older message that was not yet delivered.
+The total order property only guaranties that the order of delivered messages is the same for all replicas.
+
+\begin{figure}[!ht]
+    \centering
+    \begin{tikzpicture}[deliver/.style={fill=black,circle, inner sep=0pt, minimum size=5pt}, label/.style={font=\scriptsize}, msg/.style={->}]
+    \node (C1) {C1};
+    \node[below=1cm of C1] (R1) {R1};
+    \node[below=1cm of R1] (R2) {R2};
+
+    \node[right=6cm of C1] (C1end) {};
+    \node[right=6cm of R1] (R1end) {};
+    \node[right=6cm of R2] (R2end) {};
+
+    \path[draw, ->]
+        (C1) edge (C1end)
+        (R1) edge (R1end)
+        (R2) edge (R2end)
+        ;
+
+    \node(m1s) [right=0.5cm of C1] {};
+    \node[label] (M1) [below=0.3cm of m1s.east] {M1 = write(x)};
+    \node[deliver] (m1r1) [right=4cm of R1] {};
+    \node[deliver] (m1r2) [right=5cm of R2] {};
+
+    \path (m1s.center) edge[msg] (m1r1);
+    \path (m1s.center) edge[msg] (m1r2);
+
+    \node(m2s) [right=2cm of C1] {};
+    \node[label] (M2) [below right=0.1cm and -0.3cm of m2s] {M2 = write(x)};
+    \node[deliver] (m2r1) [right=3cm of R1] {};
+    \node[deliver] (m2r2) [right=3cm of R2] {};
+
+    \path (m2s.center) edge[msg] (m2r1);
+    \path (m2s.center) edge[msg] (m2r2);
+
+    \end{tikzpicture}
+    \caption{Atomic Multicast without synchronous clients}
+    \label{fig:1c}
+\end{figure}
+
+\subsubsection*{ii.}
+
+As causal atomic multicast also includes fifo ordering the wrong arrival order of messages when sending them asynchronously like in figure \ref{fig:1c} cannot happen.
+Messages from the same sender will arrive in the order they were sent.
+
 \section*{2 - Multicast Semantics}
@@
 
+\subsection*{a)}
+\subsubsection*{i.}
+See figure \ref{fig:3a}
+
+\begin{figure}[!ht]
+	\centering
+	\begin{tikzpicture}
+	\draw[->, thick] (0, 6) -- (12, 6);
+	\draw[->, thick] (0, 3) -- (12, 3);
+	\draw[->, thick] (0, 0) -- (12, 0);
+
+	\foreach \t/\p in {1/6,2/3,3/0} {
+		\draw node at (-0.5, \p) {\(P_\t\)};
+	};
+
+	\draw[->, thick] (0.5, 6) -- node[pos=0.1, right] {$ m_1 $} (6, 0);
+	\draw[->, thick] (0.5, 6) -- (1, 3);
+
+	\draw[->, thick] (2.5, 3) -- (5, 6);
+	\draw[->, thick] (2.5, 3) -- node[pos=0.3, left] {$ m_2 $} (4, 0);
+	\draw[->, blue, thick] (4, 0) to [out=30,in=150] (7, 0);
+
+	\draw[->, thick] (6.5, 3) -- node[pos=0.3, left] {$ m_3 $} (11, 6);
+	\draw[->, thick] (6.5, 3) -- (10, 0);
+
+	\draw[->, thick] (8, 0) -- (9.5, 3);
+	\draw[->, thick] (8, 0) -- node[pos=0.1, left] {$ m_4 $} (9.5, 6);
+
+	\draw node at (0.5, 6.3) {$ [1,0,0] $};
+	\draw node at (5, 6.3) {$ [1,1,0] $};
+	\draw node at (9.5, 6.3) {$ [1,1,1] $};
+	\draw node at (11, 6.3) {$ [1,2,1] $};
+
+	\draw node at (1, 2.7) {$ [1,0,0] $};
+	\draw node at (2, 3.3) {$ [1,1,0] $};
+	\draw node at (6, 3.3) {$ [1,2,0] $};
+	\draw node at (9.5, 3.3) {$ [1,2,1] $};
+
+	\draw node at (4, -0.3) {$ [0,0,0] $};
+	\draw node at (6, -0.3) {$ [1,0,0] $};
+	\draw node[blue] at (7, 0.3) {$ [1,1,0] $};
+	\draw node at (8, -0.3) {$ [1,1,1] $};
+	\draw node at (10, -0.3) {$ [1,2,1] $};
+
+	\end{tikzpicture}
+	\caption{Execution of CBCAST Algorithm}
+	\label{fig:3a}
+\end{figure}
+
+\subsubsection*{ii.}
+
+The delivered message $ m_2 $ on $ P_3 $ has to be delayed. The \textcolor{blue}{blue} vector clock would be updated to $ [1,1,0] $ after the recieved $ m_1 $ message.
+
+\subsection*{b)}
+\subsubsection*{i.}
+See figure \ref{fig:3b}
+
+\begin{figure}[!ht]
+	\centering
+	\begin{tikzpicture}
+	\draw[->, thick] (0, 6) -- (12, 6);
+	\draw[->, thick] (0, 3) -- (12, 3);
+	\draw[->, thick] (0, 0) -- (12, 0);
+
+	\foreach \t/\p in {1/6,2/3,3/0} {
+		\draw node at (-0.5, \p) {\(P_\t\)};
+	};
+
+	\draw[->, thick] (1, 6) -- node[pos=0.1, sloped, above] {$ m_1 $} (10, 0);
+	\draw[->, thick] (1, 6) -- node[pos=0.3, sloped, above] {$ m_1 $} (2, 3);
+
+	\draw[->, thick] (3, 6) -- node[pos=0.1, sloped, above] {$ m_2 $} (11, 0);
+	\draw[->, thick] (3, 6) -- node[pos=0.3, sloped, above] {$ m_2 $} (4, 3);
+
+	\draw[->, thick] (3, 3) -- node[pos=0.7, sloped, above] {$ < m_1||m_3> $} (7, 6);
+	\draw[->, thick] (3, 3) -- node[pos=0.6, sloped, above] {$ < m_1||m_3> $} (9, 0);
+
+	\draw[->, thick] (4.5, 3) -- node[pos=0.6, sloped, above] {$ < m_2||m_4> $} (11, 6);
+	\draw[->, thick] (4.5, 3) -- node[pos=0.7, sloped, below] {$ < m_2||m_4> $} (7, 0);
+	\end{tikzpicture}
+	\caption{Execution of Algorithm 1}
+	\label{fig:3b}
+\end{figure}
+\subsubsection*{ii.} Otherwise the total ordering can be violated, and therefore the correctness.
+\subsubsection*{iii.}
+
 \end{document}

### 2020-01-24 assignment6/submission6.tex
Added Task 2 solution

@@
 
+\subsection*{a)}
+
+\subsubsection*{i.}
+False\\
+m$_3$ is delivered to P$_$2 but not to P$_1$ and P$_3$ which are correct.
+\subsubsection*{ii.}
+True\\
+m$_1$, m$_2$ and m$_4$ are send by correct processes and delivered exactly once by P$_1$ and P$_3$
+\subsubsection*{iii.}
+True\\
+Only P$_1$ sends more than 1 messaage. The second one is send after all processes recieved the first one. So it is FIFO multicast.\\
+P$_1$ and P$_2$ are delivering m$_1$ and m$_2$ in diffrent orders. So no atomic multicast.
+\subsubsection*{iv.}
+False\\
+P$_2$ does not deliver m$_1$ but m$_4$, that violates Uniform FIFO.
+\subsection*{b)}
+
+\subsubsection*{i.}
+Reliable Multicast: fullfilled\\
+FIFO Multicast: violated\\
+Causal Multicast: violated\\
+Atomic Multicast: violated\\
+\subsubsection*{ii.}
+Reliable Multicast: violated\\
+FIFO Multicast: violated\\
+Causal Multicast: violated\\
+Atomic Multicast: violated\\
+\subsubsection*{iii.}
+Reliable Multicast: fullfilled\\
+FIFO Multicast: fullfilled\\
+Causal Multicast: fullfilled\\
+Atomic Multicast: violated\\
 \section*{3 - Causal Mulitcast}

### 2020-01-26 assignment6/submission6.tex
done 6.3.c

--- 2020-01-23 assignment6/submission6.tex
@@
 \end{figure}
-\subsubsection*{ii.} Otherwise the total ordering can be violated, and therefore the correctness.
-\subsubsection*{iii.}
+\subsubsection*{ii.} As seen in figure \ref{fig:3b} the fifo ordering can be violated, and therefore the correctness.
+\subsubsection*{iii.} Append the last sent message as causal dependency to the new message. The reciever can then guarantee fifo ordering again.
 
